### 2020-03-10 hb.tex
syntax updates. NEEDS proofreading

\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
\usepackage{hyperref}
\usepackage{minted} % pip3 install pygments, pdflatex -shell-escape hb
\newminted[elpicode]{elpi.py:ElpiLexer -x}{linenos=true,fontsize=\footnotesize,escapeinside=\%\%}
\newmintinline[elpi]{elpi.py:ElpiLexer -x}{fontsize=\small}
\newminted[coqcode]{elpi.py:CoqElpiLexer -x}{linenos=true,fontsize=\footnotesize,mathescape,escapeinside=\#\#}
\newmintinline[coq]{elpi.py:CoqElpiLexer -x}{fontsize=\small}
\usepackage{xcolor}
\usepackage{wrapfig}

\definecolor{dkgreen}{rgb}{0, 0.502, 0}

\newcommand{\coqelpi}{Coq-Elpi}
\newcommand{\HB}{\ensuremath{\mathcal{HB}}}
\newcommand{\hb}{\coq{hierarchy-builder}}

%This is a template for producing LIPIcs articles.
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling cleveref support, use "autoref"

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Hierarchy Builder: algebraic~hierarchies made easy in Coq with Elpi} %TODO Please add

\titlerunning{Hierarchy Builder}%optional, please use if title is longer than one line
\author{Cyril Cohen}{Inria, Universit\'e C\^ote d'Azur, France}{Cyril.Cohen@inria.fr}{}{}
\author{Kazuhiko Sakaguchi}{University of Tsukuba, Japan}{sakaguchi@logic.cs.tsukuba.ac.jp}{}{}
\author{Enrico Tassi}{Inria, Universit\'e C\^ote d'Azur, France}{Enrico.Tassi@inria.fr}{}{}
\authorrunning{C.\,Cohen and K.\,Sakaguchi and E.\,Tassi}
% \author{John Q. Public}{Dummy University Computing Laboratory, [optional: Address], Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional
% \author{Joan R. Public\footnote{Optional footnote, e.g. to mark corresponding author}}{Department of Informatics, Dummy College, [optional: Address], Country}{joanrpublic@dummycollege.org}{[orcid]}{[funding]}
% \authorrunning{J.\,Q. Public and J.\,R. Public}%TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Cyril Cohen and Kazuhiko Sakaguchi and Enrico Tassi}%TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{General and reference~General literature}
\ccsdesc[100]{General and reference}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
\supplement{Source code of the Coq package: \url{https://github.com/math-comp/hierarchy-builder}}
\keywords{Algebraic Hierarchy, Packed Classes, Coq, Elpi, Metaprogramming, $\lambda$Prolog}%TODO mandatory; please add comma-separated list of keywords

\category{}%optional, e.g. invited paper

\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% TODO: Write glossary
% remove distinction
% TODO: Structure -> Class
% TODO: packager -> ClassBuilder
% TODO: Builder -> Builder
% Factories ( primitive / general ) ; kill mixin

\newcommand{\Lang}{\ensuremath{\Lambda}}
\newcommand{\term}{term}
\newcommand{\mixin}{mixin}
\newcommand{\mixins}{mixins}
\newcommand{\Mixins}{Mixins}
\newcommand{\M}{\ensuremath{\mathcal{M}}}
\newcommand{\factory}{factory}
\newcommand{\factories}{factories}
\newcommand{\Factories}{Factories}
% \newcommand{\packager}{packager}
% \newcommand{\packagers}{packagers}
% \newcommand{\Packager}{Packager}
\newcommand{\phantterm}{abbreviation}
\newcommand{\phantterms}{abbreviations}
\newcommand{\Phantterms}{abbreviations}
\newcommand{\builder}{builder}
\newcommand{\mixinbuilder}{builder}
\newcommand{\mixinbuilders}{builders}
\newcommand{\factoryinstance}{factory instance}
\newcommand{\factoryinstances}{factory instances}
\newcommand{\F}{\ensuremath{\mathcal{F}}}
\newcommand{\dep}{\ensuremath{\mathit{dep}}}
\newcommand{\powerset}[1]{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}(#1)}
\newcommand{\C}{\ensuremath{\mathcal{C}}}
\newcommand{\B}{\ensuremath{\mathcal{B}}}
\newcommand{\class}{class}
\newcommand{\classes}{classes}
\newcommand{\cdef}{\ensuremath{\mathit{def}}}
\newcommand{\Str}{\ensuremath{\mathcal{S}}}
\newcommand{\structure}{structure}
\newcommand{\structures}{structures}
\newcommand{\issubclass}{\ensuremath{\leq}}
\newcommand{\subclass}{subclass}
\newcommand{\join}{\ensuremath{\mathit{join}}}
\newcommand{\requires}{\ensuremath{\mathit{requires}}}
\newcommand{\target}{\ensuremath{\mathit{target}}}
\newcommand{\provides}{\ensuremath{\mathit{provides}}}
\newcommand{\from}{\ensuremath{\mathsf{from}}}
\newcommand{\proj}[2]{\ensuremath{#1[#2]}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\enum}[2]{\ensuremath{\set{#1,\ldots,#2}}}
\newcommand{\vect}[1]{\overrightarrow{#1}}
\newcommand{\pmp}[1]{\ensuremath{\vect{\left(p : m\ T\ p_\sigma\right)}^{#1}}}
\newcommand{\hbmixin}{{\tt\color{dkgreen}HB.mixin}}
\newcommand{\hbfactory}{{\tt\color{dkgreen}HB.factory}}
\newcommand{\hbinstance}{{\tt\color{dkgreen}HB.instance}}
\newcommand{\hbstructure}{{\tt\color{dkgreen}HB.structure}}
\newcommand{\hbbuilders}{{\tt\color{dkgreen}HB.builders}}
\newcommand{\hbend}{{\tt\color{dkgreen}HB.end}}

\newtheoremstyle{implem}{\topsep}{\topsep}{}{0pt}{\sffamily\bfseries}{ }{5pt plus 1pt minus 1pt}%
{$\blacktriangleright$ \thmname{#1}\thmnumber{ #2}\thmnote{ #3}}

\theoremstyle{implem}
\newtheorem*{implementation}{Implementation of}

\theoremstyle{implem}
\newtheorem*{invariant}{Programming invariant for}

\newtheoremstyle{command}{\topsep}{\topsep}{}{0pt}{\sffamily\bfseries}{ }{5pt plus 1pt minus 1pt}%
{$\blacktriangleright$ \thmname{#1} {\coq{#3}}}
\theoremstyle{command}
\newtheorem*{command}{Coq command}

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
It is nowadays customary to organize libraries of machine checked
proofs around hierarchies of algebraic
structures~\cite{DBLP:conf/mpc/AffeldtNS19,DBLP:journals/mics/BoldoLM15,Cohen_phd,Holzl:2013,10.1145/3372885.3373824,Rouhling_phd,mathclasses}.
One influential example is the Mathematical Components library on top
of which the long and intricate proof of the Odd Order
Theorem could be fully formalized~\cite{DBLP:conf/itp/GonthierAABCGRMOBPRSTT13}.

Still, building algebraic hierarchies in a proof assistant such as Coq~\cite{Coq:manual}
requires a lot of manual labor and often a deep expertise in the internals of
the prover~\cite{DBLP:conf/tphol/GarillotGMR09,DBLP:conf/itp/MahboubiT13}.
Moreover, according to our experience~\cite{KSdraft},
making a hierarchy evolve without causing breakage in client code is equally tricky:
even a simple refactoring such as splitting a structure into two simpler ones
is hard to get right.

In this paper we describe \HB{}, a high level language
to \emph{build} hierarchies of algebraic structures and to make these hierarchies
\emph{evolve} without breaking user code. The key concepts are the ones of
\emph{\factory{}}, \emph{\mixinbuilder{}} and \emph{\phantterm{}} that let the hierarchy developer
describe an actual
interface for their library. Behind that interface the developer can provide
appropriate code to ensure backward compatibility.
We implement the \HB{} language in the \hb{} addon for the Coq
system using the Elpi~\cite{DBLP:conf/lpar/DunchevGCT15,CoqElpi}
extension language.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Modern libraries of machine checked proofs are organized around
hierarchies of algebraic structures~\cite{DBLP:conf/mpc/AffeldtNS19,DBLP:journals/mics/BoldoLM15,Cohen_phd,Holzl:2013,10.1145/3372885.3373824,Rouhling_phd,mathclasses}.
For example the Mathematical Components library for the Coq system~\cite{Coq:manual}
provides a very rich, ever growing, hierarchy of structures such as
group, ring, module, algebra, field, partial order, order, lattice\ldots
The hierarchy does not only serve the purpose of organizing knowledge, but
also to make it easy to exploit it. Indeed the interactive prover can
take advantage of the structure of the library and the relation between
its concepts to infer part of information usually left implicit
by the user, a capability that turned out to be key to tame
the complexity and size of the formal proof of the Odd Order
Theorem~\cite{DBLP:conf/itp/GonthierAABCGRMOBPRSTT13}.

The hierarchy of the Mathematical Components library
is implemented following the discipline of Packed Classes initially introduced
in \cite{DBLP:conf/tphol/GarillotGMR09} and later also adopted in
\cite{DBLP:conf/mpc/AffeldtNS19} and \cite{DBLP:journals/mics/BoldoLM15}.
We call Packed Classes a discipline, and not a language, because, in spite of
its many virtues, it is unwieldy to use. In particular it leaks to the user
many of the technical details of the Coq system. As a result, one
needs to be a Coq expert in order to build or modify a hierarchy, and even
experts make mistakes as shown in~\cite{KSdraft}.
Another inconvenience of the Packed Classes discipline is that even
simple changes to the hierarchy, such as splitting a structure into
two simpler ones, break user code.

In this paper we describe \HB{}, a high level language
to \emph{build} hierarchies of algebraic structures and to make these hierarchies
\emph{evolve} without breaking user code. The key concepts are the ones of
\emph{\factory{}} and \emph{\phantterm{}} that let the hierarchy developer describe an actual
interface for their library. Behind that interface the developer can provide
appropriate code to ensure backward compatibility.
We implement the \HB{} language by compiling it
to a variant of the Packed Classes discipline, that we call \emph{flat},
in the \hb{} addon for Coq. We write this addon using the
Elpi~\cite{DBLP:conf/lpar/DunchevGCT15,CoqElpi} extension language.\\

\noindent To sum up, the main contributions of the paper are:
\begin{itemize}
\item the design of the \HB{} language,
\item the compilation of \HB{} to the (flat) Packed Classes discipline, and
\item the implementation of \HB{} in the \hb{} addon for Coq.
\end{itemize}
The paper is organized as follows. Via an example we introduce
\HB{} and its key ideas. %We then provide a formal description of \HB{}.
We then describe the discipline
of Packed Classes and we show how \HB{} can be compiled to it.
We then discuss the implementation of the Coq addon via the Elpi extension
language and we position \HB{} in the literature.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\section{\HB{} by examples: building and evolving a hierarchy}
\label{sec:example}

The first version of our hierarchy (that we name V1) features only two
structures: \coq{Monoid} and \coq{Ring}.
%\marginpar{remove hb.end for \mixins{}, use next release of coq elpi}
\begin{coqcode}
#\hbmixin{}# Record Monoid_of_Type A := {                       #\label{demo:mixin:mot}#
  zero : A;
  add : A -> A -> A;
  addrA : associative add;             (* `add` is associative.                   *)
  add0r : left_id zero add;            (* `zero` is the left and right neutral    *)
  addr0 : right_id zero add;           (*   element with respect to `add`.        *)
}.                                                             #\label{demo:mixin:mot:end}#
#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type.axioms A }.        #\label{demo:structure:monoid}#

#\hbmixin{}# Record Ring_of_Monoid A of Monoid.axioms A := {       #\label{demo:mixin:rom}#
  one : A;
  opp : A -> A;
  mul : A -> A -> A;
  addNr : left_inverse zero opp add;   (* `opp x` is the left and right additive  *)
  addrN : right_inverse zero opp add;  (*   inverse of `x`.                       *)
  mulrA : associative mul;             (* `mul` is associative.                   *)
  mul1r : left_id one mul;             (* `one` is the left and right neutral     *)
  mulr1 : right_id one mul;            (*   element with respect to `mul`.        *)
  mulrDl : left_distributive mul add;  (* `mul` is left and right distributive    *)
  mulrDr : right_distributive mul add; (*   over `add`.                           *)
}.
#\hbstructure{}# Definition Ring := { A of Monoid.axioms A & Ring_of_Monoid.axioms A }. #\label{demo:structure:ring}#

\end{coqcode}

In order to build a structure we need to declare some \factories{} and
later assemble them. One kind of \factory{} supported by \HB{}, the simplest
one, is called \emph{\mixin{}} and is embodied by a record, by convention named
\coq{axioms}, that gathers operations and properties.

\Mixins{} are declared via the \hbmixin{} command
that takes a module name,
a type variable name and a possibly empty list of \factories{} for that type.
The code between lines~\ref{demo:mixin:mot}
and~\ref{demo:mixin:mot:end} declares a \mixin{} that can turn a naked
type \coq{A} into a monoid, hence we chose the module name to be \coq{Monoid_of_Type}.

The \hbstructure{} command takes in input a name S and a possibly
empty list of \factories{}.
It registers a structure S in the hierarchy placing
any definition specific to that structure inside a Coq module named \coq{S}.
Line~\ref{demo:structure:monoid} hence forges the \coq{Monoid} structure.

Line~\ref{demo:mixin:rom} declares a second \mixin{} collecting the operations
and properties that are needed in order to enrich a monoid to a ring, hence
the name \coq{Ring_of_Monoid}. Indeed this time the type variable \coq{A}
is followed by \coq{Monoid.axioms} that enriches
\coq{A} with the operations and properties of monoids. As a consequence
\coq{add} and \coq{zero} can be used to express the new properties.

The last line declares the \coq{Ring} structure to hold all the axioms declared
so far. We can now inspect the contents of the hierarchy and then
proceed to build a theory about abstract rings,
register examples (instances) of ring structures
and finally use the abstract theory on these examples.

\begin{coqcode}
Print Monoid.type. (* Monoid.type  :=  { sort : Type;  ... }                           *) #\label{demo:theory:print:type}#
Check @add.        (* add          :   forall M : Monoid.type, M -> M -> M             *) #\label{demo:theory:check:add}#
Check @addNr.      (* addNr        :   forall R : Ring.type, left_inverse zero opp add *) #\label{demo:theory:check:addNr}#

Lemma addrC : forall R : Ring.type, commutative (@add R).                       #\label{demo:theory:state:addrC}#
Proof. (* Proof by Hankel 1867, omitted for brevity *) Qed.

Definition Z_Monoid_axioms : Monoid_of_Type.axioms Z :=                         #\label{demo:theory:z:monoid:axioms}#
  Monoid_of_Type.Axioms Z 0 Z.add Z.add_assoc Z.add_0_l Z.add_0_r.

#\hbinstance{}# Z Z_Monoid_axioms.                                            #\label{demo:theory:z:monoid:canonical}#

Definition Z_Ring_axioms : Ring_of_Monoid.axioms Z _ :=                        #\label{demo:theory:z:ring:axioms}#
  Ring_of_Monoid.Axioms Z 1 Z.opp Z.mul
    Z.add_opp_diag_l Z.add_opp_diag_r Z.mul_assoc Z.mul_1_l Z.mul_1_r
    Z.mul_add_distr_r Z.mul_add_distr_l.

#\hbinstance{}# Z Z_Ring_axioms.                                              #\label{demo:theory:z:ring:canonical}#

Lemma exercise (m n : Z) : (n + m) - n * 1 = m.
Proof. by rewrite mulr1 (addrC n) -(addrA m) addrN addr0. Qed.
\end{coqcode}

We can print the type for monoids as forged by \HB{} (line~\ref{demo:theory:print:type}).
It packs a carrier, called \coq{sort}, and the collection of operation and
properties that we omit for brevity.  We can also look at the type of two
constants synthesized by \HB{} out of the hierarchy declaration. Remark that
while the names of the constants come from the names of \mixin{} fields,
their types differ.
In particular they are quantified over a \coq{Monoid.type} or \coq{Ring.type},
and not a simple type \coq{A} as in the \mixins{}. Moreover we evince that
the \coq{sort} projection is declared as an implicit coercion~\cite{Saibi97}
and is automatically inserted in order to make \coq{M -> M -> M} a meaningful
type for binary operations on the carrier of \coq{M}. Last, we see that
properties are quantified on (hence apply to) the structure they belong to but
use, in their statements, operations belonging to simpler structures.
For example \coq{addNr} is a property of a ring but  its statement mentions \coq{add},\
the operation of the underlying monoid.

We then follow the proof of Hankel~\cite{nearrings} to show that
ring axioms imply the commutativity of the underlying monoid
(line~\ref{demo:theory:state:addrC}). This simple example shows we can
populate the theory of abstract rings with new results.

We use the \coq{Monoid_of_Type.Axioms} \emph{\phantterm{}}
(line~\ref{demo:theory:z:monoid:axioms})
in order to build an instance of the \coq{Monoid}
structure for binary integers \coq{Z}.
We then register that monoid instance as the canonical one on \coq{Z}
(line~\ref{demo:theory:z:monoid:canonical}) via the command
\coq{hb.canonical}.
We can similarly declare that \coq{Z} forms a ring by using
the \coq{Ring_of_Monoid.Axioms} \phantterm{}
(lines~\ref{demo:theory:z:ring:axioms} and~\ref{demo:theory:z:ring:canonical}).
Note that the \coq{Ring_of_Monoid.Axioms} \phantterm{} is not
a plain record constructor for \coq{Ring_of_Monoid.axioms}, since that
would require more arguments, namely the monoid ones (see the \coq{_}
at line~\ref{demo:theory:z:ring:axioms}). The
\phantterm{} synthesized by \HB{} infers them automatically
(as in~\cite[Section 7]{DBLP:conf/itp/MahboubiT13}) thanks to the \coq{hb.canonical}
declaration given just above.

From now on the axioms as well as the abstract theory of rings apply to
integers, as shown in lemma \coq{exercise}. The details of the proof
do not matter here, what is worth pointing out is that in a single
statement we mix monoid (e.g. \coq{+}) and ring (e.g. \coq{-}) operations and in the
proof we use monoid axioms (e.g. \coq{addrA}), ring axioms
(e.g. \coq{addrN}) and ring lemmas (e.g. \coq{addrC}), all seamlessly.

\subsection{Evolution of the hierarchy}\label{subsec:evolution}

We proceed by accommodating the intermediate
structure of Abelian groups.

\begin{coqcode}
#\hbmixin{}# Record Monoid_of_Type A := { ... (* unchanged *) ... }.
#\hbstructure{}# Definition Monoid := { A & Monoid_of_Type.axioms A }.

#\hbmixin{}# Record AbelianGroup_of_Monoid A of Monoid.axioms A := {
  opp : A -> A;
  addrC : commutative (add : A -> A -> A);
  addNr : left_inverse zero opp add;
}.
#\hbstructure{}# Definition AbelianGroup := { A of Monoid.axioms A & AbelianGroup_of_Monoid.axioms A }.

#\hbmixin{}# Record Ring_of_AbelianGroup A of AbelianGroup.axioms A := {
  one : A;
  mul : A -> A -> A;
  mulrA : associative mul;
  mul1r : left_id one mul;              mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;   mulrDr : right_distributive mul add;
}.
#\hbstructure{}# Definition Ring := { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.

Lemma addrN {R : AbelianGroup.type} : right_inverse zero opp add.               #\label{demo2:proof:addrN}#
Proof. by move=>x; rewrite addrC addNr. Qed.
\end{coqcode}

Some operations and properties were moved from the old \mixin{} for rings
into a newborn \mixin{} \coq{AbelianGroup_of_Monoid} that gathers the
axioms needed to turn a monoid into an Abelian group. Consequently the \mixin{}
for rings is now called \coq{Ring_of_AbelianGroup} (instead of \coq{Ring_of_Monoid})
since it expects the type \coq{A} to be already an Abelian group and hence
gathers fewer axioms. \\
While operations moved from one structure to another, some properties
undergo a deep change in their status.
The lemma \coq{addrC} part of the abstract theory of
rings is now an axiom of Abelian groups, while \coq{addrN} is no more
an axiom of rings, but rather a theorem of the abstract theory of Abelian
groups.

With this new version of the hierarchy, that we name V2, code written
for version V1 breaks. For example the declaration of the canonical ring over
the integers fails, if only
because we do not have a \coq{Ring_of_Monoid.Axioms} \phantterm{} anymore.

Our objective is to obtain a version of the hierarchy, that we name V3, that does
not only feature Abelian groups but that is also backward compatible
with V1.

\subsection{The missing puzzle piece}\label{subsec:v3}

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{puzzle.pdf}
  \end{center}
  \caption{\label{fig:puzzle}The evolution of the hierarchy. V3 is backward compatible with V1, while V2 is not.}
\end{figure}

The key to make a hierarchy evolve without breaking user code is the full
fledged notion of \emph{\factory{}} (the \mixins{} seen so far are degenerate,
trivial, \factories{}).
\Factories{}, like \mixins{}, are packages for operations and properties but are
not directly used in the definition of structures. Instead a \factory{} is
equipped with \emph{\builder{}s}: user provided pieces of code that extract
from the \factory{} the contents of \mixins{}, so that existing
\phantterm{}s can be used.

As depicted in figure~\ref{fig:puzzle} we change again the hierarchy
by declaring a \coq{Ring_of_Monoid} \factory{}, that, from the user point of view,
will look indistinguishable from the old \coq{Ring_of_Monoid} \mixin{}
and hence grant backward compatibility between version V3 and version V1.

\begin{coqcode}
#\hbfactory{}# Record Ring_of_Monoid A of Monoid.axioms A := { ... (* unchanged, see page 2 line 13 *) ... }.
#\hbbuilders{}# Context A (f : Ring_of_Monoid.axioms A).                          #\label{demo3:variable:f}#

  Lemma addrC : commutative add. Proof. (* The same proof as before *) Qed.

  Definition to_AbelianGroup_of_Monoid :=
    AbelianGroup_of_Monoid.Axioms A (opp f) addrC (addNr f). (* addrN unused *)
  #\hbinstance{}# A to_AbelianGroup_of_Monoid.

  Definition to_Ring_of_AbelianGroup :=
    Ring_of_AbelianGroup.Axioms A (one f) (mul f)
      (mulrA f) (mul1r f) (mulr1 f) (mulrDl f) (mulrDr f).
  #\hbinstance{}# A to_Ring_of_AbelianGroup.

#\hbend{}#.
\end{coqcode}

The record \coq{Ring_of_Monoid.axioms} is the same we declared as a \mixin{}
in version V1. In order to make a \factory{} out of it we equip it with
two definitions that embody the \emph{\builder{}s}.
The first is \coq{to_AbelianGroup_of_Monoid} and
explains how to build an \coq{AbelianGroup} structure out of the \factory{}
axioms (named \coq{f}, line~\ref{demo3:variable:f}).
This construction is also registered as canonical for \coq{A},
so that the next construction \coq{to_Ring_of_AbelianGroup} can call
the \coq{Ring_of_AbelianGroup.Axioms} \phantterm{} that requires \coq{A} to be
an \coq{AbelianGroup}.
It is worth pointing out that the proof of \coq{addrC} we had in V1 is now
required in order to write the \builder{} for Abelian groups, while
the \coq{addrN} field is not used (the same statement is already part of
the theory of Abelian groups, see line~\ref{demo2:proof:addrN} of the previous
code snippet).

Thanks to this \factory{} we can now declare \coq{Z} to be an instance
of a ring using the \coq{Ring_of_Monoid.Axioms} \phantterm{}.
The associated \builder{}s generate, behind the scenes, instances of the
\coq{Ring_of_AbelianGroup.axioms} and \coq{AbelianGroup_of_Monoid.axioms} \mixins{}
that in turn are used to build instances of the \coq{AbelianGroup} and \coq{Ring}
structures. Indeed, when used in the context of the hierarchy version V3,
the command \hbinstance{} \coq{Z Z_Ring_axioms}
makes \coq{Z} an instance of \emph{both} structures, and not just the \coq{Ring}
one as in version V1.
Thanks to that the proof of \coq{example} can use the theory of
both structures, for example \coq{addrC} holds on Abelian groups,
\coq{addrA} holds on monoids, while \coq{addr1} holds on rings.
As a result the very same proof works on both
version V1 and V3.

Last, it is worth pointing out that the new \factory{} makes the following
two lines equivalent (the former declares
rings in V1) since they both describe the same set of \mixins{}.

\begin{coqcode}
#\hbstructure{}# Definition Ring := { A of Monoid.axioms A & Ring_of_Monoid.axioms A }.
#\hbstructure{}# Definition Ring := { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.
\end{coqcode}

This is another example of client code that would not break: the client of the hierarchy is allowed to declare new
structures on top of existing ones.

\subsection{\HB{} in a nutshell}

By using \HB{} the hierarchy designer has the following freedoms and advantages.

\begin{itemize}
\item Operations and properties (axioms) are made available to the user
      as soon as they are used in a structure. The hierarchy designer
      is free to move them from one to another and replace an axiom
      by a lemma and vice versa.
\item Structures cannot disappear but the way they are built may change.
      The hierarchy designer is free to split structures into smaller
      ones in order to better factor and reuse parts of the hierarchy
      and the library that follows it.
\item \Mixins{} cannot disappear but can change considerably
      in their implementation. A \mixin{} can become a full fledged
      \factory{} equipped with \builder{}s to ensure backward compatibility.
\item \HB{} high level commands compile to
      the discipline of Packed Classes (Coq modules, records, coercions,
      implicit arguments, canonical structure instances, notations,
      \phantterms{}).
      This process lifts a considerable
      burden from the shoulders of the hierarchy designer and the final user
      who are no longer required to master the details of Packed Classes.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\HB{} language}\label{sec:language}

The Coq terms handled by \HB{} are subdivided in five categories, the
\mixins{} \(\M{}\), the \factories{}~\(\F{}\), \mixinbuilders{}
\(\B{}\), the \classes{} \(\C{}\) and the \structures{}
\(\Str{}\). \Mixins{}, \factories{} and instances are tagged by the
user, through the commands \hbmixin{}, \hbfactory{} and \hbinstance{},
and the user may rely on their implementation. However structures and
classes are generated with the command \hbstructure{} and builders are
generated when using \hbinstance{} while declaring a \factory{}, and
the user may only refer to structure types, but should never rely on
their implementation, neither should they rely on explicit builders.

\subsection{\HB{} \mixins{}, \factories{} and instances}
\label{sec:hb-mixins-factories}

\begin{definition}[\M{}, \mixins{}]
A \emph{\mixin{}} \(m \in \M{}\) is a Coq record with one or more parameters.
The first parameter must be a \coq{(T : Type)}, while the other parameters
are \mixins{} applied to \coq{T} and possibly $n$ other distinct \mixins{}. I.e.
\begin{coqcode}
Record #$m$# (T : Type) #\(\pmp{n}\)# : Type := { .. }.
\end{coqcode}
where \(\pmp{n} = \vect{\left(p_i : m_i\ T\ p_{\sigma(i,1)}\ \ldots\ p_{\sigma(i,n_i)}\right)}_{i \in \enum1n}\).
For all \(i \in \enum1n\) we have \(m_i \in \M{}\) and its
arguments consist in $n_i$ of the previously quantified
mixin parameters $p_k$, i.e. for all \(k \in \enum1{n_i}\), we
have \(\sigma(i,k) \in \enum1{i-1}\).
\end{definition}

\begin{definition}[\dep{}, \mixin{} dependencies]
  Given \(m \in \M{}\), we define \(\dep{}(m) \in \powerset\M{}\)
  as the set of all \mixins{} that occur as parameters of \(m\), i.e.
  \(\dep{}(m) = \enum{m_1}{m_n}\).
\end{definition}

\begin{remark}For all \(i \in \enum1n\), we have
\(\dep{}(m_i) = \enum{m_{\sigma(i,1)}}{m_{\sigma(i,m_i)}}\).
\end{remark}

\begin{definition}If \(f : \mathcal{X} \to \powerset{\mathcal{Y}}\), then
\(f^\star : \powerset{\mathcal{X}} \to \powerset{\mathcal{Y}}\) is defined as
\(f^\star(X) = \bigcup_{x \in X} f(x).\)
\end{definition}

\begin{proposition}
\dep{} is transitively closed since records are well typed in the
empty context and describes a DAG since Coq does not admit circular
definitions: \(\dep{}^\star(\dep{}^\star(M)) \subseteq \dep{}^\star(M)\)
\end{proposition}

\begin{definition}[\factories{} \F{}, \mixinbuilders{} \B{} and \from{}]
  A \emph{\mixinbuilder{}} \(\mu \in \B{}\) is any function which parameters are the carrier
  type \coq{(T: Type)}, an arbitrary number of distinct \mixins{}
  \(\enum{m_1}{m_n}\) and an additional argument \(f\), which we call
  a \emph{\factory{}}, and which depends on the \mixins{}. Its return type is
  a \mixin{} \(m_{n+1} \in \M{}\) i.e.
\begin{coqcode}
Definition f (T : Type) #$\pmp{n}$# : Type := ...
Definition #$\mu$# (T : Type) #$\pmp{n}$#: #\(f\ T\ p_1 \ldots p_n\)# -> #\(m_{n+1}\ T\ p_{\sigma(n+1,1)}\ \ldots\ p_{\sigma(n+1,n_i)}\)# := ...
\end{coqcode}

  We define \(\from{} (f, m_{n+1}) = \mu\) to be the (unique) \mixinbuilder{}
  for \(m_{n+1}\) from the \factory{} \(f\).
\end{definition}

\begin{definition}[$\provides{}$, $\requires{}$]
  We define the following functions \begin{itemize}
  \item\( \requires{}(f) = \enum{m_1}{m_n}  \in \powerset\M{}\)
    is the set of mixins a \factory{} \(f \in \F{}\) depends on.
    As a consequence all the
    \mixinbuilders{} of a given \factory{} have the same set of
    dependencies.
  \item\( \provides{}(f) = \{m | \from{}(f,m) \textrm{ is defined}\}
    \in \powerset\M{}\) the set of mixins that a factory
    \(f \in \F{}\) provides, through its \mixinbuilders{}.
  \end{itemize}
Moreover the following properties hold for all \(f \in \F{}\):
  %\begin{proposition}
  %  They validate the following properties for all \(f \in \F{}\):
    \begin{enumerate}
      \item\( \dep{}^\star(\requires{}(f)) \subseteq \requires{}(f)\)
      \item\( \requires{}(f) \cap \provides{}(f) = \emptyset \)
      \item\( \forall m \in \provides{}(f),~ \from{}(f,m) \) is defined
    \end{enumerate}
  %\end{proposition}
\end{definition}

\Mixins{} are declared by the user as the fundamental building blocks of
a hierarchy. As the next proposition shows they shall not be distinguished
from regular \factories{}, since they are a special cases.

\begin{proposition}[\( \M{} \subseteq \F{} \)] For all \(m \in \M{}\)
  we have \(\requires{}(m) = \dep{}(m)\), \(\provides{}(m) = \{m\}\)
  and
  \(\from{}(m,m) = \left(\coq{fun T }\ \pmp{n} \ \coq{(x :}\ m\
  \coq{T }\ p_1\ \ldots p_n\ \coq{) => x}\right) \in \B{}.\)
\end{proposition}

\begin{command} to declare a new mixin:
\begin{coqcode}
#\hbmixin{}# Record M T #$f_1\ \ldots\ f_n$# := { .. }. #\hbend{}#.
\end{coqcode}
  Creates a module \coq{M} with a record
  \coq{M.axioms}, which depends on the \mixins{}
  \(\requires{}(\coq{M.axioms}) =
  \provides{}^\star\left(\enum{f_1}{f_n}\right)\), and registers this
  new record both as a \mixin{} and a \factory{}. It also creates a
  \phantterm{} \coq{M.Axioms.}
\end{command}

\begin{command} to declare an instance: \hbinstance{}
  \(X\ b_1 \ldots b_k\), synthesizes terms corresponding to all the
  \mixins{} that can be built from the $b_i$. Indeed if
  \(b_i : f_i\ T\ \ldots\), then this command creates elements of type
  \(\provides{}^\star\left(\enum{f_1}{f_k}\right).\) This command also
  generates unification hints as described in
  Section~\ref{sec:target-lang}.
\end{command}

\begin{command} to declare a new \factory{} and generate new \mixinbuilders:
\begin{coqcode}
#\hbfactory{}# Record M T #$f_1\ \ldots\ f_n$# := { .. }.
#\hbbuilders{}# Context T (a : M.axioms A).

  Definition #$b_{n+1}$# : #$f_{n+1}$# T .. := ...
  #\hbinstance{}# T #$b_{n+1}$#.
  ..
  Definition #$b_{n+k}$# : #$f_{n+k}$# T .. := ...
  #\hbinstance{}# T #$b_{n+k}$#.
#\hbend{}#.
\end{coqcode}
  Creates a module \coq{M} with a record
  \coq{M.axioms}, which depends on the mixins
  \(\requires{}(\coq{M.axioms}) =
  \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and registers this
  new record as a factory. We then use the \factoryinstances{}
  \(b_{n+1}, \ldots, b_{n+k}\) provided by the user and derive
  \mixinbuilders{} from them, so that
  \(\provides{}(\coq{M.axioms}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)

  It is thus necessary that
  \(\requires{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right) \subseteq
  \provides{}^\star\left(\enum{f_1}{f_n}\right).\)
\end{command}

Note that the \(b_i\) are not \mixinbuilders{} since their return types
are not necessarily \mixins{}, but could be a \factory{}. However, since all
\factories{} provide \mixin{} through \mixinbuilders{}, we can
obtain \mixinbuilders{} out of each \(b_i\) by function composition.

\subsection{\HB{} classes and structures}
\label{sec:hb-class-struct}

\begin{definition}[\C{}, class]
A \class{} \coq{c} \(\in \C{}\) is a Coq record with one parameter
\coq{(T : Type)}. The type of each field is a \mixin{} in \M{} applied to
\coq{T} and, if needed, any number of other fields:
\begin{coqcode}
Record #$c$# (T : Type) := { #\(p_1 : m_1\ T\)#; .. ; #\(p_i : m_i\ T\ p_{\sigma(i,1)}\ \ldots p_{\sigma(i,n_i)}\)#; ..}.
\end{coqcode}
\end{definition}

\begin{definition}[\cdef{}, class definition]
  We call \(\cdef{}(c) \in \powerset\M{}\) the set of \mixins{}
  mentioned in the fields of the \class{}, i.e.
  \(\{m_1, \ldots,m_n\}\).  Given that \class{} records are well typed
  in the empty context the set of \mixin{} records is closed
  transitively.  The implementation enforces that no two \class{}
  records contains the same set of \mixins{} (disregarding the order of
  the fields), i.e. \cdef{} is injective.
\end{definition}

\begin{invariant}[\cdef{}]
  For all \(f \in \F{}\):
  \begin{enumerate}
    \item\( \exists c \in \C{},~ \cdef{}(c) = \requires{}^\star(f) \cup \provides{}^\star(f),\)
    \item\( \exists C \subseteq \C{},~\cdef{}^\star(C) = \requires{}^\star(f).\)
  \end{enumerate}
\end{invariant}

Classes could also be seen as factories.
\begin{proposition}[\( \C{} \subseteq \F{} \)]
  forall \(m \in \cdef{}(c)\) we have \(\from{}(c,m_i) = p_i,\) and it
  follows that \(\requires{}(c)=\emptyset \),
  \(\provides{}(c)=\cdef{}(c).\)
\end{proposition}
However since classes are generated, their implementation may change,
thus users should not rely on constructors of classes. Hence the only
way users may refer to a class is as an argument of \hbmixin{},
\hbfactory{} or \hbstructure{}.

\begin{definition}[\Str{} Structure]
A \structure{} \coq{s} \(\in \Str{}\) is a dependent pair: a Coq record where
the value of the first field occurs in the type of the second.
The first field is \coq{(sort : Type)} and the second field is
\coq{(class : c sort)} for some \coq{c} \(\in \C{}\).
As a consequence \structures{} are in bijection with \classes{}.
\end{definition}

\begin{command} to declare a class and structure:
  \hbstructure{} \coq#Definition M := { A of# \(f_1 \ldots f_n\) \coq#}.#
  crafts a class \coq{M.class_of} $\in \C{}$ where
  $\cdef{}(c) = \provides{}^\star\left(\enum{f_1}{f_n}\right) $ and the
  corresponding structure \coq{M.type} $\in \Str{}$, together with
  unification hints as described in Section~\ref{sec:target-lang}.
\end{command}

\begin{definition}[\issubclass{} \(\in \C{}\times\C{}\), \subclass{}]\label{def:subclass}
  \(c1~\issubclass{}~c2\) iff \(\cdef{}(c2) \subseteq \cdef{}(c1)\)
  \end{definition}

\subsection{Automatic inference of \mixins{}}
\label{sec:autom-infer-mixins}

Since \mixins{} may change but \factories{} stay the same, \HB{}
commands must never rely on a particular set of \mixins{} as
arguments, and \factory{} arguments must never be given explicitly by
the user. As described in Sections~\ref{sec:hb-mixins-factories}
and~\ref{sec:hb-class-struct}, \HB{} commands take a list of
\factories{} as arguments, which they expand into lists of \mixins{}
behind the scene. However \factory{} types and constructors have
\mixin{} arguments that must be inferred automatically when used. To
this end, the commands \hbmixin{} and \hbfactory{} generate
\phantterms{} for the user to replace uses of constructors of \factories{}.
These commands first create a record \(f_{aux}\) with a
constructor \(F_{aux}\) and then create abbreviations \(f\) and \(F\)
that automatically fill the mixin arguments of \(f_{aux}\) and \(F_{aux}\)
respectively.
See \cite[Section 7]{DBLP:conf/itp/MahboubiT13} for a detailed description
of how to implement these abbreviations in Coq.

\begin{coqcode}
Record #\(f_{aux}\)# T #\(\pmp{n}\)# := #\(F_{aux}\)# { ... }.
Notation #$f$# T := (#\(f_{aux}\)# T ...(* $p_1 \ldots p_n$ inferred when $T$ is known *)).
Notation #$F$# T  #\(x_1\ \ldots\ x_k\)# := (#\(F_{aux}\)# T ...(* $p_1 \ldots p_n$ inferred when $T$ is known *) #\(x_1\ \ldots\ x_k\)#).
Definition b : #$f$# T := #$F$# T #\(x_1\ \ldots\ x_n\)#
\end{coqcode}

% As described in \cite{affeldt:hal-02463336},... % the only way to define
% % inheritance of structure in a deterministic fashion is to replace

% Cite \cite{KSdraft}

% \begin{definition}[\(\join{} \in \C{}\times\C{} \to \C{}\)]\label{def:join}
%   if \(c1, c2 \in \C{}\) then \(\join{}(c1,c2)\) is the smallest,
%   w.r.t. set inclusion, \class{} \(c\) such that
%   \(\cdef{}(c1) \cup \cdef{}(c2) \subseteq \cdef{}(c)\) if it exists.
% \end{definition}

% \begin{remark}[]
% \(\join{}(c1,c2)\) is a \subclass{} of both \(c1\) and \(c2\).
% \end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The target language: Coq with Packed, flat, Classes}
\label{sec:target-lang}

The language of packed classes were introduced in
\cite{DBLP:conf/tphol/GarillotGMR09} and used directly to describe the
algebraic hierarchy of the Mathematical Components library.
It is based on a disciplined use of records and projections and
on the possibility of extending the elaborator of Coq
via the declaration of Canonical Structures instances.
In this section we describe the \emph{flat} variant of
Packed Classes, the target language of \HB{}.

\subsection{Describing structures with records and projections}

We describe mathematical structures with three kinds of dependent records: mixins, classes, and structures.
As shown in section \ref{sec:example}, a \mixin{} gathers operators and axioms newly introduced by a structure.
As in \cite[section 2.4]{DBLP:conf/tphol/GarillotGMR09}\cite[section 2]{KSdraft}, a class record is parametrized by the carrier type \coq{(T : Type)} and gathers all the operators and axioms of a structure by assembling mixins, and a \coq{Structure type} (a record) bundles a carrier and its class instance, as follows.
\begin{coqcode}
Module Monoid.
Record axioms (T : Type) : Type :=
  Class { Monoid_of_Type_mixin : Monoid_of_Type.axioms T; }.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End Monoid.
\end{coqcode}
The \coq{Monoid} module plays the role of a name space and forces us to write
qualified names such as \coq{Monoid.type}; as a consequence we can reuse the
same unqualified names for other structures, i.e., class and structure record
can always be named as \coq{axioms} and \coq{type} respectively.

Mixins and classes are internal definitions to structures; in contrast,
\coq{Monoid.type} is part of the interface of the monoid structure.
As seen in section \ref{sec:example}, we declare \coq{Monoid.sort} as an
implicit coercion and lift monoid operators and axioms from projections for
the monoid \mixin{} to definitions and lemmas for \coq{Monoid.type} as follows.
\begin{coqcode}
Coercion Monoid.sort : Monoid.type >-> Sortclass.

Definition zero {T : Monoid.type} : T :=
  Monoid_of_Type.zero T (Monoid.Monoid_of_Type_mixin T (Monoid.class T)).
Definition add {T : Monoid.type} : T -> T -> T :=
  Monoid_of_Type.add T (Monoid.Monoid_of_Type_mixin T (Monoid.class T)).
Lemma addrA {T : Monoid.type} : associative (@add T).
(* Two monoid axioms `add0r` and `addr0` are omitted. *)
\end{coqcode}

Next we define Abelian groups.
Since the monoid structure is the bottom of this hierarchy its class record
\coq{Monoid.axioms} consists just one \mixin{}. In contrast the class record
of Abelian groups consists of two \mixins{} where the second one depends
 on the former (since Abelian groups inherit from monoids).
\begin{coqcode}
Module AbelianGroup.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type.axioms T;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid.axioms T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort }.
End AbelianGroup.
\end{coqcode}
As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
coercion and lift additive inverse \coq{opp} and Abelian group axioms.
Since Abelian groups inherit from monoids, we also declare an implicit coercion
from Abelian groups to monoids. A coercion between structures can be defined
in two steps: first a coercion between the class record of the
superclass to the class record of the subclass (line~\ref{demo4:coe:ag:monoid:class});
and a coercion between
structure records (line~\ref{demo4:coe:ag:monoid:type}) relying on the first one.
\begin{coqcode}
Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup.axioms T) : #\label{demo4:coe:ag:monoid:class}#
  Monoid.axioms T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
  Monoid.Pack (AbelianGroup.sort T) (AbelianGroup.class T).

Definition opp {T : AbelianGroup.type} : T -> T := ... .
(* Two Abelian group axioms `addrC` and `addNr` are omitted. *)
\end{coqcode}

Generally, a coercion from a structure A to another structure B can (and should) have the following form, thanks to the corresponding coercion between classes \coq{A_class_to_B_class}.
\begin{coqcode}
Coercion A_to_B (T : A.type) : B.type :=
  B.Pack (A.sort T) ((* A_class_to_B_class _ *) (A.class T)).
\end{coqcode}

\subsection{Multiple inheritance}
In order to introduce multiple inheritance, we extend the hierarchy described in
section~\ref{subsec:v3} with the structure of semirings%, which is another
%intermediate structure between monoids and rings 
as depicted in figure \ref{fig:puzzle_v4}. Semirings introduce the binary
multiplication operator \coq{mul} and its identity element \coq{one}.

\begin{wrapfigure}[3]{r}{.4\textwidth}
 \includegraphics[width=.4\textwidth]{v4.pdf}
 \caption{V4 introduces multiple inheritance. For brevity we omit the factories/builders for the upper arrows.}
 \label{fig:puzzle_v4}
\end{wrapfigure}
\
\begin{coqcode}
#\hbmixin{}# Record SemiRing_of_Monoid A of Monoid.axioms A := {
  one : A;
  mul : A -> A -> A;
  (* 7 axioms are omitted. *)
}.

#\hbfactory{}# Record Ring_of_AbelianGroup A of AbelianGroup.axioms A := {
  (* 2 operators and 5 axioms are omitted. *)
}.
#\hbbuilders{}# Context A (a : Ring_of_AbelianGroup A).
..
#\hbend{}#.
\end{coqcode}

Since semirings and Abelian groups do not inherit from each other,
the definition of semirings and Abelian groups are quite similar:
\begin{coqcode}
Module SemiRing.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type.axioms T;
  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid.axioms T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End SemiRing.
\end{coqcode}
We define implicit coercions from the semiring structure to the carrier
and the monoid structure, and we lift semiring operators and axioms as follows:
\begin{coqcode}
Coercion SemiRing.sort : SemiRing.type >-> Sortclass.
Coercion SemiRing_to_Monoid : SemiRing.type >-> Monoid.type.

Definition one {T : SemiRing.type} : T := ... .
Definition mul {T : SemiRing.type} : T -> T -> T := ... .
(* 7 semiring axioms are omitted. *)
\end{coqcode}

The class record is defined by gathering the monoid, Abelian group, and
semiring \mixins{}. Since the rings inherit from monoids, semirings, and
Abelian groups, we define implicit coercions from the ring structure to those
three structures.
\begin{coqcode}
Module Ring.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type.axioms T;
  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid.axioms T Monoid_of_Type_mixin;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid.axioms T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End Ring.

Coercion Ring.sort : Ring.type >-> Sortclass.
Coercion Ring_to_Monoid : Ring.type >-> Monoid.type.
Coercion Ring_to_SemiRing : Ring.type >-> SemiRing.type.
Coercion Ring_to_AbelianGroup : Ring.type >-> AbelianGroup.type.
\end{coqcode}

\subsection{Linking structures and instances via Canonical Structures}

We recall here how to use the Canonical
Structures~\cite{DBLP:conf/itp/MahboubiT13,Saibi99phd} mechanism, which
lets the user extend the elaborator of Coq, in order to handle
inheritance and inference of structure~\cite{affeldt:hal-02463336,KSdraft,DBLP:conf/tphol/GarillotGMR09}.
This software component
takes in input a term as written by the user and has to infer all the missing information that is necessary in order to make the term well typed.

A first example of elaboration that requires canonical structures is \(0 + 1\).
After removing all syntactic facilities, the underlying Coq term is
\coq{(add _ (zero _)) (one _)} where \coq{_} stands for an implicit piece
of information to be inferred and the constants \coq{add} and \coq{zero} come
from the monoid structure while \coq{one} comes from semirings.
When the term is type checked each \coq{_} is replaced by a unification
variable written $?_v$.
Respectively, the head and the argument of the top application
can be typed as follows:
\begin{coqcode}
  add #$?_{\texttt{M}}$# (zero #$?_{\texttt{M}}$#) : Monoid.sort #$?_{\texttt{M}}$# -> Monoid.sort #$?_{\texttt{M}}$#
  one #$?_{\texttt{SR}}$# : SemiRing.sort #$?_{\texttt{SR}}$#
\end{coqcode}
where \(?_{\texttt{M}}~\coq{: Monoid.type}\) and
\(?_{\texttt{SR}}~\coq{: SemiRing.type}\).
In order to type check the application Coq has to unify \(\coq{Monoid.sort}~?_{\texttt{M}}\) with
\(\coq{SemiRing.sort}~?_{\texttt{SR}}\), which is not trivial: it amounts at
finding a structure that is both a monoid and a semiring, possibly the
smallest one~\cite[Sect.~4]{KSdraft}.
This piece of information can be inferred from the hierarchy and its inheritance
relation (Definition~\ref{def:subclass}) and we can tell Coq to exploit it by
declaring \coq{SemiRing_to_Monoid : SemiRing.type -> Monoid.type} as
canonical. With that hint Coq will pick \(?_{\texttt{M}}\) to be
\(\coq{SemiRing_to_Monoid}~?_{\texttt{SR}}\) as the canonical solution for this
unification problem.
In general, all the coercions between structures must be declared
as canonical. % to make this inference work.

Another example of elaboration problem is \(- 1\), which hides the term~\coq{opp _ (one _)}.
Here \coq{opp} and \coq{one} are respectively from Abelian groups and semirings,
which do not inherit each other but whose smallest common substructure is
rings; thus we have to extend the unifier to solve a unification problem
\(\coq{AbelianGroup.sort}~?_{\texttt{AbG}}~\coq{= SemiRing.sort}~?_{\texttt{SR}}\) by instantiating \(?_{\texttt{AbG}}\) and \(?_{\texttt{SR}}\) with \(\coq{Ring_to_AbelianGroup}~?_{\texttt{R}}\) and \(\coq{Ring_to_SemiRing}~?_{\texttt{R}}\) respectively where \(?_{\texttt{R}}\) is a fresh unification variable of type \coq{Ring.type}.
This hint can be given as follows:% can be done by the following canonical declaration.
\begin{coqcode}
Canonical AbelianGroup_to_SemiRing (T : Ring.type) :=
  SemiRing.Pack (AbelianGroup.sort (Ring_to_AbelianGroup T)) (Ring.class T)).
\end{coqcode}

Similarly, one can apply an algebraic theory to an instance (an example) of that structure, e.g., \(2 \times 3\) where \(2\) and \(3\) have type \(\mathbb{Z}\).
The same mechanism of canonical structures let us extend the unifier to solve \(\coq{SemiRing.type}~? = \mathbb{Z}\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The implementation in Coq-Elpi}\label{sec:implementation}

The implementation is based on the Elpi
extension language for Coq. In this section we introduce the features of the
programming language that came in handy in the development of \HB{} and
comment a few code snippets.

Coq-Elpi~\cite{CoqElpi} is a Coq plugin embedding
Elpi and providing an
extensive, high level, API to access and script the Coq system at the
vernacular level.
This API lives in the \elpi{coq} namespace and lets one easily declare
records, coercions, canonical structures, modules, implicit arguments, etc.
The most basic Coq data type exposed to Elpi is the one of references to global
declarations:

\begin{elpicode}
kind gref  type.                 % The data type of references to global terms
type indt  inductive -> gref.    % eg: Coq.Init.Datatypes.nat
type indc  constructor -> gref.  % eg: Coq.Init.Datatypes.O
type const constant -> gref.     % eg: Coq.Init.Peano.plus
\end{elpicode}

The arguments of the three constructors are opaque to Elpi, that can only use
values of these types via dedicated APIs. For example the API for declaring
an inductive type will generate a value of type \elpi{inductive} that
is printed as, for example, \elpi{«nat»}.

Elpi~\cite{DBLP:conf/lpar/DunchevGCT15} is a dialect
of $\lambda$Prolog~\cite{Miller:2012:PHL:2331097}, an higher order
logic programming language that makes it easy to manipulate abstract syntax
tree with binders. Coq-Elpi takes full advantage of this capability by
representing Coq terms in Higher Order Abstract
Syntax~\cite{10.1145/53990.54010} style, reusing the binder of the programming
language in order to represent Coq's ones. Here is an excerpt of the data
type of Coq terms:

\begin{elpicode}
kind term type.                              % The data type of Coq terms
type global gref -> term.                    % eg: nat, O, S, plus, ...
type fun    term -> (term -> term) -> term.  % eg: fun x : t => b(x)
type app    list term -> term.               % eg: app [hd|args]
... % all other term constructors are omitted for brevity
\end{elpicode}

Note that the \elpi{fun} constructor holds a $\lambda$Prolog function.
In this syntax the Coq term \coq{(fun x : nat => x x)} becomes
\elpi{(fun (global (indt «nat»)) x\ app[x, x])} where \elpi{x\} binds \elpi{x}
in the body of the function. Substitution of a bound variable for a term
can be computed by applying a term (of function type) to an argument.

Data types with binders are also used as input to high level APIs that build
terms behind the scenes. For example a Coq record is just an inductive type
and the API to declare one must allow the type of a field to depend on the
fields that comes before it. Note that the \elpi{field} constructor
takes a coercion flag, the name of the field, its type and binds a term
in the remaining record declaration.

\begin{elpicode}
kind indt-decl type.              % The type of an inductive type declaration
type record      string -> term -> string -> record-decl -> indt-decl.
type field       bool -> string -> term -> (term -> record-decl) -> record-decl.
type end-record  record-decl.
... % constructors for non-record inductive types are omitted for brevity

external pred coq.env.add-indt i:indt-decl, o:inductive.
external pred coq.CS.canonical-projections i:inductive, o:list (option constant).
\end{elpicode}

The \elpi{pred} keyword documents types and modes (input or output) of the
arguments of a predicate, while \elpi{external} signals that the
predicate is a builtin (in other words it is implemented in
OCaml rather than $\lambda$Prolog).

We comment these two builtin predicates and the \elpi{indt-decl} type
while looking at the code of \elpi{declare-structure} that is in charge
of scripting the following Coq code:% (given as a parameter \coq{axioms}):

\begin{coqcode}
Structure type : Type := Pack { sort : Type; class : axioms sort }.
\end{coqcode}

The following Elpi code builds the declaration, type checks it, adds it to the
Coq environment and finally returns the projections for the sort and the class
fields.% (it is handy to have them in the rest of the code).

\begin{elpicode}
pred declare-structure i:gref, o:term, o:term, o:term.

declare-structure ClassName Structure SortProjection ClassProjection :- std.do! [
  StructureDeclaration =
    record "type" {{ Type }} "Pack" (
      field tt "sort" {{ Type }} s\                        %\label{binder:sort}%
      field ff "class" (app [global ClassName, s]) _\
    end-record),
  coq.typecheck-indt-decl StructureDeclaration,
  coq.env.add-indt StructureDeclaration StructureName,     %\label{record:decl}%
  coq.CS.canonical-projections StructureName [some SortP, some ClassP],
  Structure = global (indt StructureName),
  SortProjection = global (const SortP),
  ClassProjection = global (const ClassP),
].
\end{elpicode}

Note that the binder \elpi{s\} at line \ref{binder:sort} lets one mention
the first field in the type of the second.
The syntax \elpi{{{ Type }}} is a quotation: it lets
one use the syntax of Coq to write an Elpi expression of type \elpi{term}.
The API \elpi{coq.CS.canonical-projections} lets us find the projections
automatically generated by Coq for a given record.
The last detail worth mentioning is that this program makes no use of
backtraking: the \elpi{std.do!} combinator signals that.

In the simple case of the structure record, the number of fields, and hence
the number of binders, is fixed. On the contrary the class record has one
field per \mixin{} and each of them can depend on the previous ones.
In order to synthesize terms with binders in an inductive fashion (using
a recursive predicate) $\lambda$Prolog lets one postulate fresh nominal
constants using the \elpi{pi} operator and attach to the
nominal some knowledge in the form of a clause via the \elpi{=>} operator.
This process is called binder mobility: the binder is moved from
the data (that we are building) to the program (the context of the
current computation).
This feature is key to the following code that synthesizes the
declaration of the fields of the class record.

\begin{elpicode}
pred synthesize-fields.field-for-mixin i:mixinname, o:term.
pred synthesize-fields i:list mixinname, i:term, o:record-decl.

synthesize-fields [] _ Decl :- Decl = end-record.
synthesize-fields [M|ML] T Decl :- std.do! [
  get-mixin-modname M ModName, Name is ModName ^ "_mixin",                          %\label{craft:name}%
  dep1 M Deps,                                         %\label{deps:fetch}%
  std.map Deps synthesize-fields.field-for-mixin Args, %\label{deps:satisfy}%
  Type = app[ global M, T | Args ],                    %\label{build:type}%
  Decl = (field ff Name Type f\ Fields f),
  pi m\                                                %\label{postulate:m}%
    synthesize-fields.field-for-mixin M m =>           %\label{postulate:satisfy}%
    synthesize-fields ML T (Fields m)                  %\label{rec:call}%
].
\end{elpicode}

The first predicate \elpi{synthesize-fields.field-for-mixin} is used
to link a \mixin{} to a nominal that corresponds to the record field
for that mixin. It has no clauses in the base program but some clauses
are added dynamically by \elpi{synthesize-fields}.

The second predicate proceeds by recursion on the (topologically sorted) list of \mixins{},
and terminates when the list is empty. If the list contains a \mixin{} \elpi{M}
then it crafts a \elpi{Name} for it (line~\ref{craft:name}),
fetches its dependencies (line~\ref{deps:fetch}) and
finds the (previously declared) record fields holding these \mixins{}
(line~\ref{deps:satisfy}).
The \elpi{(std.map L1 P L2)} predicate relates the two lists \elpi{L1} and
\elpi{L2} point wise using the predicate \elpi{P}.

Line~\ref{build:type} builds the type of the field: the \mixin{} name applied
to the type (sort) and all its dependencies.
Note that the \elpi{Fields} variable, representing the declaration of the
next fields, is under the binder for \elpi{f} (the current field).
In order to make the recursive call under that binder (line~\ref{rec:call})
and recursively process \elpi{ML} we postulate a nominal
\elpi{m} (line~\ref{postulate:m}) that is a term satisfying any
future dependency on the current \mixin{} (line~\ref{postulate:satisfy})
and we replace \elpi{f} by \elpi{m} in \elpi{Fields} by writing
\elpi{(Fields m)}.

% The code of \elpi{synthesize-fields} makes an idiomatic use
% of \elpi{=>} in $\lambda$Prolog: the current program is temporarily
% extended with a clause.
%
% \marginpar{this may go}The same operation, but with a persistent
% effect, is made available by Coq-Elpi via a dedicated API to extend
% data bases. A data base is a named set of clauses and
% commands are typically built by accumulating specific clauses on top
% of a data base. Here an excerpt of the \verb+hierarchy.db+ data base and
% of the \elpi{declare_mixin} command. Text between \verb+lp:{{+ and \verb+}}+
% is Elpi code embedded in a Coq file. All the vernacular commands beginning
% with \coq{Elpi} are provided by the Coq-Elpi plugin.
%
% \begin{coqcode}
% Elpi Db hierarchy.db lp:{{
%   pred dep1 o:mixinname, o:list mixinname.
%   ... % for brevity we omit class and join
% }}.
% Elpi Command declare_mixin.
% Elpi Accumulate Db hierarchy.db.
% Elpi Accumulate lp:{{
%   main [str S] :- !, std.do! [
%     coq.locate S M,
%     coq.env.typeof M Ty,
%     gather-mixin-dendencies Ty [] ML, % implementation omitted for brevity
%     coq.elpi.accumulate "hierarchy.db" (clause _ _ (dep1 M ML)),
%     ...
%   ].
%   main _ :- coq.error "Usage: declare_mixin <mixin>".
% }}.
% \end{coqcode}
%
% Line 11 computes the dependencies of \mixin{} \elpi{M} given
% its type and line 12 crafs a clause for \elpi{dep1} and
% adds it to the data base \elpi{"hierarchy.db"}.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Benchmark}
%if we can port ssralg, or not... It would be nice, but maybe its a bit late.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}
The most closely related work is the one about Packed Classes~\cite{DBLP:conf/tphol/GarillotGMR09} on which we
build. The main differences are that \HB{} is a higher level language
that is compiled down to (flat) Packed Classes. The systematic use of
\factories{} makes the user interface of a hierarchy stable under the insertion of
structures, a property that Packed Classes lacks.
Finally many of the intricacies of Packed Classes are hidden to the user by
the compilation step, in particular creating all the necessary coercions and canonical structure declarations,
especially in the case of diamonds or when merging several libraries, which used to cause the need for \textit{a posteriori}
validation of a hierarchy design~\cite{KSdraft}. It also opens the way
to automatically detect and solve problems tied to non judgmental commutative diagrams when forgetting
structure~\cite{affeldt:hal-02463336}.

In~\cite{CaretteCombinators} Carette and O'Connor describe the language of
Theory Presentation Combinators that can be used to describe a hierarchy of
algebraic structures.
They focus on the categorical semantics of the language that is built upon
the category of context.
They do not describe any actual compilation to the language of a mainstream
interactive prover, indeed they claim their language to be mostly type theory
independent. We know they considered targeting type theory and the language
of unification hints~\cite{10.1007/978-3-642-03359-9_8}
(a super set of the one of Canonical Structures),
but we could not find any written trace of that. Language wise they provides
keywords such as \verb+combine+ and \verb+over+ to share (reuse) a property
when defining a new structure. For example in order to avoid restating
the commutativity property when defining Abelian groups they combine
a commutative monoid and a group forcing the subjacent monoid to coincide: %
\verb-CommutativeGroup := combine CommutativeMonoid , Group over Monoid.-
In our language \HB{} the same role is played by \emph{\mixins{}}.
A \mixin{} lets one write once and for all a property and abstract it over types
and operations so that it can be reused in all the structures that need it.
One operation \HB{} allows for but that does not seem to be possible in the
setting of Presentation Combinators is the one of replacing an axioms with a
lemma and vice versa. As shown in \autoref{subsec:evolution} \HB{} supports
that.

The MMT system~\cite{RABE20131} provides a framework to describe formal
languages in a logical framework, providing good support for binders and
notations. It also provides an expressive module system to organize
theories and express relations among them in the form of functors.
At the time of writing
it provides limited support for elaborating user input taking systematic
advantage of the contents of the theories. The elaborator can be extended
by the user writing Scala code, and in principle use the contents of the
libraries to make sense of an incomplete expressions, but no higher level
language or mechanism is provided.

The library of the Mizar system features a hierarchy of algebraic
structures~\cite{7733265}. In spite of lacking dependent types, Mizar
provides the concept of attributed types and adjectives
that can be used to describe the signature of structures as one would
do with a dependently typed record and their properties as
one would define a conjunctive predicate.
The Mizar language also provides the notion of cluster that is used
to link structures: by showing that property $P$ implies property $Q$
one can inform automation that structures characterized by $P$ are
instances of structures characterized by $Q$. The foundational theory of Mizar
features an extensional notion of equality that makes it easy
to share the signature or the properties of structures by just requiring
a proof of their equivalence that is in turn used by automation to treat
equivalent structures as equal.

The concepts of \factory{} and \builder{} presented here is akin to the
\verb-AbstractFactory- and \verb-Builder- pattern from "the Gang Of Four" design
patterns~\cite{gamma1995design} in the sense that \factories{} are used
to build an arbitrary number of objects (here \mixin{} instances).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
In this paper we design and implement \HB{}, a high level language to describe
hierarchies of algebraic structures. The implementation of \HB{} is
based on the Elpi extension language for the Coq system and is available
at \url{https://github.com/math-comp/hierarchy-builder}.
The implementation amounts to approximately a thousand lines of (commented)
Elpi code and tree hundred lines of Coq vernacular. It took less than one month
to implement \HB{} while it took several years of attempts and fruitful
discussions with the other developers of the Mathematical Components library
to design it, as well as Coq users and developers and members of Dagstuhl seminar 18341.
In particular, discussions with Georges Gonthier, Assia Mahboubi and
Florent Hivert were instrumental in coining the concepts formalized here.

The \HB{} language is only loosely tied to Coq or even Type Theory.
We believe it could be adopted with no major change to other tools. Indeed the
properties and invariants that link \factories{}, \mixins{} and \classes{}
are key to rule out meaningless or ambiguous sentences and are not
specific to our logic setting. Moreover
most logics feature packing construction similar to records.

The compilation scheme we present in \autoref{sec:target-lang} is tied to dependent records,
that are available in Coq but also
in other provers based on Type Theory such a Matita~\cite{DBLP:conf/cade/AspertiRCT11}
and Lean~\cite{DBLP:conf/cade/MouraKADR15}.
We chose the flat variant of Packed Classes as the target for \HB{} because the
Mathematical Components library uses Packed Classes as well: As of today
Packed Classes offer the best compromise between flexibility and
performance~\cite[Section 8]{DBLP:conf/itp/MahboubiT13} in Coq.
Of course one could imagine a future were other approaches such as
telescopes~\cite{telescopes} or unbundled
classes~\cite{mathclasses} would offer equal or better performances in Coq, or in
another system.
It is a virtue of our work to provide a user language that is separate
from the implementation one. We believe it would take a minor coding effort to
retarget \HB{} to another bundling approach.

We leave to future work extending \HB{} to support structures parametrized
by structures such as the one of module over a ring.
We also leave to future
work the automatic synthesis of the notion of morphism between structures
of the hierarchy.

\bibliography{biblio}
\newpage
\appendix

\section{Coq reference}

\begin{coqcode}
Section OperationProperties.
Variable T : Type.
Variable e : T.
Variable inv : T -> T.
Variable op : T -> T -> T.
Variable add : T -> T -> T.

Definition left_id  := forall x, op e x = x.
Definition right_id := forall x, op x e = x.

Definition left_inverse := forall x, op (inv x) x = e.

Definition commutative := forall x y, op x y = op y x.
Definition associative := forall x y z, op x (op y z) = op (op x y) z.

Definition left_distributive  := forall x y z, op (add x y) z = add (op x z) (op y z).
Definition right_distributive := forall x y z, op x (add y z) = add (op x y) (op x z).
\end{coqcode}

\section{Proof of addrC}
\begin{coqcode}
Lemma addrC {R : Ring.type} : commutative (@add R).
Proof.
have innerC (a b : R) : (a + b) + (a + b) = (a + a) + (b + b).
  by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
have addKl (a b c : R) : a + b = a + c -> b = c.
  apply: can_inj (add a) (add (-a)) _ _ _.
  by move=> x; rewrite addrA addNr add0r.
have addKr (a b c : R) : b + a = c + a -> b = c.
  apply: can_inj (add ^~ a) (add ^~ (-a)) _ _ _.
  by move=> x; rewrite /= -addrA addrN addr0.
move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
by rewrite -!addrA [in RHS]addrA innerC !addrA.
Qed.
\end{coqcode}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "hb.tex"
%%% TeX-command-extra-options: "-shell-escape"
%%% ispell-local-dictionary: "english"
%%% End:

### 2020-03-10 hb.tex
minor typesetting
some fixes

@@
 \Mixins{} are declared via the \hbmixin{} command
-that takes a module name,
+that takes a record declaration with
 a type variable name and a possibly empty list of \factories{} for that type.
@@
 and~\ref{demo:mixin:mot:end} declares a \mixin{} that can turn a naked
-type \coq{A} into a monoid, hence we chose the module name to be \coq{Monoid_of_Type}.
-
-The \hbstructure{} command takes in input a name S and a possibly
-empty list of \factories{}.
+type \coq{A} into a monoid, hence we chose the name to be \coq{Monoid_of_Type}.
+
+The \hbstructure{} command takes in input the definition S for a type
+\coq{A} together with a list of \factories{} for \coq{A}.
 It registers a structure S in the hierarchy placing
@@
 }.
-#\hbstructure{}# Definition AbelianGroup := { A of Monoid.axioms A & AbelianGroup_of_Monoid.axioms A }.
+#\hbstructure{}# Definition AbelianGroup :=
+  { A of Monoid.axioms A & AbelianGroup_of_Monoid.axioms A }.
 
@@
 }.
-#\hbstructure{}# Definition Ring := { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.
+#\hbstructure{}# Definition Ring :=
+  { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.
 
@@
 \begin{coqcode}
-#\hbfactory{}# Record Ring_of_Monoid A of Monoid.axioms A := { ... (* unchanged, see page 2 line 13 *) ... }.
+#\hbfactory{}# Record Ring_of_Monoid A of Monoid.axioms A := {
+  ... (* unchanged, see page 2 line 13 *) ...
+}.
 #\hbbuilders{}# Context A (f : Ring_of_Monoid.axioms A).                          #\label{demo3:variable:f}#
@@
 \begin{coqcode}
-#\hbstructure{}# Definition Ring := { A of Monoid.axioms A & Ring_of_Monoid.axioms A }.
-#\hbstructure{}# Definition Ring := { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.
+#\hbstructure{}# Definition Ring :=
+  { A of Monoid.axioms A & Ring_of_Monoid.axioms A }.
+#\hbstructure{}# Definition Ring :=
+  { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.
 \end{coqcode}
@@
 \begin{coqcode}
-#\hbmixin{}# Record M T #$f_1\ \ldots\ f_n$# := { .. }. #\hbend{}#.
+#\hbmixin{}# Record M T of #$f_1\ \ldots\ f_n$# := { .. }.
 \end{coqcode}
@@
 \begin{coqcode}
-#\hbfactory{}# Record M T #$f_1\ \ldots\ f_n$# := { .. }.
+#\hbfactory{}# Record M T of #$f_1\ \ldots\ f_n$# := { .. }.
 #\hbbuilders{}# Context T (a : M.axioms A).
@@
 \begin{coqcode}
-#\hbmixin{}# Record SemiRing_of_Monoid A of Monoid.axioms A := {
+#\hbmixin{}# Record SemiRing_of_Monoid A 
+    of Monoid.axioms A := {
   one : A;
@@
 
-#\hbfactory{}# Record Ring_of_AbelianGroup A of AbelianGroup.axioms A := {
+#\hbfactory{}# Record Ring_of_AbelianGroup A
+    of AbelianGroup.axioms A := {
   (* 2 operators and 5 axioms are omitted. *)

### 2020-03-11 hb.tex
update syntax to v0.9.0

@@
 }.                                                             #\label{demo:mixin:mot:end}#
-#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type.axioms A }.        #\label{demo:structure:monoid}#
-
-#\hbmixin{}# Record Ring_of_Monoid A of Monoid.axioms A := {       #\label{demo:mixin:rom}#
+#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type A }.        #\label{demo:structure:monoid}#
+
+#\hbmixin{}# Record Ring_of_Monoid A of Monoid A := {       #\label{demo:mixin:rom}#
   one : A;
@@
 }.
-#\hbstructure{}# Definition Ring := { A of Monoid.axioms A & Ring_of_Monoid.axioms A }. #\label{demo:structure:ring}#
+#\hbstructure{}# Definition Ring := { A of Monoid A & Ring_of_Monoid A }. #\label{demo:structure:ring}#
 
@@
 the name \coq{Ring_of_Monoid}. Indeed this time the type variable \coq{A}
-is followed by \coq{Monoid.axioms} that enriches
+is followed by \coq{Monoid} that enriches
 \coq{A} with the operations and properties of monoids. As a consequence
@@
 
-Definition Z_Monoid_axioms : Monoid_of_Type.axioms Z :=                         #\label{demo:theory:z:monoid:axioms}#
-  Monoid_of_Type.Axioms Z 0 Z.add Z.add_assoc Z.add_0_l Z.add_0_r.
+Definition Z_Monoid_axioms : Monoid_of_Type Z :=                         #\label{demo:theory:z:monoid:axioms}#
+  Monoid_of_Type.Build Z 0 Z.add Z.add_assoc Z.add_0_l Z.add_0_r.
 
@@
 
-Definition Z_Ring_axioms : Ring_of_Monoid.axioms Z _ :=                        #\label{demo:theory:z:ring:axioms}#
-  Ring_of_Monoid.Axioms Z 1 Z.opp Z.mul
+Definition Z_Ring_axioms : Ring_of_Monoid Z _ :=                        #\label{demo:theory:z:ring:axioms}#
+  Ring_of_Monoid.Build Z 1 Z.opp Z.mul
     Z.add_opp_diag_l Z.add_opp_diag_r Z.mul_assoc Z.mul_1_l Z.mul_1_r
@@
 
-We use the \coq{Monoid_of_Type.Axioms} \emph{\phantterm{}}
+We use the \coq{Monoid_of_Type.Build} \emph{\phantterm{}}
 (line~\ref{demo:theory:z:monoid:axioms})
@@
 We can similarly declare that \coq{Z} forms a ring by using
-the \coq{Ring_of_Monoid.Axioms} \phantterm{}
+the \coq{Ring_of_Monoid.Build} \phantterm{}
 (lines~\ref{demo:theory:z:ring:axioms} and~\ref{demo:theory:z:ring:canonical}).
-Note that the \coq{Ring_of_Monoid.Axioms} \phantterm{} is not
-a plain record constructor for \coq{Ring_of_Monoid.axioms}, since that
+Note that the \coq{Ring_of_Monoid.Build} \phantterm{} is not
+a plain record constructor for \coq{Ring_of_Monoid}, since that
 would require more arguments, namely the monoid ones (see the \coq{_}
@@
 #\hbmixin{}# Record Monoid_of_Type A := { ... (* unchanged *) ... }.
-#\hbstructure{}# Definition Monoid := { A & Monoid_of_Type.axioms A }.
-
-#\hbmixin{}# Record AbelianGroup_of_Monoid A of Monoid.axioms A := {
+#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type A }.
+
+#\hbmixin{}# Record AbelianGroup_of_Monoid A of Monoid A := {
   opp : A -> A;
@@
 }.
-#\hbstructure{}# Definition AbelianGroup :=
-  { A of Monoid.axioms A & AbelianGroup_of_Monoid.axioms A }.
-
-#\hbmixin{}# Record Ring_of_AbelianGroup A of AbelianGroup.axioms A := {
+#\hbstructure{}# Definition AbelianGroup := { A of Monoid A & AbelianGroup_of_Monoid A }.
+
+#\hbmixin{}# Record Ring_of_AbelianGroup A of AbelianGroup A := {
   one : A;
@@
 }.
-#\hbstructure{}# Definition Ring :=
-  { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.
+#\hbstructure{}# Definition Ring := { A of AbelianGroup A & Ring_of_AbelianGroup A }.
 
@@
 the integers fails, if only
-because we do not have a \coq{Ring_of_Monoid.Axioms} \phantterm{} anymore.
+because we do not have a \coq{Ring_of_Monoid.Build} \phantterm{} anymore.
 
@@
 \begin{coqcode}
-#\hbfactory{}# Record Ring_of_Monoid A of Monoid.axioms A := {
+#\hbfactory{}# Record Ring_of_Monoid A of Monoid A := {
   ... (* unchanged, see page 2 line 13 *) ...
 }.
-#\hbbuilders{}# Context A (f : Ring_of_Monoid.axioms A).                          #\label{demo3:variable:f}#
+#\hbbuilders{}# Context A (f : Ring_of_Monoid A).                          #\label{demo3:variable:f}#
 
@@
   Definition to_AbelianGroup_of_Monoid :=
-    AbelianGroup_of_Monoid.Axioms A (opp f) addrC (addNr f). (* addrN unused *)
+    AbelianGroup_of_Monoid.Build A (opp f) addrC (addNr f). (* addrN unused *)
   #\hbinstance{}# A to_AbelianGroup_of_Monoid.
@@
   Definition to_Ring_of_AbelianGroup :=
-    Ring_of_AbelianGroup.Axioms A (one f) (mul f)
+    Ring_of_AbelianGroup.Build A (one f) (mul f)
       (mulrA f) (mul1r f) (mulr1 f) (mulrDl f) (mulrDr f).
@@
 
-The record \coq{Ring_of_Monoid.axioms} is the same we declared as a \mixin{}
+The record \coq{Ring_of_Monoid} is the same we declared as a \mixin{}
 in version V1. In order to make a \factory{} out of it we equip it with
@@
 so that the next construction \coq{to_Ring_of_AbelianGroup} can call
-the \coq{Ring_of_AbelianGroup.Axioms} \phantterm{} that requires \coq{A} to be
+the \coq{Ring_of_AbelianGroup.Build} \phantterm{} that requires \coq{A} to be
 an \coq{AbelianGroup}.
@@
 Thanks to this \factory{} we can now declare \coq{Z} to be an instance
-of a ring using the \coq{Ring_of_Monoid.Axioms} \phantterm{}.
+of a ring using the \coq{Ring_of_Monoid.Build} \phantterm{}.
 The associated \builder{}s generate, behind the scenes, instances of the
-\coq{Ring_of_AbelianGroup.axioms} and \coq{AbelianGroup_of_Monoid.axioms} \mixins{}
+\coq{Ring_of_AbelianGroup} and \coq{AbelianGroup_of_Monoid} \mixins{}
 that in turn are used to build instances of the \coq{AbelianGroup} and \coq{Ring}
@@
 \begin{coqcode}
-#\hbstructure{}# Definition Ring :=
-  { A of Monoid.axioms A & Ring_of_Monoid.axioms A }.
-#\hbstructure{}# Definition Ring :=
-  { A of AbelianGroup.axioms A & Ring_of_AbelianGroup.axioms A }.
+#\hbstructure{}# Definition Ring := { A of Monoid A & Ring_of_Monoid A }.
+#\hbstructure{}# Definition Ring := { A of AbelianGroup A & Ring_of_AbelianGroup A }.
 \end{coqcode}
@@
   Creates a module \coq{M} with a record
-  \coq{M.axioms}, which depends on the \mixins{}
-  \(\requires{}(\coq{M.axioms}) =
+  \coq{M}, which depends on the \mixins{}
+  \(\requires{}(\coq{M}) =
   \provides{}^\star\left(\enum{f_1}{f_n}\right)\), and registers this
   new record both as a \mixin{} and a \factory{}. It also creates a
-  \phantterm{} \coq{M.Axioms.}
+  \phantterm{} \coq{M.Build.}
 \end{command}
@@
 #\hbfactory{}# Record M T of #$f_1\ \ldots\ f_n$# := { .. }.
-#\hbbuilders{}# Context T (a : M.axioms A).
+#\hbbuilders{}# Context T (a : M A).
 
@@
   Creates a module \coq{M} with a record
-  \coq{M.axioms}, which depends on the mixins
-  \(\requires{}(\coq{M.axioms}) =
+  \coq{M}, which depends on the mixins
+  \(\requires{}(\coq{M}) =
   \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and registers this
@@
   \mixinbuilders{} from them, so that
-  \(\provides{}(\coq{M.axioms}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)
+  \(\provides{}(\coq{M}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)
 
@@
 Record axioms (T : Type) : Type :=
-  Class { Monoid_of_Type_mixin : Monoid_of_Type.axioms T; }.
+  Class { Monoid_of_Type_mixin : Monoid_of_Type T; }.
 Structure type : Type := Pack { sort : Type; class : axioms sort; }.
@@
 Since the monoid structure is the bottom of this hierarchy its class record
-\coq{Monoid.axioms} consists just one \mixin{}. In contrast the class record
+\coq{Monoid} consists just one \mixin{}. In contrast the class record
 of Abelian groups consists of two \mixins{} where the second one depends
@@
 Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin : Monoid_of_Type.axioms T;
-  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid.axioms T Monoid_of_Type_mixin; }.
+  Monoid_of_Type_mixin : Monoid_of_Type T;
+  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
 Structure type : Type := Pack { sort : Type; class : axioms sort }.
@@
 Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
-Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup.axioms T) : #\label{demo4:coe:ag:monoid:class}#
-  Monoid.axioms T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
+Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup T) : #\label{demo4:coe:ag:monoid:class}#
+  Monoid T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
 Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
@@
 #\hbmixin{}# Record SemiRing_of_Monoid A 
-    of Monoid.axioms A := {
+    of Monoid A := {
   one : A;
@@
 #\hbfactory{}# Record Ring_of_AbelianGroup A
-    of AbelianGroup.axioms A := {
+    of AbelianGroup A := {
   (* 2 operators and 5 axioms are omitted. *)
@@
 Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin : Monoid_of_Type.axioms T;
-  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid.axioms T Monoid_of_Type_mixin; }.
+  Monoid_of_Type_mixin : Monoid_of_Type T;
+  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin; }.
 Structure type : Type := Pack { sort : Type; class : axioms sort; }.
@@
 Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin : Monoid_of_Type.axioms T;
-  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid.axioms T Monoid_of_Type_mixin;
-  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid.axioms T Monoid_of_Type_mixin; }.
+  Monoid_of_Type_mixin : Monoid_of_Type T;
+  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin;
+  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
 Structure type : Type := Pack { sort : Type; class : axioms sort; }.

### 2020-03-11 elpi.py

@@ -161,7 +161,7 @@ class CoqElpiLexer(RegexLexer):
     # 'raise', 'rec', 'sig', 'struct', 'then', 'to', 'true', 'try',
     # 'type', 'val', 'virtual', 'when', 'while', 'with'
     keyopts = (
-        '!=', '#', '&', '&&', r'\(', r'\)', r'\*', r'\+', ',', '-', r'-\.',
+        '!=', '#', '&&', r'\(', r'\)', r'\*', r'\+', ',', '-', r'-\.',
         '->', r'\.', r'\.\.', ':', '::', ':=', ':>', ';', ';;', '<', '<-',
         '<->', '=', '>', '>]', r'>\}', r'\?', r'\?\?', r'\[', r'\[<', r'\[>',
         r'\[\|', ']', '_', '`', r'\{', r'\{<', r'\|', r'\|]', r'\}', '~', '=>',
@@ -186,6 +186,7 @@ class CoqElpiLexer(RegexLexer):
             (words(keywords5, prefix=r'\b', suffix=r'\b'), Keyword.Pseudo),
             (words(keywords6, prefix=r'\b', suffix=r'\b'), Keyword.Reserved),
             # (r'\b([A-Z][\w\']*)(\.)', Name.Namespace, 'dotted'),
+            (r'&', Keyword),
             (r'\b([A-Z][\w\']*)', Name),
             (r'(%s)' % '|'.join(keyopts[::-1]), Operator),
             (r'(%s|%s)?%s' % (infix_syms, prefix_syms, operators), Operator),

### 2020-03-11 hb.tex
fix minor issues

@@
 \emph{evolve} without breaking user code. The key concepts are the ones of
-\emph{\factory{}} and \emph{\phantterm{}} that let the hierarchy developer describe an actual
+\emph{\factory{}}, \emph{\builder{}} and \emph{\phantterm{}} that let the hierarchy developer describe an actual
 interface for their library. Behind that interface the developer can provide
@@
 later assemble them. One kind of \factory{} supported by \HB{}, the simplest
-one, is called \emph{\mixin{}} and is embodied by a record, by convention named
-\coq{axioms}, that gathers operations and properties.
+one, is called \emph{\mixin{}} and is embodied by a record
+that gathers operations and properties.
 
@@
 that takes a record declaration with
-a type variable name and a possibly empty list of \factories{} for that type.
+a type parameter and a possibly empty list of \factories{} for that type.
 The code between lines~\ref{demo:mixin:mot}
@@
 
-The \hbstructure{} command takes in input the definition S for a type
-\coq{A} together with a list of \factories{} for \coq{A}.
+The \hbstructure{} command takes in input a definition for a sigma type S
+that equips \coq{A} with a list of \factories{}.
 It registers a structure S in the hierarchy placing
@@
 \phantterm{} synthesized by \HB{} infers them automatically
-(as in~\cite[Section 7]{DBLP:conf/itp/MahboubiT13}) thanks to the \coq{hb.canonical}
+(as in~\cite[Section 7]{DBLP:conf/itp/MahboubiT13}) thanks to the \hbinstance{}
 declaration given just above.

### 2020-03-30 rebuttal.txt
Import initial rebuttal by Enrico
Co-Authored-By: Enrico Tassi <enrico.tassi@inria.fr>

Here the message I propose, it is ~600 words so there is space if you
want to write more:

Dear PC members, first of all we would like to thank the reviewers for
their valuable feedback. We believe we will be able to address all
their requests for the final version, namely clarify the exposition in
sections 4 and 5 (as per review 1 and 2), and add a few extra details
in section 6 (as per review 2).

Specific answer to reviewer 1
-----------------------------

- p4-5, evolution of the hierarchy:
     Could you make clearer the distinction (or absence of) between
     axioms and theorems in your implementation of a theory (problem
     mentioned on line 121). I.e., does it matter or not for the user
     of the hierarchy.

It does not matter: addrN is name he can use in her proofs. What
changed is how it is "implemented" by the hierarchy designer. Line
121 is very much related to the remark at lines 421-425, we shall
hint that in section 2 as well.

- Propositions 5 and 8: there is no proof, but is there anything to
prove?

We agree that 5 is a remark. On the contrary 8 shows M \subset F
by providing an implementation of requires, provides and from.
This proposition identifies *factories* as the central notion of HB,
since mixins are just a special case. Trivia: in an early draft the
name of the language was, indeed, Factory.

- Section 4:
     What are you describing here?
     Is it the code your tool generate?
     Does it correspond (exactly) to examples in Section 2?
     Has anything been omitted?

4.1 show Coq code that is equivalent to the HB commands issued
in Section 2.
HB commands do not synthesize that text, but have the same effect.

Many details were omitted. For example the abbreviations mentioned in
section 3.3 are not there. Another omission is that all "exported"
notions are gathered into an inner module, eg Monoid.Exports, etc.
We believe these details are not very interesting, but we will clarify
the status of the Coq code in section 4.

General answer to all reviewers
-------------------------------
All reviewers acknowledge that HB looks useful. Reviewer 1 flatters
us with "the tool is invaluable" and acknowledges that a contribution
of HB is "providing a way for normal Coq users to write hierarchies
using the packed class discipline". Reviewer 2 finds it "promising".
Reviewer 3 finds the machinery "impressive".

The main critic to HB, by Reviewer 3, is that it looks not only
impressive,
but also intimidating and Reviewer 3 wonders if a Coq user would run
the
risk of depending on HB in his work.

We use the remaining space to comment on that, that we believe is a
very pertinent remark.

In an early prototype of HB we considered generating text (Coq source
code).
We believe this approach could let users take advantage of HB without
depending on it (permanently). We do not provide this feature in the
current code because, honestly, it is not the most intriguing feature
to
implement or to describe in a paper.

We shall consider that feature again in the future, as an extreme
measure,
if the software stack behind HB turns out to severely hinder its
adoption.
We hope Reviewer 1 is right in saying that even regular Coq users could
be interested in using HB, and if the tool turns out to be popular its
maintenance will be assured by its authors and the community of users,
as it is the case for the Coq system or the Mathematical Components
library.

### 2020-03-30 rebuttal.txt
reworked from google docs

@@
-Here the message I propose, it is ~600 words so there is space if you
-want to write more:
+Dear PC members, first of all we would like to thank the reviewers for their
+valuable feedback. We believe we will be able to address all their requests for
+the final version, namely clarify the exposition in sections 4 and 5 (as per
+review 1 and 2), and add a few extra details in section 6 (as per review 2).
 
-Dear PC members, first of all we would like to thank the reviewers for
-their valuable feedback. We believe we will be able to address all
-their requests for the final version, namely clarify the exposition in
-sections 4 and 5 (as per review 1 and 2), and add a few extra details
-in section 6 (as per review 2).
-
-Specific answer to reviewer 1
+Answer to reviewer 1
 -----------------------------
@@
 
-It does not matter: addrN is name he can use in her proofs. What
-changed is how it is "implemented" by the hierarchy designer. Line
-121 is very much related to the remark at lines 421-425, we shall
-hint that in section 2 as well.
+There are two distinct use cases for a hierarchy:
+1. one assumes a structure as
+     an hypothesis. In this case, all its axioms are theorems so the distinction
+     axiom-theorem does not matter.
+2. one instantiates a structure on a concrete
+     type in order to access all its theory. Here, however, one must pick an
+     axiomatic (i.e. a set of axioms) to show Coq that some type is indeed an
+     instance of a structure. If for each structure there is only one axiomatic
+     (which is its internal representation) and if we keep changing it, some theorems
+     may become axioms and increase the burden on the user (this is what happens with
+     addrC in V2, in order to instantiate a ring one has to prove it, while it was
+     not required in V1). We fix that by letting the designer provide as many
+     axiomatics as she wants, so that the user can pick the most convenient one, in
+     particular V3 also provides the axiomatic from V1 as a factory. Behind the
+     scenes, hierarchy-builder compiles it to the axiomatic currently used as the
+     internal representation (made of mixins).
+
+We plan to integrate a variation of the text above as the transition from
+section 2.1 to 2.2.
 
 - Propositions 5 and 8: there is no proof, but is there anything to
-prove?
+Prove?
 
-We agree that 5 is a remark. On the contrary 8 shows M \subset F
-by providing an implementation of requires, provides and from.
-This proposition identifies *factories* as the central notion of HB,
-since mixins are just a special case. Trivia: in an early draft the
-name of the language was, indeed, Factory.
+Albeit trivial, they are both theorems. Our mistake was to mix the proof with
+the statement.
+
+Proposition 5) dep represents a transitively closed DAG, in particular
+  dep*(dep*(M)) ⊆ dep*(M).
+Proof.
+dep is transitively closed since records are well typed in the empty context.
+dep describes a DAG since Coq does not admit circular definitions.
+Qed.
+
+Proposition 8) There is a way to see mixins as factories (M ⊆  F).
+Proof.
+For each m ∈ M we pose
+* requires(m) = dep(m)
+* provides(m) = {m}
+* from (m, m) = fun T (p : m T p σ ) (x : m T p 1 . . . p n ) => x ∈ B
+Qed.
+
+This proposition identifies *factories* as the central notion of HB, since
+mixins are just a special case. Trivia: in an early draft the name of the
+language was, indeed, Factory.
 
@@
          What are you describing here?
-     Is it the code your tool generate?
+         Is it the code your tool generates?
          Does it correspond (exactly) to examples in Section 2?
@@
 
-4.1 show Coq code that is equivalent to the HB commands issued
-in Section 2.
-HB commands do not synthesize that text, but have the same effect.
+Section 4.1 shows Coq code that is equivalent to the HB commands given in
+Section 2. These commands do not synthesize that text, but have the same effect.
 
-Many details were omitted. For example the abbreviations mentioned in
-section 3.3 are not there. Another omission is that all "exported"
-notions are gathered into an inner module, eg Monoid.Exports, etc.
-We believe these details are not very interesting, but we will clarify
-the status of the Coq code in section 4.
+Many details were omitted:
+* the abbreviations mentioned in section 3.3
+* the declarations of canonical structures are postponed to 4.3
+* "Exported" notions are gathered into an inner module, eg Monoid.Exports.
 
-General answer to all reviewers
+We believe these details are not very interesting, but we will clarify the
+status of the Coq code in section 4 in the final version.
+
+Comment to all reviewers
 -------------------------------
-All reviewers acknowledge that HB looks useful. Reviewer 1 flatters
-us with "the tool is invaluable" and acknowledges that a contribution
-of HB is "providing a way for normal Coq users to write hierarchies
-using the packed class discipline". Reviewer 2 finds it "promising".
-Reviewer 3 finds the machinery "impressive".
 
-The main critic to HB, by Reviewer 3, is that it looks not only
-impressive,
-but also intimidating and Reviewer 3 wonders if a Coq user would run
-the
-risk of depending on HB in his work.
+We could not avoid to remark that, by lack of clarity on our side, an important
+aspect of HB was not sufficiently put forward. Indeed two reviewers write
+seemingly opposite remarks:
 
-We use the remaining space to comment on that, that we believe is a
-very pertinent remark.
+* Reviewer 1 : "The first [contribution], surprisingly little advertised, [is]
+     providing a way for normal Coq users to write hierarchies using the packed class
+     discipline"
 
-In an early prototype of HB we considered generating text (Coq source
-code).
-We believe this approach could let users take advantage of HB without
-depending on it (permanently). We do not provide this feature in the
-current code because, honestly, it is not the most intriguing feature
-to
-implement or to describe in a paper.
+* Reviewer 3 :  “I can imagine a few experts using these tools to build
+     hierarchies, but I would not expect the average Coq user to invest time into
+     learning about them”.
 
-We shall consider that feature again in the future, as an extreme
-measure,
-if the software stack behind HB turns out to severely hinder its
-adoption.
-We hope Reviewer 1 is right in saying that even regular Coq users could
-be interested in using HB, and if the tool turns out to be popular its
-maintenance will be assured by its authors and the community of users,
-as it is the case for the Coq system or the Mathematical Components
-library.
+We agree sections 4 and 5 of the paper can look intimidating to a non Coq +
+Packed Classes expert, but these details are essentially the ones a user of HB
+does not need to understand in order to use the tool. Indeed, the implementation
+of HB takes care of these details and avoids most of the common mistakes one can
+hardly avoid when writing Packed Classes code by hand. In the final version of
+the paper, we intend to clarify this aspect in the introduction.
+
+Finally, It is true that the software stack is large and may discourage the
+adoption of HB, as remarked by Reviewer 3. In an early prototype of HB we
+considered generating text (Coq source code). We believe this approach could let
+users take advantage of HB without depending on it (permanently). We do not
+provide this feature in the current code base because, honestly, it is not the
+most intriguing feature to implement or to describe in a paper. We shall
+consider that feature again in the future, as an extreme measure, if the
+software stack behind HB turns out to severely hinder its adoption. However, we
+hope Reviewer 1 is right in saying that the tool is invaluable, and if the tool
+turns out to be popular its maintenance will be assured by its authors and the
+community of users, as it is the case for the Coq system or the Mathematical
+Components library.

### 2020-03-30 rebuttal.txt
typos, clarifications, typesetting

@@
 
-Answer to reviewer 1
------------------------------
+*Answer to reviewer 1*
 
-- p4-5, evolution of the hierarchy:
-         Could you make clearer the distinction (or absence of) between
-         axioms and theorems in your implementation of a theory (problem
-         mentioned on line 121). I.e., does it matter or not for the user
-         of the hierarchy.
+- p4-5, evolution of the hierarchy: Could you make clearer the
+  distinction (or absence of) between axioms and theorems in your
+  implementation of a theory (problem mentioned on line 121). I.e.,
+  does it matter or not for the user of the hierarchy.
 
 There are two distinct use cases for a hierarchy:
-1. one assumes a structure as
-     an hypothesis. In this case, all its axioms are theorems so the distinction
-     axiom-theorem does not matter.
-2. one instantiates a structure on a concrete
-     type in order to access all its theory. Here, however, one must pick an
-     axiomatic (i.e. a set of axioms) to show Coq that some type is indeed an
-     instance of a structure. If for each structure there is only one axiomatic
-     (which is its internal representation) and if we keep changing it, some theorems
-     may become axioms and increase the burden on the user (this is what happens with
-     addrC in V2, in order to instantiate a ring one has to prove it, while it was
-     not required in V1). We fix that by letting the designer provide as many
-     axiomatics as she wants, so that the user can pick the most convenient one, in
-     particular V3 also provides the axiomatic from V1 as a factory. Behind the
-     scenes, hierarchy-builder compiles it to the axiomatic currently used as the
-     internal representation (made of mixins).
+1. one assumes a structure as an hypothesis. In this case, all its
+   axioms are theorems so the distinction axiom-theorem does not
+   matter.
+2. one instantiates a structure on a concrete type in order to access
+   all its theory. Here, however, one must pick an axiomatic (i.e. a
+   set of axioms) to show Coq that some type is indeed an instance of
+   a structure. If for each structure there is only one axiomatic
+   (which is its internal representation) and if we keep changing it
+   across versions, some theorems may become axioms and increase the
+   burden on the user (this is what happens with addrC, in order to
+   instantiate a ring in V2 one has to prove it, while it was not
+   required in V1). We fix that by letting the designer provide as
+   many axiomatics (i.e. factories) as she wants, so that the user can
+   pick the most convenient one, in particular V3 also provides the
+   axiomatic from V1. Behind the scenes, hierarchy-builder compiles it
+   to the axiomatic currently used as the internal representation
+   (made of mixins).
 
@@
 
-- Propositions 5 and 8: there is no proof, but is there anything to
-Prove?
+- Propositions 5 and 8: there is no proof, but is there anything to Prove?
 
 Albeit trivial, they are both theorems. Our mistake was to mix the proof with
-the statement.
+the statement. Here is the modification we will make:
 
@@
 
-This proposition identifies *factories* as the central notion of HB, since
-mixins are just a special case. Trivia: in an early draft the name of the
-language was, indeed, Factory.
+The latter proposition identifies *factories* as the central notion of
+HB, since mixins are just a special case. Trivia: in an early draft
+the name of the language was, indeed, Factory.
 
-- Section 4:
-         What are you describing here?
-         Is it the code your tool generates?
-         Does it correspond (exactly) to examples in Section 2?
+- Section 4: What are you describing here?  Is it the code your tool
+  generates?  Does it correspond (exactly) to examples in Section 2?
   Has anything been omitted?
 
-Section 4.1 shows Coq code that is equivalent to the HB commands given in
+Section 4.1 shows Coq code that is almost equivalent to the HB commands given in
 Section 2. These commands do not synthesize that text, but have the same effect.
@@
 Many details were omitted:
-* the abbreviations mentioned in section 3.3
-* the declarations of canonical structures are postponed to 4.3
-* "Exported" notions are gathered into an inner module, eg Monoid.Exports.
+- the abbreviations mentioned in section 3.3
+- the declarations of canonical structures are postponed to 4.3
+- "Exported" notions are gathered into an inner module, e.g. Monoid.Exports.
 
-We believe these details are not very interesting, but we will clarify the
+We believe these details are not very interesting, and we will clarify the
 status of the Coq code in section 4 in the final version.
 
-Comment to all reviewers
--------------------------------
+*Comment to all reviewers*
 
@@
 seemingly opposite remarks:
-
-* Reviewer 1 : "The first [contribution], surprisingly little advertised, [is]
-     providing a way for normal Coq users to write hierarchies using the packed class
-     discipline"
-
-* Reviewer 3 :  “I can imagine a few experts using these tools to build
-     hierarchies, but I would not expect the average Coq user to invest time into
-     learning about them”.
+- Reviewer 1 : "The first [contribution], surprisingly little
+   advertised, [is] providing a way for normal Coq users to write
+   hierarchies using the packed class discipline"
+- Reviewer 3 : “I can imagine a few experts using these tools to build
+   hierarchies, but I would not expect the average Coq user to invest
+   time into learning about them”.
 

### 2020-03-31 rebuttal.txt
Clearly quote the reviewers comments without any modification (except for "comment to all reviewers")

@@
 
-- p4-5, evolution of the hierarchy: Could you make clearer the
-  distinction (or absence of) between axioms and theorems in your
-  implementation of a theory (problem mentioned on line 121). I.e.,
-  does it matter or not for the user of the hierarchy.
+> p4-5, evolution of the hierarchy:
+> Could you make clearer the distinction (or absence of) between
+> axioms and theorems in your implementation of a theory (problem
+> mentioned on line 121). I.e., does it matter or not for the user of
+> the hierarchy.
 
@@
 
-- Propositions 5 and 8: there is no proof, but is there anything to Prove?
+> Propositions 5 and 8: there is no proof, but is there anything to Prove?
 
@@
 
-- Section 4: What are you describing here?  Is it the code your tool
-  generates?  Does it correspond (exactly) to examples in Section 2?
-  Has anything been omitted?
+> Section 4:
+> What are you describing here?
+> Is it the code your tool generate?
+> Does it correspond (exactly) to examples in Section 2?
+> Has anything been omitted?
 

### 2020-04-17 hb.tex
all items for =ET= except the one on section 5 are done

@@
 is implemented following the discipline of Packed Classes initially introduced
-in \cite{DBLP:conf/tphol/GarillotGMR09} and later also adopted in
-\cite{DBLP:conf/mpc/AffeldtNS19} and \cite{DBLP:journals/mics/BoldoLM15}.
+by Garillot, Gonthier, Mahboubi and Rideau~\cite{DBLP:conf/tphol/GarillotGMR09}
+and later also adopted by Affeldt, Nowak, and Saikawain
+to describe a hierarchy of monadic effects~\cite{DBLP:conf/mpc/AffeldtNS19}
+and by Boldo, Lelay, and Melquiond in the Coquelicot library of real
+analysis~\cite{DBLP:journals/mics/BoldoLM15}.
 We call Packed Classes a discipline, and not a language, because, in spite of
@@
 (line~\ref{demo:theory:z:monoid:canonical}) via the command
-\coq{hb.canonical}.
+\coq{HB.instance}.
 We can similarly declare that \coq{Z} forms a ring by using
@@
 since it expects the type \coq{A} to be already an Abelian group and hence
-gathers fewer axioms. \\
+gathers fewer axioms.
 While operations moved from one structure to another, some properties
@@
 an axiom of rings, but rather a theorem of the abstract theory of Abelian
-groups.
-
-With this new version of the hierarchy, that we name V2, code written
-for version V1 breaks. For example the declaration of the canonical ring over
-the integers fails, if only
-because we do not have a \coq{Ring_of_Monoid.Build} \phantterm{} anymore.
+groups proved at line~\ref{demo2:proof:addrN}.
+
+It is worth clarifying here what axioms and lemmas are by looking at the
+two distinct use cases for a hierarchy: 1) develop the abstract theory of some
+structure; 2) apply the abstract theory on some concrete example.
+In the first case
+all the axioms of the structure are (assumed to be) theorems so the distinction
+between axiom and theorem does not matter. This is what happens
+in the proof of \coq{addNr} by taking \coq{R} to be of type
+\coq{AbelianGroup.type} as an assumption.
+In the second case,
+in order to show Coq that a data type and some operations for a structure,
+one must pick the axioms of the structure an prove them for that specific
+type and operations. For example the \coq{Ring_of_Monoid.Build} \phantterm{}
+is used at page~\pageref{demo:theory:z:ring:axioms} to package the set
+of proofs that make \coq{Z} a ring (proviso \coq{Z} is already a monoid).
+
+With this new version of the hierarchy, that we name V2, the axiomatic that
+was previously exposed to user changed, and indeed
+code written for version V1 breaks. For example the declaration of the
+canonical ring over the integers fails, if only
+because we do not have a \coq{Ring_of_Monoid.Build} \phantterm{}
+anymore.
 
@@
   \begin{center}
-    \includegraphics[width=0.8\textwidth]{puzzle.pdf}
+    \includegraphics[width=\textwidth]{puzzle.pdf}
   \end{center}
@@
 
+\begin{coqcode}
+#\hbmixin{}# Record SemiRing_of_Monoid A of Monoid A := {
+  one : A;
+  mul : A -> A -> A;
+  (* 7 axioms are omitted. *)
+}.
+\end{coqcode}
+
 \begin{wrapfigure}[3]{r}{.4\textwidth}
@@
 \begin{coqcode}
-#\hbmixin{}# Record SemiRing_of_Monoid A 
-    of Monoid A := {
-  one : A;
-  mul : A -> A -> A;
-  (* 7 axioms are omitted. *)
-}.
-
 #\hbfactory{}# Record Ring_of_AbelianGroup A
@@
 Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin : Monoid_of_Type T;
-  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin; }.
+  Monoid_of_Type_mixin :
+    Monoid_of_Type T;
+  SemiRing_of_Monoid_mixin :
+    SemiRing_of_Monoid T Monoid_of_Type_mixin;
+}.
 Structure type : Type := Pack { sort : Type; class : axioms sort; }.
@@
 The last detail worth mentioning is that this program makes no use of
-backtraking: the \elpi{std.do!} combinator signals that.
+backtracking: the \elpi{std.do!} combinator signals that.
 
@@
 of unification hints~\cite{10.1007/978-3-642-03359-9_8}
-(a super set of the one of Canonical Structures),
-but we could not find any written trace of that. Language wise they provides
+(a superset of the one of Canonical Structures),
+but we could not find any written trace of that. Language-wise they provide
 keywords such as \verb+combine+ and \verb+over+ to share (reuse) a property
@@
 One operation \HB{} allows for but that does not seem to be possible in the
-setting of Presentation Combinators is the one of replacing an axioms with a
+setting of Presentation Combinators is the one of replacing an axiom with a
 lemma and vice versa. As shown in \autoref{subsec:evolution} \HB{} supports

### 2020-04-23 coqws/easychair.tex
init ws

% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

\usepackage{doc}
\usepackage{minted} % pip3 install pygments, pdflatex -shell-escape hb
\newminted[elpicode]{elpi.py:ElpiLexer -x}{linenos=true,fontsize=\footnotesize,escapeinside=\%\%}
\newmintinline[elpi]{elpi.py:ElpiLexer -x}{fontsize=\small}
\newminted[coqcode]{elpi.py:CoqElpiLexer -x}{linenos=true,fontsize=\footnotesize,mathescape,escapeinside=\#\#}
\newmintinline[coq]{elpi.py:CoqElpiLexer -x}{fontsize=\small}
\definecolor{dkgreen}{rgb}{0, 0.502, 0}
\bibliographystyle{plainurl}% the mandatory bibstyle

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Hierarchy Builder}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%

\author{Cyril Cohen\inst{1} \and Kazuhiko Sakaguchi\inst{2} \and Enrico Tassi\inst{3}}

\institute{
 Inria, Universit\'e C\^ote d'Azur, France\\ \email{Cyril.Cohen@inria.fr}
\and
 University of Tsukuba, Japan\\ \email{sakaguchi@logic.cs.tsukuba.ac.jp}
\and
 Inria, Universit\'e C\^ote d'Azur, France\\\ \email{Enrico.Tassi@inria.fr}
}

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{C.\,Cohen and K.\,Sakaguchi and E.\,Tassi}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{Hierarchy Builder}

\begin{document}

\maketitle

\begin{abstract}
  We present a tool that makes it easy to declare and evolve hierarchies
  of algebraic structures in Coq.
\end{abstract}

% The table of contents below is added for your convenience. Please do not use
% the table of contents if you are preparing your paper for publication in the
% EPiC Series or Kalpa Publications series

%\setcounter{tocdepth}{2}
%{\small
%\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

\section{Context}

bla bla librarires.
bla bla structures.
bla bla packed classes.
bla bla difficult.
bla bla HB language.
bla bla imlemented in Elpi.
In the talk we give a demo of the tool, details are in the paper cite FSCD.

\newpage
\section{An example of HB}

\begin{coqcode}
HB.mixin Record AddComoid_of_Type A := {
  zero : A;
  add : A -> A -> A;
  addrA : forall x y z, add x (add y z) = add (add x y) z;
  addrC : forall x y, add x y = add y x;
  add0r : forall x, add zero x = x;
}.

HB.structure Definition AddComoid := { A of AddComoid_of_Type A }.

Notation "0" := zero.
Infix "+" := add.

Check forall (M : AddComoid.type) (x : M), x + x = 0.
\end{coqcode}

where we are, where we go

\begin{coqcode}
HB.mixin Record AbelianGrp_of_AddComoid A of AddComoid A := {
  opp : A -> A;
  addNr : forall x, opp x + x = 0;
}.

HB.structure Definition AbelianGrp := { A of AbelianGrp_of_AddComoid A & }.

Notation "- x" := (opp x).

Lemma example (G : AbelianGrp.type) (x : G) : x + (- x) = - 0.
Proof. by rewrite addrC addNr -[LHS](addNr zero) addrC add0r. Qed.
\end{coqcode}

and now examples

\begin{coqcode}

Definition Z_CoMoid := AddComoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
HB.instance Z Z_CoMoid.
Definition Z_AbGrp := AbelianGrp_of_AddComoid.Build Z Z.opp Z.add_opp_diag_l.
HB.instance Z Z_AbGrp.

Lemma example2 (x : Z) : x + (- x) = - 0.
Proof. by rewrite example. Qed.
\end{coqcode}

\bibliography{biblio}

\end{document}

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}

The {\easychair} class was designed to be easy to use, and specifically favors
electronic and on-paper publishing by the EasyChair conference system
\cite{easychair}, including the
\href{http://www.easychair.org/publications/EPiC}{EasyChair EPiC
  Series} and \href{http://www.easychair.org/publications/Kalpa}{EasyChair Kalpa
  Publications}. 
EasyChair is a free conference management system that is flexible, easy to use,
and has many features to make it suitable for various conference models. It is
currently probably the most commonly used conference management system
\cite{easychair}.

The {\easychair} class was designed according to some requirements, which
are described in Appendix~\ref{sect:easychair-requirements}. 
An article that occupies approximately 15 LNCS-formatted pages
takes up approximately 14 {\easychair} pages.

\begin{figure}[tb]
	\begin{centering}
	\includegraphics[width=0.5\textwidth]{logoEC}
	\caption{EasyChair logo}
	\label{fig:easychair-logo}
	\end{centering}
\end{figure}

%------------------------------------------------------------------------------
\section{Typesetting}
\label{sect:typesetting}

Typesetting with {\easychair} is, well, easy.  Just by using the
document class entry in the document's preamble as follows:
\verb+\documentclass{easychair}+ the typesetting work is nearly done.
The {\easychair} class is a relatively conservative extension of the
standard \texttt{article} class, so most of the environments, section
headers, etc. defined by \texttt{article} are available.

%------------------------------------------------------------------------------
\subsection{Generalities}
\label{sect:generalities}

The following are the general default parameters {\easychair}
introduces into the typesetting aspect of articles. If you use
\easychair\ for proceedings or other kinds of publishing through
EasyChair, do not alter these -- papers deviating from the formatting
standards will be rejected by EasyChair.

\begin{enumerate}
\item
The default paper size is US letter. It can be explicitly set to A4 
(\texttt{a4paper}) or letter (\texttt{letterpaper}) paper in the
document class entry, e.g.:\\\verb+\documentclass[a4paper]{easychair}+

\item
The print area for both letter and A4 paper sizes is 145x224 mm. This size
has been selected to allow for inexpensive printing using our current
print-on-demand publisher.

\item
The base font is Computer Modern. The base font size is 10pt. If you
use any other font size, there is no guarantee that the produced
document will look nice or fit into our standard page size.

\item
The references list is condensed. The default bibliography styles, such as
\texttt{plain}, \texttt{abbrv}, and \texttt{alpha}, are suggested.

\item
PNG, JPG, and PDF images are supported, i.e., those that are supported
by the standard \texttt{graphicx} package \cite{graphicx-package}, and
render nicely in online versions of PDF documents.  This document
shows some examples of JPG and PDF images, for example in
Figure~\ref{fig:easychair-logo}. If the papers are designed for
publishing in print, the images should be at least 300dpi in
resolution.

\end{enumerate}

%------------------------------------------------------------------------------
\subsection{Front Matter}
\label{sect:front-matter}

The front matter of an {\easychair} article follows the \texttt{article}
style, augmented with the \verb+\titlerunning+ and \verb+\authorrunning+
commands for use by authors. 
For the \verb+\author+ command with multiple authors, use \verb+\and+ to
separate authors from different institutions, as done in this document.
Institutions are defined using \verb+\institute+ in a similar way and affiliations of authors assigned using \verb+\inst+.
Authors must set the \verb+\titlerunning+ and \verb+\authorrunning+.
For example, the front matter of this document defined the authors
and title as follows:

%---------------------------------------------------------------

\noindent\small
\begin{verbatim}
\title{The {\easychair} Class File\\
       Documentation and Guide for Authors%
\thanks{Other people who contributed to this document include Maria Voronkov
  (Imperial College and EasyChair) and Graham Gough (The University of
  Manchester).}}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%
\author{
Serguei A. Mokhov\inst{1}\thanks{Designed and implemented the class style}
\and
   Geoff Sutcliffe\inst{2}\thanks{Did numerous tests and provided a lot of suggestions}
\and
   Andrei Voronkov\inst{3}\inst{4}\inst{5}\thanks{Masterminded EasyChair and created 
     versions 3.0--3.5 of the class style}
}

% Institutes for affiliations are also joined by \and,
\institute{
  Concordia University,
  Montreal, Quebec, Canada\\
  \email{mokhov@cse.concordia.ca}
\and
   University of Miami,
   Miami, Florida, U.S.A.\\
   \email{geoff@cs.miami.edu}\\
\and
   University of Manchester,
   Manchester, U.K.\\
   \email{andrei@voronkov.com}\\
\and
   Chalmers University of Technology,
   Gothenburg, Sweden
\and
   EasyChair
 }

\authorrunning{Mokhov, Sutcliffe and Voronkov}
\titlerunning{The {\easychair} Class File}
\end{verbatim}
\normalsize

%------------------------------------------------------------------------------
\subsection{Page Numbering}
\label{sect:page-numbering}

Page numbers are at the bottom of every page. Authors must leave the
page numbers in as-is. When EasyChair proceedings and Procedia volumes are processed by
EasyChair, the correct volume page numbers will be inserted
automatically.

%------------------------------------------------------------------------------
\subsection{Section Headings and Capitalization}
\label{sect:section-headings}

Section and paragraph headings in {\easychair} are invoked via the standard 
commands, such as
\verb+\section+,
\verb+\subsection+,
\verb+\subsubsection+, and
\verb+\paragraph+.
Generally, every non-trivial word in a heading must be capitalized according to
general capitalization guidelines. A reasonable rule to use is that
all prepositions, coordinating conjunctions and articles having four
or fewer letters should not be capitalized. If you do not know what it
means, simply do not capitalize the following words:
\textit{amid, anti, as, at, atop, but, by, down, for, from, in, into, like,
near, next, of, off, on, onto, out, over, pace, past, per, plus, qua,
save, than, till, to, up, upon, via, with, for, and, nor, but, or,
yet, so, a, an, the.} For example, if you want to call your paper
``oldest but goldest'', then the proper title for it is ``Oldest but
Goldest''. ``Oldest but goldest'' is wrong (since ``goldest'' is not
capitalized) and ``Oldest But Goldest'' is wrong (``but'' should stay
lower-cased since it belongs to the list of words above.) Needless to
say, ``OLDEST BUT GOLDEST'' is very wrong.

Paragraph headings should not be capitalized and should have a
trailing period. That is, you should write

\begin{quote}
\verb|\paragraph{EasyChair is cool.}|
\end{quote}
rather than

\begin{quote}
\verb|\paragraph{EasyChair is cool}|
\end{quote}
unless your aim was to write something like

\begin{quote}
\verb|\paragraph{EasyChair is cool} when you use it for publishing.| 
\end{quote}
Of course, you are welcome to replace the trailing period by any other
punctuation mark, for example

\begin{quote}
\verb|\paragraph{EasyChair is cool!}|
\end{quote}

See the examples in this document, e.g.,
Section~\ref{sect:typesetting} is a section, this 
(Section~\ref{sect:section-headings}) is a subsection, and
Section~\ref{sect:subsubsection-headings} is a subsubsection.

%------------------------------------------------------------------------------
\subsubsection{Subsubsection Header}
\label{sect:subsubsection-headings}

This is a subsubsection.

\paragraph{Paragraph header.}

This is a paragraph. 
One way of saving space when hyper-references are not essential is to 
use paragraphs instead of subsubsections.

%------------------------------------------------------------------------------
\subsection{Mathematics}
\label{sect:mathematics}

Mathematics can be done inline for simple things, e.g., an equation
$x = 0$, possibly with super and subscripts, e.g., $x^2_k \approx 27$,
Greek letters, e.g., $\alpha \cup \Theta \ne \gamma$, etc.
Larger formulae must be done using {\tt \verb|\|[}~~{\tt \verb|\|]}
bracketing, e.g.,
\[
\int_{0}^{1} x dx = \left[ \frac{1}{2}x^2 \right]_{0}^{1} = \frac{1}{2}
\]
or using {\tt \verb|\|begin\{equation\}} and {\tt \verb|\|end\{equation\}} for
numbered equations, e.g.,
\begin{equation}
e^x = \sum_{n=0}^\infty \frac{x^n}{n!} = \lim_{n\rightarrow\infty} (1+x/n)^n
\end{equation}

Use {\tt \verb|\|begin\{align*\}} and {\tt \verb|\|end\{align*\}} (or without
the {\tt *} include number) to align equations, e.g.,
\begin{align*}
x^2 + y^2 &= 1 \\
y &= \sqrt{1 - x^2}
\end{align*}

Fonts, using \verb|\matcal| and others can also be used in the math mode: $\mathcal{ALC}$.

%------------------------------------------------------------------------------
\subsection{Tables}
\label{sect:tables}

Table~\ref{tab:ltbexample} shows an example of a table of data that was
conveniently available (i.e., the data has nothing to do with
{\easychair}) apart from being related to the research of two of its
authors.

\begin{table}[htp]
	\begin{centering}
		\begin{tabular}{lrrrrrrrr}
		\hline
		ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
		                      & /100& time & out & Con. & Eff. & /35 & /40 & /25 \\
		\hline
		Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
		iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
		SInE 0.4              &  64 & 75.3 &  64 & 0.32 &  8.5 &  26 &  13 &  25 \\
		leanCoP-SInE 2.1      &  35 &110.8 &  35 & 0.23 &  3.2 &  23 &   1 &  11 \\
		E-LTB 1.1pre          &  18 & 63.4 &   0 & 0.21 &  2.8 &   7 &   9 &   2 \\
		EP-LTB 1.1pre         &  18 & 77.8 &  18 & 0.21 &  2.3 &   7 &   9 &   2 \\
		E-KRH'-LTB 1.1.3      &   0 &   -- &  -- &   -- &   -- &   0 &   0 &   0 \\
		\hline
		\end{tabular}
		\caption{LTB division results}
		\label{tab:ltbexample}
	\end{centering}
\end{table}

%------------------------------------------------------------------------------
\subsection{References}
\label{sect:references}

References must be provided in a {\tt .bib} file, so that \BibTeX\ can
be used to generate the references in a consistent style in a volume.
The preferred styles are {\tt plain} and {\tt alpha}.
For example, the references for this paper are generated from the
lines
\begin{verbatim}
    \bibliographystyle{plain}
    \bibliography{easychair}
\end{verbatim}
and a way to compose the entires, e.g. citing this class style~\cite{easychair-latex-class}
is below:
\small
\begin{verbatim}
    @misc
    {
      easychair-latex-class,
      author       = {Serguei A. Mokhov and Geoff Sutcliffe and Andrei Voronkov},
      title        = {The {\sf easychair} Class File Documentation and Guide
                      for Authors},
      year         = {2008--2011},
      howpublished = {[online]},
      note         = {Available at \url{http://easychair.org/easychair.zip}}
    }
\end{verbatim}
\normalsize

%------------------------------------------------------------------------------
\section{Installation and Usage Instructions}
\label{sect:installation-usage}

%------------------------------------------------------------------------------
\subsection{Installation}

The ``installation'' of the {\easychair} document class is easy.
Download the latest version of the \texttt{easychair.zip} package 
from \url{http://www.easychair.org/easychair.zip}
and unzip it in the directory where you will prepare your paper.
You will get the following files, out of which you may need to keep only 
the \texttt{easychair.cls} style class if you are familiar with the rest 
of the files and do not require them to get started.
We are also working to make {\easychair} available from CTAN \cite{ctan},
such that it can be installed with the popular \TeX Live \cite{texlive} and
{\miktex} \cite{miktex} \LaTeX\ package management systems.

\begin{itemize}
\item
\texttt{easychair.cls} -- the class file that this is all about.

\item
\texttt{easychair-a4.pdf} -- the PDF version of this guide rendered using 
\texttt{a4paper} option.

\item
\texttt{easychair.tex} -- the {\LaTeX} source of this guide, 
and
\texttt{easychair.bib} -- the supporting bibliography entries found starting 
on page~\pageref{sect:bib}.

\item
\texttt{logoEC.pdf} -- the PDF version of the EasyChair logo rendered in
Figure~\ref{fig:easychair-logo} and
\texttt{throneEC.jpg} -- the JPG version of the easy throne rendered in 
Figure~\ref{fig:easythrone}.

\end{itemize}

%------------------------------------------------------------------------------
\subsection{Required Packages}

The {\easychair} class relies only on packages deemed standard and shipped by
most {\LaTeX} distributions in the worlds of Linux (current \texttt{texlive} \cite{texlive}
or older \texttt{tetex}), MacOS X,
and Windows (via Cygwin or {\miktex}).
If for some reason your distribution is old or doesn't have the packages
listed below, you can always obtain a copy from CTAN \cite{ctan}. Note
that EasyChair loads this packages automatically so you don't have to
use \verb|\usepackage| to use any of these packages.

\begin{itemize}
\item
\texttt{url} \cite{url-package} (included also by \texttt{hyperref} automatically) -- to provide URL rendering support for the
monospaced font, which takes care of special characters as well as line 
wrapping.

\item
\texttt{hyperref} \cite{hyperref-package} -- to allow hyperlinking of URLs and
cross references within an article.
Its options are set to either \verb+letterpaper+ or \verb+a4paper+, depending
on the \verb+\documentclass+ options.

\item
\texttt{graphicx} \cite{graphicx-package} -- the standard package for rendering
PNG, JPG, and PDF graphic images, primarily in \texttt{figure} environments.

\item
optional \texttt{mathptmx} \cite{mathptmx-package} -- Times base font for compactness
(use with the \texttt{withtimes} {\easychair} option).

\item
\texttt{listings} \cite{listings-package} -- to allow highlighted source code 
listing styles.

\item
\texttt{latexsym} \cite{latexsym-package} -- to provide common math and other 
symbols.

\item
\texttt{amsthm} \cite{amsthm-package} -- to provide {\AmS} theorem-like 
environments.

\item
\texttt{empheq} \cite{empheq-package} -- to provide equation environments, etc.

\item
\texttt{geometry} \cite{geometry-package} -- to set {\easychair} margins, 
outlined in Section~\ref{sect:generalities}.

\item
\texttt{lastpage} \cite{lastpage-package} -- to allow computationally
referencing the last page.

\item
\texttt{fancyhdr} \cite{fancyhdr-package} -- for running heads.

\item
\texttt{footmisc} \cite{footmisc-package} -- to ensure that footnotes are
always at the bottom.

\item
optional \texttt{makeidx} \cite{makeidx-package} -- for index generation
(use with the \texttt{thesis} {\easychair} option).

\item
\texttt{eso-pic} \cite{eso-pic-package} -- for draft versions and checking page
overlflows vs. a border drawn around the headers, footers, and the main body of
the article.

\end{itemize}

%------------------------------------------------------------------------------
\subsection{Recommended Packages}
\label{sect:recommended-packages}

Here is a list of some packages that this guide's authors have experimented 
with, and which are suitable for inclusion if needed by article authors.
These packages must be loaded using \verb+\usepackage+.
In general, authors may use any standard packages provided they do not change
the basic layout and font settings established by the {\easychair} class.
Such packages must be provided with the submission of articles.

\begin{itemize}
\item
\texttt{rotating} \cite{rotating-package} -- to rotate floats (figures and
tables) on the page, when wide tables or figures do not fit in portrait layout.

\item
\texttt{pdflscape} \cite{pdflscape-package} -- similar to \texttt{rotating}, 
but also allows rotating text to make it conveniently viewable in a PDF 
viewer that supports individual rotated pages.
A possible disadvantage is that a page break is forced, which may create
gaps before or after the landscape page.

\item
\texttt{algorithm2e} \cite{algorithm2e-package} -- provides a figure-like
algorithm environment for formal algorithm presentation with highlighting.

\end{itemize}

%------------------------------------------------------------------------------
\subsection{Compiling}
\label{sect:compiling}

\texttt{pdflatex} \cite{pdflatex-instructions} is the preferred tool for
producing PDF files with {\easychair} class documents.
The author kit (\texttt{easychair.zip}) includes some minimal automation 
that authors can use at their discretion.

\begin{itemize}
\item
Linux and UNIX-like platforms (also works under Cygwin and MacOS X):
A {\makefile} is provided for the GNU \texttt{make} \cite{gmake} utility,
so this document can be compiled by typing \texttt{make} at the terminal 
prompt (on the systems where both GNU and non-GNU versions of \texttt{make} 
are installed, one may need to use \texttt{gmake}).

\item
Microsoft Windows:
{\texniccenter} \cite{texniccenter} or {\latexeditor} \cite{led} and
{\miktex} \cite{miktex} as their backend are common tools
for {\LaTeX} processing under Microsoft Windows. 
The former provide a GUI front-end to {\LaTeX}, and the latter is the 
Windows native-compiled binaries and standard packages with 
a comprehensive package update tool. 
The \texttt{easychair.tcp} project file is provided for {\texniccenter} users,
as well as \texttt{easychair.lpr} for {\latexeditor} users.

\item
MacOS X:
TeXShop \cite{texshop} is a tool for {\LaTeX} processing under Mac OS X.
It provides a GUI front-end to {\LaTeX}. The backend can be installed
through the \texttt{fink} \cite{fink} repository or the Darwin Ports.
\end{itemize}

Some authors use packages that require the use of \texttt{latex}
instead of \texttt{pdflatex}, most notably \texttt{pstricks} and its
derivatives. If switching to modern packages, such as
\texttt{tikz} is not an option, these authors can use the {\easychair}
class with \texttt{latex}.

%------------------------------------------------------------------------------
\subsection{Bug Reports}
\label{sec:bug-reports}

Please report bugs, errors, and omissions you find with the {\easychair} 
class to its current maintainer, Andrei Voronkov,
at \url{andrei@voronkov.com}. Any \emph{constructive} feedback is
always welcome. If anybody has a very hard question, we might be
unable to answer it without Sergei. If anybody knows his whereabouts,
Geoff and Andrei have the ransom money ready.

%------------------------------------------------------------------------------

\section{Avoiding Text Overflows}

When used in EasyChair, the {\easychair} class is intended for
publishing, including printing. This implies that the text of your
document should not overflow the page dimensions. This section gives
some practical advice on how you can avoid text overflows. When you
run \texttt{pdflatex} or \texttt{latex} on your files, you sometimes
get warnings like

\small
\begin{verbatim}
  LaTeX Warning: Overfull \hbox (7.43138pt too wide) in paragraph at lines 994--994
\end{verbatim}
\normalsize
or

\small
\begin{verbatim}
  LaTeX Warning: Float too large for page by 55.21666pt on input line 638.
\end{verbatim}
\normalsize
They mean what they say: your text overflows the allowed space. When
you prepare a PDF document that will only be used online or as a
draft, in many cases you can ignore these warnings, since they only
make your document uglier than it could be but it remains still
readable. However, for documents that are supposed to be printed,
overflows make a lot of difference. There is no universal recipe on
how to make your text fit into the page. However, there are some
common cases that can be easily fixed using the tricks described
here.

\subsection{Using the \texttt{debug} Option to Check for Overflows}

A general guideline is that you should first check how the alleged
overflows looks in reality. For example, some LaTeX overflows are
invisible or hardly noticeable for people and overlows by
less than 4pt look OK anyhow. The {\easychair} class has an option
\texttt{debug} that adds to each page of your documents a red frame
corresponding to the allowed page
dimensions. Figure~\ref{fig:redframe} on page~\pageref{fig:redframe}
shows an example of a page created using this style. With such a red
frame, you can clearly see any overflows apart from very minor ones.

\begin{figure}
  \begin{centering}
    \includegraphics[width=0.89\textwidth]{debug.pdf}
    \caption{A page of a document created using the \texttt{debug}
      option} 
    \label{fig:redframe}
  \end{centering}
\end{figure}

\subsection{Tables}

Many page overflows happen because of large tables. In many case these
overflows can be easily removed by slightly reducing padding added by
\LaTeX\ to every column. It is controlled by the \LaTeX\ command
\verb|\tabcolsep| whose value by default is 6pt. Even small changes in
the value of this command may give drastic reductions in the width of
tables. This is illustrated in Figure~\ref{fig:tabcolsep} on
page~\pageref{fig:tabcolsep}. Note though that there is no free lunch:
smaller values for this command may result in lower redability.

%------------------------------------------------------------
\begin{figure}[tb]\small
\begin{center}
  \begin{tabular}{lrrrrrrrr}
    \hline
    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
   \hline
    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
   \hline
  \end{tabular}
\end{center}

\begin{center}\renewcommand{\tabcolsep}{5pt}
  \begin{tabular}{lrrrrrrrr}
    \hline
    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
   \hline
    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
   \hline
  \end{tabular}
\end{center}

\begin{center}\renewcommand{\tabcolsep}{3pt}
  \begin{tabular}{lrrrrrrrr}
    \hline
    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
   \hline
    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
   \hline
  \end{tabular}
\end{center}

\begin{center}\renewcommand{\tabcolsep}{1pt}
  \begin{tabular}{lrrrrrrrr}
    \hline
    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
   \hline
    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
   \hline
  \end{tabular}
\end{center}
\normalsize

\caption{Original table and tables with \texttt{tabcolsep} set to 5pt,
  3pt, and 1pt
  \label{fig:tabcolsep}}

\end{figure}
%------------------------------------------------------------

\subsection{Images}

Images included using \verb|\includegraphics| are easy to resize since
one can specify the size of the result explicitly. For example,
Figure~\ref{fig:easythrone} shows three copies of the same image
having different sizes obtained using the following commands:

\begin{verbatim}
  \includegraphics[width=0.5\textwidth]{throneEC.jpg}
  \includegraphics[width=0.3\textwidth]{throneEC.jpg}
  \includegraphics[width=0.15\textwidth]{throneEC.jpg}
\end{verbatim}

\subsection{A Universal Recipe}

\LaTeX\ has a very powerful weapon for reducing the size of almost
anythings. More precisely, it can reduce anything producing what
\LaTeX\ considers a box. This weapon is called
\verb|\scalebox|. Consider an example (check the source of this file
to see how it was produced).

\begin{center}
\begin{tabular}{|c|c|}
\hline \Huge
\begin{tabular}[b]{cr}
year & users \\ \hline
2007 &    47,753 \\
 2008 &    114,494 \\
 2009 &    207,506 \\
 2010 &   371,054 \\
\end{tabular}
&
\includegraphics[width=0.28\textwidth]{chairEC}\\
\hline
\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
  logos,}\\
\multicolumn{2}{|c|}{scaled to the number of users in 2010} \\
\hline
\end{tabular}
\end{center}
This is what happens when we put (almost) the same \LaTeX\ code in 
\verb|\scalebox{0.55923}{...}| to scale it down to the number of users
in 2009:

\begin{center}
\scalebox{0.55923}{%
\begin{tabular}{|c|c|}
\hline \Huge
\begin{tabular}[b]{cr}
year & users \\ \hline
2007 &    47,753 \\
 2008 &    114,494 \\
 2009 &    207,506 \\
 2010 &   371,054 \\
\end{tabular}
&
\includegraphics[width=0.28\textwidth]{chairEC}\\
\hline
\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
  logos,}\\
\multicolumn{2}{|c|}{scaled to the number of users in 2009} \\
\hline
\end{tabular}}
\end{center}
We can scale it down even further to the 2008 figure using
\verb|\scalebox{0.30856}{...}|:

\begin{center}
\scalebox{0.30856}{%
\begin{tabular}{|c|c|}
\hline \Huge
\begin{tabular}[b]{cr}
year & users \\ \hline
2007 &    47,753 \\
 2008 &    114,494 \\
 2009 &    207,506 \\
 2010 &   371,054 \\
\end{tabular}
&
\includegraphics[width=0.28\textwidth]{chairEC}\\
\hline
\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
  logos,}\\
\multicolumn{2}{|c|}{scaled to the number of users in 2008} \\
\hline
\end{tabular}}
\end{center}
or further down to 2007:

\begin{center}
\scalebox{0.12870}{%
\begin{tabular}{|c|c|}
\hline \Huge
\begin{tabular}[b]{cr}
year & users \\ \hline
2007 &    47,753 \\
 2008 &    114,494 \\
 2009 &    207,506 \\
 2010 &   371,054 \\
\end{tabular}
&
\includegraphics[width=0.28\textwidth]{chairEC}\\
\hline
\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
  logos,}\\
\multicolumn{2}{|c|}{scaled to the number of users in 2008} \\
\hline
\end{tabular}}
\end{center}

This size reduction technique is very efficient: using the right scale
you may post your whole article on Twitter in a single tweet. However,
it may also may parts of your text virtually unreadable with an
unfortunate side effect of annoying reviewers.

\section{Submitting Your Article Through EasyChair}

This section is intended only for the authors and editors of EasyChair
proceedings and Procedia volumes.

When you prepare an article for either of these, it should be
submitted through EasyChair. EasyChair automates the submission
process as much as possible and goes to a great length to ensure that
your article can be published and printed. Publication for EasyChair
means much more than just putting a PDF of your article online. It
collects a some meta-information about the article to classify it,
find similar articles, make it easily searchable, and index it in
various Web services, such as DBLP. This section explains how
EasyChair processes your article.

\subsection{Submitting the Article}

Submitting the article is easy. All you should do is to put the source
of your article in a single zip file. The source must contain all
auxiliary files required to create a PDF file of your article: this
includes images, bibliography, and all non-standard \LaTeX\ packages
you used\footnote{ 
  A non-standard \LaTeX\ package is a package that is not included in
  CTAN. 
} For example, suppose that your main \LaTeX\ file is
\texttt{main.tex}, it inputs another file \texttt{macros.tex} and uses
the file \texttt{biblio.bib} to produce the bibliography. Suppose it
also uses two images \texttt{images/easy.jpg} and
\texttt{images/easy.jpg}. Then you should create a zip archive
containing all these files. Suppose all these files are put in a
directory \texttt{mypaper} on your computer, where \texttt{images} is
a subdirectory of \texttt{mypaper}

On almost any operating system (Linux,
Windows, or Mac) you can achieve this by using the following sequence
of commands:

\begin{verbatim}
cd mypaper
zip -r mypaper.zip *
\end{verbatim}
This will create a zip archive \texttt{mypaper.zip} including all
files in the directory \texttt{mypaper} and its subdirectories.

\subsection{EPiC Series and Kalpa Publications}
\label{subsec:EPiC}

If you publish your article in the EasyChair EPiC series, our aim is to produce
a nicely looking EPiC Series header, see Figure~\ref{fig:epicheader} on page
\pageref{fig:epicheader} for an idea. The images and information in the header
will be added by EasyChair automatically and depend on the subseries and the
volume information.

\begin{figure}
  \begin{centering}
    \includegraphics[width=0.89\textwidth]{epic.pdf}
    \caption{This document with an EPiC Series header} 
    \label{fig:epicheader}
  \end{centering}
\end{figure}

To ensure that the layout of your document will not change when the header is
added, you should use the option \texttt{EPiC} as follows:

\small
\begin{verbatim}
\documentclass[EPiC]{easychair}
\end{verbatim}
\normalsize
This will reserve space for the header on the top of the first page of your
article.

%------------------------------------------------------------------------------
\section{Future Work}
\label{sect:future-work}

We plan to further strengthen the {\easychair} class and promote it for 
electronic publishing for EasyChair-powered conferences and workshops,
and take over the world, as shown in Figure~\ref{fig:easythrone}. We
aim at creating a new model of \emph{affordable publishing}, where
anybody can become a low-cost publisher. The
\href{http://www.easychair.org/publications/EPiC}{EasyChair EPiC
  Series} and
the
\href{http://www.easychair.org/publications/Kalpa}{EasyChair Kalpa Publications} 
are our first contributions to affordable open access publishing.

\begin{figure}[tb]
  \begin{centering}
    \includegraphics[width=0.5\textwidth]{throneEC.jpg}
    \includegraphics[width=0.3\textwidth]{throneEC.jpg}
    \includegraphics[width=0.15\textwidth]{throneEC.jpg}
  \end{centering}
  \caption{Why one should use EasyChair}
  \label{fig:easythrone}
\end{figure}

%------------------------------------------------------------------------------
\section{Acknowledgments}
\label{sect:acks}

\begin{itemize}
\item
Aleksander Kosenkov for the graphics that are used here and on the EasyChair 
website~\cite{easychair}.

\item
Graham Gough (The University of Manchester) for implementing title-related
commands.

\item
Thomas F. Sturm (Universit\"at der Bundeswehr M\"unchen) for implementing the
\LaTeX\ part of the headers used for the EasyChair EPiC series.

\item
The CTAN \cite{ctan} and {\LaTeX} communities \cite{texniccenter,miktex}.

\item
Guilin Qi, Jasmin Christian Blanchette, Leslie Lamport, Uwe Pfeiffer,
and others for constructive feedback on the style, most of which had been
incorporated in version 2 of the class style.
\end{itemize}

\label{sect:bib}
\bibliographystyle{plain}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{easychair}

%------------------------------------------------------------------------------
\appendix
\section{{\easychair} Requirements Specification}
\label{sect:easychair-requirements}

The following high-level requirements were set for the development of 
the {\easychair} class, and were refined as development went along.

\begin{enumerate}
\item
The style should be easy to use. 
The average {\LaTeX} user should not need to read a long manual.

\item
It should be economical in space but the text should be nice-to-read.

\item
It should use fonts producing a reasonable-quality PDF.

\item
The bibliography should produce hyperlinks.

\item
Sections should produce menu sections in PDF.

\item
The text should look good on both A4 and letter paper.

\item
The style should be single-column for convenience of scrolling.

\item
The print area should be convenient for printing using print-on-demand publishers.

\item
Running heads.
\end{enumerate}

%------------------------------------------------------------------------------
% Index
%\printindex

%------------------------------------------------------------------------------
\end{document}

### 2020-04-23 coqws/elpi.py

@@ -1,0 +1,1 @@
+../elpi.py

### 2020-04-23 elpi.py

@@ -117,7 +117,15 @@ class CoqElpiLexer(RegexLexer):
         'Qed', 'Defined', 'Hint', 'Resolve', 'Rewrite', 'View', 'Search',
         'Show', 'Print', 'Printing', 'All', 'Graph', 'Projections', 'inside',
         'outside', 'Check', 'Global', 'Instance', 'Existing',
-        'Universe', 'Polymorphic', 'Monomorphic', 'Context', 'Elpi', 'Accumulate', 'Db', 'Command'
+        'Universe', 'Polymorphic', 'Monomorphic', 'Context', 'Elpi',
+        'Accumulate', 'Db', 'Command', 'Infix', )
+    hb = (
+        'HB.mixin',
+        'HB.factory',
+        'HB.instance',
+        'HB.structure',
+        'HB.builders',
+        'HB.end',
     )
     keywords2 = (
         # Gallina
@@ -185,6 +193,7 @@ class CoqElpiLexer(RegexLexer):
             (words(keywords4, prefix=r'\b', suffix=r'\b'), Keyword),
             (words(keywords5, prefix=r'\b', suffix=r'\b'), Keyword.Pseudo),
             (words(keywords6, prefix=r'\b', suffix=r'\b'), Keyword.Reserved),
+            (words(hb, prefix=r'\b', suffix=r'\b'), Name.Class),
             # (r'\b([A-Z][\w\']*)(\.)', Name.Namespace, 'dotted'),
             (r'&', Keyword),
             (r'\b([A-Z][\w\']*)', Name),

### 2020-04-24 hb.tex
Address the =KS= items except explanation of the flat variant

@@
 
-The language of packed classes were introduced in
-\cite{DBLP:conf/tphol/GarillotGMR09} and used directly to describe the
-algebraic hierarchy of the Mathematical Components library.
-It is based on a disciplined use of records and projections and
-on the possibility of extending the elaborator of Coq
-via the declaration of Canonical Structures instances.
-In this section we describe the \emph{flat} variant of
-Packed Classes, the target language of \HB{}.
+The language of Packed Classes~\cite{DBLP:conf/tphol/GarillotGMR09} is used directly to describe the algebraic hierarchy of the Mathematical Components library.
+It is based on a disciplined use of records and projections and on the possibility of extending the elaborator of Coq via the declaration of Canonical Structures~\cite{DBLP:conf/itp/MahboubiT13} instances.
+In this section, we describe the \emph{flat} variant of Packed Classes, the target language of \hb{}, through the hierarchy V3 extended with semirings.
+
+The \hb{} addon can generate all the Coq declarations in this section automatically, but some details are omitted for brevity in this section.
 
@@
 Since the monoid structure is the bottom of this hierarchy its class record
-\coq{Monoid} consists just one \mixin{}. In contrast the class record
+\coq{Monoid} consists of just one \mixin{}. In contrast the class record
 of Abelian groups consists of two \mixins{} where the second one depends
@@
 In order to introduce multiple inheritance, we extend the hierarchy described in
-section~\ref{subsec:v3} with the structure of semirings%, which is another
+section~\ref{subsec:v3} with the structure of semirings %, which is another
 %intermediate structure between monoids and rings 
@@
 This software component
-takes in input a term as written by the user and has to infer all the missing information that is necessary in order to make the term well typed.
+takes as input a term as written by the user and has to infer all the missing information that is necessary in order to make the term well typed.
 

### 2020-04-24 hb.tex
splitting, refactoring and clarifications in Sect 3.1

--- 2020-04-17 hb.tex
@@
 
-\definecolor{dkgreen}{rgb}{0, 0.502, 0}
+\definecolor{dkgreen}{rgb}{0, 0, 1}
 
@@
 
-\ccsdesc[100]{General and reference~General literature}
-\ccsdesc[100]{General and reference}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
+\ccsdesc{Software and its engineering~Formal language definitions}
+\ccsdesc{Theory of computation~Type theory}
+\ccsdesc{Theory of computation~Constraint and logic programming}
+\ccsdesc{Computing methodologies~Symbolic and algebraic manipulation}
+
 \supplement{Source code of the Coq package: \url{https://github.com/math-comp/hierarchy-builder}}
@@
 
-\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+\relatedversion{System Description}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
 %\relatedversion{A full version of the paper is available at \url{...}.}
-
-\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
 
@@
 %Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{John Q. Open and Joan R. Access}
-\EventNoEds{2}
-\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
-\EventShortTitle{CVIT 2016}
-\EventAcronym{CVIT}
-\EventYear{2016}
-\EventDate{December 24--27, 2016}
-\EventLocation{Little Whinging, United Kingdom}
+\EventEditors{Zena M. Ariola}
+\EventNoEds{1}
+\EventLongTitle{5th International Conference on Formal Structures for Computation and Deduction (FSCD 2020)}
+\EventShortTitle{FSCD 2020}
+\EventAcronym{FSCD}
+\EventYear{2020}
+\EventDate{June 29--July 5, 2020}
+\EventLocation{Paris, France}
 \EventLogo{}
-\SeriesVolume{42}
-\ArticleNo{23}
+\SeriesVolume{167}
+\ArticleNo{8}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\newtheoremstyle{commands}{\topsep}{\topsep}{}{0pt}{\sffamily\bfseries}{ }{5pt plus 1pt minus 1pt}%
+{$\blacktriangleright$ \thmname{#1} {\coq{#3}}}
+\theoremstyle{commands}
+\newtheorem*{commands}{Coq commands}
+
 \begin{document}
@@
 
+In this section, we call ``distinguished'' a Coq definition or record
+that the developer of a library has labeled ``\mixin{}'', ``\factory{}'',
+``\mixinbuilder{}'', ``\class{}'' or ``\structure{}''.
+
 \begin{definition}[\M{}, \mixins{}]
-A \emph{\mixin{}} \(m \in \M{}\) is a Coq record with one or more parameters.
-The first parameter must be a \coq{(T : Type)}, while the other parameters
-are \mixins{} applied to \coq{T} and possibly $n$ other distinct \mixins{}. I.e.
+  A \emph{\mixin{}} \(m \in \M{}\) is a distinguished Coq record with one or more
+  parameters.  The first parameter must be a \coq{(T : Type)}, while
+  the other parameters are \mixins{} \(\left(m_i\right)_{i \in \enum1n}\), each of which
+  is applied to \coq{T} and possibly previous mixin variables. I.e.
 \begin{coqcode}
@@
 \end{coqcode}
-where \(\pmp{n} = \vect{\left(p_i : m_i\ T\ p_{\sigma(i,1)}\ \ldots\ p_{\sigma(i,n_i)}\right)}_{i \in \enum1n}\).
-For all \(i \in \enum1n\) we have \(m_i \in \M{}\) and its
-arguments consist in $n_i$ of the previously quantified
-mixin parameters $p_k$, i.e. for all \(k \in \enum1{n_i}\), we
+where \(\pmp{n} = (p_1 : m_1\ T) \ldots \left(p_n : m_n\ T\ p_{\sigma(n,1)}\ \ldots\ p_{\sigma(n,q_n)}\right)\)
+and where for all \(i~\in~\enum1n\) we have \(q_i \in \mathbb{N}\) and the
+arguments of \(m_i \in \M{}\) consist in $q_i$ of the previously quantified
+mixin parameters $p_k$, i.e. for all \(k \in \enum1{q_i}\), we
 have \(\sigma(i,k) \in \enum1{i-1}\).
@@
 \begin{remark}For all \(i \in \enum1n\), we have
-\(\dep{}(m_i) = \enum{m_{\sigma(i,1)}}{m_{\sigma(i,m_i)}}\).
+  \(\dep{}(m_i) = \enum{m_{\sigma(i,1)}}{m_{\sigma(i,q_i)}}\).
 \end{remark}
@@
 
-\begin{proposition}
-\dep{} is transitively closed since records are well typed in the
-empty context and describes a DAG since Coq does not admit circular
-definitions: \(\dep{}^\star(\dep{}^\star(M)) \subseteq \dep{}^\star(M)\)
+\begin{proposition}[\dep{} is transitively closed and describes a DAG]
+  \[\forall m \in \M{}, \qquad \dep{}^\star(\dep{}^\star(m)) \subseteq
+  \dep{}^\star(m) \quad \textrm{and} \quad m \notin \dep{}(m).\]
 \end{proposition}
-
-\begin{definition}[\factories{} \F{}, \mixinbuilders{} \B{} and \from{}]
-  A \emph{\mixinbuilder{}} \(\mu \in \B{}\) is any function which parameters are the carrier
-  type \coq{(T: Type)}, an arbitrary number of distinct \mixins{}
-  \(\enum{m_1}{m_n}\) and an additional argument \(f\), which we call
-  a \emph{\factory{}}, and which depends on the \mixins{}. Its return type is
-  a \mixin{} \(m_{n+1} \in \M{}\) i.e.
-\begin{coqcode}
-Definition f (T : Type) #$\pmp{n}$# : Type := ...
-Definition #$\mu$# (T : Type) #$\pmp{n}$#: #\(f\ T\ p_1 \ldots p_n\)# -> #\(m_{n+1}\ T\ p_{\sigma(n+1,1)}\ \ldots\ p_{\sigma(n+1,n_i)}\)# := ...
-\end{coqcode}
-
+\begin{proof}
+  Indeed \dep{} is transitively closed because records are well typed
+  in the empty context and describes a DAG since Coq does not admit
+  circular definitions.
+\end{proof}
+
+\begin{definition}[\factories{} \F{}]
+A \emph{\factory{}} \(f \in \F{}\) is a distinguished Coq record or definition with one or more parameters.
+The first parameter must be a \coq{(T : Type)}, while the other parameters
+are $n$ \mixins{}, applied to \coq{T} and previous mixin variables. I.e.
+\begin{coqcode}
+Record (* or Definition *) #$f$# (T : Type) #\(\pmp{n}\)# : Type := ...
+\end{coqcode}
+\end{definition}
+
+\begin{definition}[$\requires{}$, factory requirements.]
+  Given \(f \in \F{}\), we define \(\requires{}(f) \in \powerset\M{}\)
+  as the set of all \mixins{} that occur as parameters of \(f\), i.e.
+  \(\dep{}(f) = \enum{m_1}{m_n}\).
+\end{definition}
+
+The following property holds,
+\begin{proposition}[\requires{} is closed under \dep{}]
+\(\forall f \in \F{}, \dep{}^\star(\requires{}(f)) \subseteq \requires{}(f)\)
+\end{proposition}
+\begin{proof}
+Because Coq records and global definitions are well typed in the empty context.
+\end{proof}
+
+\begin{definition}[\mixinbuilders{} \B{}]
+  A \emph{\mixinbuilder{}} \(\mu \in \B{}\) is a distinguished function whose
+  return type is a mixin \(m_{n+1} \in \M{}\) and whose parameters are
+  the carrier type \coq{(T : Type)}, \mixins{}
+  \(\enum{m_1}{m_n} \in \powerset\M{}\) and a \emph{\factory{}}
+  \(f \in \F{}\), such that \( \requires{}(f) =
+  \enum{m_1}{m_n}\). I.e.
+\begin{coqcode}
+Definition #$\mu$# (T : Type) #$\pmp{n}$#: #\(f\ T\ p_1 \ldots p_n\)# -> #\(m_{n+1}\ T\ p_{\sigma(n+1,1)}\ \ldots\ p_{\sigma(n+1,q_{n+1})}\)# := ...
+\end{coqcode}
+\end{definition}
+
+Note that the \mixinbuilders{} of a given \factory{} have the same set
+of dependencies.
+
+\begin{definition}[\from]
   We define \(\from{} (f, m_{n+1}) = \mu\) to be the (unique) \mixinbuilder{}
-  for \(m_{n+1}\) from the \factory{} \(f\).
+  for \(m_{n+1}\) from the \factory{} \(f\), when it exists.
 \end{definition}
 
-\begin{definition}[$\provides{}$, $\requires{}$]
-  We define the following functions \begin{itemize}
-  \item\( \requires{}(f) = \enum{m_1}{m_n}  \in \powerset\M{}\)
-    is the set of mixins a \factory{} \(f \in \F{}\) depends on.
-    As a consequence all the
-    \mixinbuilders{} of a given \factory{} have the same set of
-    dependencies.
-  \item\( \provides{}(f) = \{m | \from{}(f,m) \textrm{ is defined}\}
+Note that \(\from\) is not a total function, and that \(\from{}(f, m)\) is
+defined if and only if there is a declared builder that shows how to
+build $m$ from $f$. In this case, we say $f$ provides $m$.
+
+\begin{definition}[$\provides{}$]
+  \( \provides{}(f) = \{m | \from{}(f,m) \textrm{ is defined}\}
     \in \powerset\M{}\) the set of mixins that a factory
     \(f \in \F{}\) provides, through its \mixinbuilders{}.
-  \end{itemize}
-Moreover the following properties hold for all \(f \in \F{}\):
-  %\begin{proposition}
-  %  They validate the following properties for all \(f \in \F{}\):
-    \begin{enumerate}
-      \item\( \dep{}^\star(\requires{}(f)) \subseteq \requires{}(f)\)
-      \item\( \requires{}(f) \cap \provides{}(f) = \emptyset \)
-      \item\( \forall m \in \provides{}(f),~ \from{}(f,m) \) is defined
-    \end{enumerate}
-  %\end{proposition}
 \end{definition}
@@
 
-\begin{proposition}[\( \M{} \subseteq \F{} \)] For all \(m \in \M{}\)
-  we have \(\requires{}(m) = \dep{}(m)\), \(\provides{}(m) = \{m\}\)
-  and
+\begin{proposition}[\( \M{} \subseteq \F{} \)] There is a way to see mixins as factories.
+\end{proposition}
+\begin{proof}
+  For all \(m \in \M{}\) we have \(\requires{}(m) = \dep{}(m)\),
+  \(\provides{}(m) = \{m\}\) and
   \(\from{}(m,m) = \left(\coq{fun T }\ \pmp{n} \ \coq{(x :}\ m\
     \coq{T }\ p_1\ \ldots p_n\ \coq{) => x}\right) \in \B{}.\)
-\end{proposition}
+\end{proof}
 
@@
 \end{coqcode}
-  Creates a module \coq{M} with a record
-  \coq{M}, which depends on the \mixins{}
-  \(\requires{}(\coq{M}) =
+  Creates a module \coq{M} with a record which is exported as
+  \coq{M}, and which depends on the \mixins{}
+  \(\requires{}(\coq{M}) = \dep{}(\coq{M}) =
   \provides{}^\star\left(\enum{f_1}{f_n}\right)\), and registers this
-  new record both as a \mixin{} and a \factory{}. It also creates a
+  new record both as a \mixin{} and a \factory{}. It also exports an
   \phantterm{} \coq{M.Build.}
@@
   \mixins{} that can be built from the $b_i$. Indeed if
-  \(b_i : f_i\ T\ \ldots\), then this command creates elements of type
+  \(b_i : f_i\ T\ \ldots\), then this command creates elements of types
   \(\provides{}^\star\left(\enum{f_1}{f_k}\right).\) This command also
@@
 
-\begin{command} to declare a new \factory{} and generate new \mixinbuilders:
-\begin{coqcode}
-#\hbfactory{}# Record M T of #$f_1\ \ldots\ f_n$# := { .. }.
-#\hbbuilders{}# Context T (a : M A).
+\begin{commands} to declare a new \factory{} and generate new \mixinbuilders:
+\begin{coqcode}
+#\hbfactory{}# Record F T of #$f_1\ \ldots\ f_n$# := { .. }.
+#\hbbuilders{}# Context T (a : F T).
 
@@
 \end{coqcode}
-  Creates a module \coq{M} with a record
-  \coq{M}, which depends on the mixins
+  Creates a module \coq{F} with a record which is exported as
+  \coq{F}, and which depends on the mixins
   \(\requires{}(\coq{M}) =
   \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and registers this
-  new record as a factory. We then use the \factoryinstances{}
+  new record as a factory.  It also exports an
+  \phantterm{} \coq{F.Build.} We then use the \factoryinstances{}
   \(b_{n+1}, \ldots, b_{n+k}\) provided by the user and derive
   \mixinbuilders{} from them, so that
-  \(\provides{}(\coq{M}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)
+  \(\provides{}(\coq{F}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)
 
@@
   \provides{}^\star\left(\enum{f_1}{f_n}\right).\)
-\end{command}
+\end{commands}
 
 Note that the \(b_i\) are not \mixinbuilders{} since their return types
-are not necessarily \mixins{}, but could be a \factory{}. However, since all
-\factories{} provide \mixin{} through \mixinbuilders{}, we can
+are not necessarily \mixins{}, but could be \factories{}. However, since all
+\factories{} provide \mixins{} through \mixinbuilders{}, we
 obtain \mixinbuilders{} out of each \(b_i\) by function composition.
@@
 \begin{definition}[\C{}, class]
-A \class{} \coq{c} \(\in \C{}\) is a Coq record with one parameter
+A \class{} \coq{c} \(\in \C{}\) is distinguished a Coq record with one parameter
 \coq{(T : Type)}. The type of each field is a \mixin{} in \M{} applied to
@@
 \begin{coqcode}
-Record #$c$# (T : Type) := { #\(p_1 : m_1\ T\)#; .. ; #\(p_i : m_i\ T\ p_{\sigma(i,1)}\ \ldots p_{\sigma(i,n_i)}\)#; ..}.
+Record #$c$# (T : Type) := { #\(p_1 : m_1\ T\)#; .. ; #\(p_i : m_i\ T\ p_{\sigma(i,1)}\ \ldots p_{\sigma(i,q_i)}\)#; ..}.
 \end{coqcode}
@@
   \begin{enumerate}
-    \item\( \exists c \in \C{},~ \cdef{}(c) = \requires{}^\star(f) \cup \provides{}^\star(f),\)
+    \item\( \exists c \in \C{},~ \cdef{}(c) = \requires{}^\star(f)
+      \cup \provides{}^\star(f),\)
     \item\( \exists C \subseteq \C{},~\cdef{}^\star(C) = \requires{}^\star(f).\)
@@
 \begin{definition}[\Str{} Structure]
-A \structure{} \coq{s} \(\in \Str{}\) is a dependent pair: a Coq record where
+A \structure{} \coq{s} \(\in \Str{}\) is a distinguished dependent pair: a Coq record where
 the value of the first field occurs in the type of the second.
@@
 \begin{command} to declare a class and structure:
-  \hbstructure{} \coq#Definition M := { A of# \(f_1 \ldots f_n\) \coq#}.#
+  \hbstructure{} \coq#Definition M := { A of# \(f_1 \ldots f_n\) \coq#}#
   crafts a class \coq{M.class_of} $\in \C{}$ where

### 2020-04-24 coqws/easychair.tex
coq ws draft

@@
 %\documentclass[letterpaper]{easychair}
+\usepackage{wrapfig}
 
@@
 \newmintinline[elpi]{elpi.py:ElpiLexer -x}{fontsize=\small}
-\newminted[coqcode]{elpi.py:CoqElpiLexer -x}{linenos=true,fontsize=\footnotesize,mathescape,escapeinside=\#\#}
+\newminted[coqcode]{elpi.py:CoqElpiLexer -x}{firstnumber=last,linenos=true,fontsize=\footnotesize,mathescape,escapeinside=\#\#}
 \newmintinline[coq]{elpi.py:CoqElpiLexer -x}{fontsize=\small}
@@
 \titlerunning{Hierarchy Builder}
+\newcommand{\HB}{\ensuremath{\mathcal{HB}}}
 
@@
 
-\begin{abstract}
-  We present a tool that makes it easy to declare and evolve hierarchies
-  of algebraic structures in Coq.
-\end{abstract}
+%\begin{abstract}
+%  We present a tool that makes it easy to build hierarchies
+%  of algebraic structures in Coq.
+%\end{abstract}
 
@@
 
-bla bla librarires.
-bla bla structures.
-bla bla packed classes.
-bla bla difficult.
-bla bla HB language.
-bla bla imlemented in Elpi.
-In the talk we give a demo of the tool, details are in the paper cite FSCD.
+There is no questioning the importance of abstract algebra in the mathematical
+vernacular. And it is with no surprise that libraries of machine checked code
+are, nowadays, organized around hierarchies of structures. Unfortunately the
+language of Type Theory and the features provided by the Coq system make the
+construction of a hierarchy hard even for expert users. The difficulty
+begins with the non orthogonal choices of the encoding style, from bundled to
+unbundled, and inference mechanism, from Type Classes to Canonical Structures.
+To this one may add the concerns about performance and finally the ones about
+the usability of the final hierarchy.
 
-\newpage
-\section{An example of HB}
+In this talk we give an extensive demo of the Hierarchy Builder (\HB{})
+extension for the Coq system version 8.10 and above.
+\HB{} gives the library designer a language to describe the building blocks
+of algebraic structures and to assemble them into a hierarchy. Similarly it
+provides the final user linguistic constructs to build instances (examples) of
+structures and to teach the elaborator of Coq how to take advantage of this
+knowledge during type inference. Finally \HB{} lets the library designer
+improve the usability of his library by providing alternative interfaces
+to the primitive ones, a feature that can also be used to accommodate changes
+without breaking user code.
+
+The details of \HB{}, including its compilation to the (Flat) Packed Classes
+discipline and its implementation in the Elpi extension language for Coq,
+are given in~\cite{cohen:hal-02478907}.
+In this talk we focus on how to take advantage
+of \HB{} in a Coq development, and we plan to not detail how it works
+behind the scenes, unless specific questions arise from the audience.
+
+\section{A hierarchy with a diamond in 50 lines}
 
 \begin{coqcode}
-HB.mixin Record AddComoid_of_Type A := {
+HB.mixin Record CMonoid_of_Type A := { (* The set of axioms making A a commutative monoid *)
   zero : A;
   add : A -> A -> A;
-  addrA : forall x y z, add x (add y z) = add (add x y) z;
-  addrC : forall x y, add x y = add y x;
-  add0r : forall x, add zero x = x;
+  addrA : associative add;
+  addrC : commutative add;
+  add0r : left_id zero add;
 }.
 
-HB.structure Definition AddComoid := { A of AddComoid_of_Type A }.
+HB.structure Definition CMonoid := { A of CMonoid_of_Type A }. (* The structure thereof *)
 
@@
 
-Check forall (M : AddComoid.type) (x : M), x + x = 0.
+Check 0.     (* ?M : CMonoid.type |- 0 : ?M *)
+Check addrC. (* ?M : CMonoid.type |- forall x y : ?M, x + y = y + x *)
 \end{coqcode}
 
-where we are, where we go
+\newpage
 
+\begin{wrapfigure}[0]{r}{.4\textwidth}
+  \includegraphics[width=.4\textwidth]{diagram.pdf}
+  %\caption{The hierarchy}
+  %\label{fig:hierarchy}
+\end{wrapfigure}
+\
 \begin{coqcode}
-HB.mixin Record AbelianGrp_of_AddComoid A of AddComoid A := {
+HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
   opp : A -> A;
-  addNr : forall x, opp x + x = 0;
+  addNr : left_inverse zero opp add; (* add is available here *)
 }.
 
-HB.structure Definition AbelianGrp := { A of AbelianGrp_of_AddComoid A & }.
+HB.structure Definition AbelianGrp :=
+  { A of AbelianGrp_of_CMonoid A & }.
 
@@
 
-Lemma example (G : AbelianGrp.type) (x : G) : x + (- x) = - 0.
-Proof. by rewrite addrC addNr -[LHS](addNr zero) addrC add0r. Qed.
+Check forall G : AbelianGrp.type, forall x : G, x + - x = 0.
+
+HB.mixin Record SemiRing_of_CMonoid A of CMonoid A := {
+  one : A;
+  mul : A -> A -> A;
+  mulrA : associative mul;
+  mul1r : left_id one mul;
+  mulr1 : right_id one mul;
+  mulrDl : left_distributive mul add;
+  mulrDr : right_distributive mul add;
+  mul0r : left_zero zero mul;
+  mulr0 : right_zero zero mul;
+}.
+
+HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & }.
+
+Notation "1" := one.
+Infix "*" := mul.
+
+Check forall S : SemiRing.type, forall x : S, x * x = 0 + 1. (* can I use - here? *)
+
+HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A & }.
+
+Check forall R : Ring.type, forall x : R, x * - (1 + x) = 0 + 1.
+
+Definition Z_CMonoid := CMonoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
+HB.instance Z Z_CMonoid.
+Definition Z_AbelianGrp := AbelianGrp_of_CMonoid.Build Z Z.opp Z.add_opp_diag_l.
+HB.instance Z Z_AbelianGrp.
+Definition Z_SemiRing := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l ...
+HB.instance Z Z_SemiRing.
+
+Check forall x : Z, x * - (1 + x) = 0 + 1.
 \end{coqcode}
 
-and now examples
-
-\begin{coqcode}
-
-Definition Z_CoMoid := AddComoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
-HB.instance Z Z_CoMoid.
-Definition Z_AbGrp := AbelianGrp_of_AddComoid.Build Z Z.opp Z.add_opp_diag_l.
-HB.instance Z Z_AbGrp.
-
-Lemma example2 (x : Z) : x + (- x) = - 0.
-Proof. by rewrite example. Qed.
-\end{coqcode}
+For space constraints we omit the 30 lines in which we use
+the \coq{HB.factory} command to declare an alternative interface
+to build examples of the \coq{Ring} structure (the green
+arrow in the picture) by providing all proofs at once but for the
+one on the commutativity of the Monoid operation:
+the other Ring axioms imply it~\cite{nearrings}. Factories can also be used to
+preserve backward compatibility when an hierarchy evolves to accommodate
+new intermediate  structures.
 
@@
 \end{document}
-
-%------------------------------------------------------------------------------
-\section{Introduction}
-\label{sect:introduction}
-
-The {\easychair} class was designed to be easy to use, and specifically favors
-electronic and on-paper publishing by the EasyChair conference system
-\cite{easychair}, including the
-\href{http://www.easychair.org/publications/EPiC}{EasyChair EPiC
-  Series} and \href{http://www.easychair.org/publications/Kalpa}{EasyChair Kalpa
-  Publications}. 
-EasyChair is a free conference management system that is flexible, easy to use,
-and has many features to make it suitable for various conference models. It is
-currently probably the most commonly used conference management system
-\cite{easychair}.
-
-The {\easychair} class was designed according to some requirements, which
-are described in Appendix~\ref{sect:easychair-requirements}. 
-An article that occupies approximately 15 LNCS-formatted pages
-takes up approximately 14 {\easychair} pages.
-
-\begin{figure}[tb]
-	\begin{centering}
-	\includegraphics[width=0.5\textwidth]{logoEC}
-	\caption{EasyChair logo}
-	\label{fig:easychair-logo}
-	\end{centering}
-\end{figure}
-
-%------------------------------------------------------------------------------
-\section{Typesetting}
-\label{sect:typesetting}
-
-Typesetting with {\easychair} is, well, easy.  Just by using the
-document class entry in the document's preamble as follows:
-\verb+\documentclass{easychair}+ the typesetting work is nearly done.
-The {\easychair} class is a relatively conservative extension of the
-standard \texttt{article} class, so most of the environments, section
-headers, etc. defined by \texttt{article} are available.
-
-%------------------------------------------------------------------------------
-\subsection{Generalities}
-\label{sect:generalities}
-
-The following are the general default parameters {\easychair}
-introduces into the typesetting aspect of articles. If you use
-\easychair\ for proceedings or other kinds of publishing through
-EasyChair, do not alter these -- papers deviating from the formatting
-standards will be rejected by EasyChair.
-
-\begin{enumerate}
-\item
-The default paper size is US letter. It can be explicitly set to A4 
-(\texttt{a4paper}) or letter (\texttt{letterpaper}) paper in the
-document class entry, e.g.:\\\verb+\documentclass[a4paper]{easychair}+
-
-\item
-The print area for both letter and A4 paper sizes is 145x224 mm. This size
-has been selected to allow for inexpensive printing using our current
-print-on-demand publisher.
-
-\item
-The base font is Computer Modern. The base font size is 10pt. If you
-use any other font size, there is no guarantee that the produced
-document will look nice or fit into our standard page size.
-
-\item
-The references list is condensed. The default bibliography styles, such as
-\texttt{plain}, \texttt{abbrv}, and \texttt{alpha}, are suggested.
-
-\item
-PNG, JPG, and PDF images are supported, i.e., those that are supported
-by the standard \texttt{graphicx} package \cite{graphicx-package}, and
-render nicely in online versions of PDF documents.  This document
-shows some examples of JPG and PDF images, for example in
-Figure~\ref{fig:easychair-logo}. If the papers are designed for
-publishing in print, the images should be at least 300dpi in
-resolution.
-
-\end{enumerate}
-
-%------------------------------------------------------------------------------
-\subsection{Front Matter}
-\label{sect:front-matter}
-
-The front matter of an {\easychair} article follows the \texttt{article}
-style, augmented with the \verb+\titlerunning+ and \verb+\authorrunning+
-commands for use by authors. 
-For the \verb+\author+ command with multiple authors, use \verb+\and+ to
-separate authors from different institutions, as done in this document.
-Institutions are defined using \verb+\institute+ in a similar way and affiliations of authors assigned using \verb+\inst+.
-Authors must set the \verb+\titlerunning+ and \verb+\authorrunning+.
-For example, the front matter of this document defined the authors
-and title as follows:
-
-%---------------------------------------------------------------
-
-\noindent\small
-\begin{verbatim}
-\title{The {\easychair} Class File\\
-       Documentation and Guide for Authors%
-\thanks{Other people who contributed to this document include Maria Voronkov
-  (Imperial College and EasyChair) and Graham Gough (The University of
-  Manchester).}}
-
-% Authors are joined by \and. Their affiliations are given by \inst, which indexes
-% into the list defined using \institute
-%
-\author{
-Serguei A. Mokhov\inst{1}\thanks{Designed and implemented the class style}
-\and
-   Geoff Sutcliffe\inst{2}\thanks{Did numerous tests and provided a lot of suggestions}
-\and
-   Andrei Voronkov\inst{3}\inst{4}\inst{5}\thanks{Masterminded EasyChair and created 
-     versions 3.0--3.5 of the class style}
-}
-
-% Institutes for affiliations are also joined by \and,
-\institute{
-  Concordia University,
-  Montreal, Quebec, Canada\\
-  \email{mokhov@cse.concordia.ca}
-\and
-   University of Miami,
-   Miami, Florida, U.S.A.\\
-   \email{geoff@cs.miami.edu}\\
-\and
-   University of Manchester,
-   Manchester, U.K.\\
-   \email{andrei@voronkov.com}\\
-\and
-   Chalmers University of Technology,
-   Gothenburg, Sweden
-\and
-   EasyChair
- }
-
-\authorrunning{Mokhov, Sutcliffe and Voronkov}
-\titlerunning{The {\easychair} Class File}
-\end{verbatim}
-\normalsize
-
-%------------------------------------------------------------------------------
-\subsection{Page Numbering}
-\label{sect:page-numbering}
-
-Page numbers are at the bottom of every page. Authors must leave the
-page numbers in as-is. When EasyChair proceedings and Procedia volumes are processed by
-EasyChair, the correct volume page numbers will be inserted
-automatically.
-
-%------------------------------------------------------------------------------
-\subsection{Section Headings and Capitalization}
-\label{sect:section-headings}
-
-Section and paragraph headings in {\easychair} are invoked via the standard 
-commands, such as
-\verb+\section+,
-\verb+\subsection+,
-\verb+\subsubsection+, and
-\verb+\paragraph+.
-Generally, every non-trivial word in a heading must be capitalized according to
-general capitalization guidelines. A reasonable rule to use is that
-all prepositions, coordinating conjunctions and articles having four
-or fewer letters should not be capitalized. If you do not know what it
-means, simply do not capitalize the following words:
-\textit{amid, anti, as, at, atop, but, by, down, for, from, in, into, like,
-near, next, of, off, on, onto, out, over, pace, past, per, plus, qua,
-save, than, till, to, up, upon, via, with, for, and, nor, but, or,
-yet, so, a, an, the.} For example, if you want to call your paper
-``oldest but goldest'', then the proper title for it is ``Oldest but
-Goldest''. ``Oldest but goldest'' is wrong (since ``goldest'' is not
-capitalized) and ``Oldest But Goldest'' is wrong (``but'' should stay
-lower-cased since it belongs to the list of words above.) Needless to
-say, ``OLDEST BUT GOLDEST'' is very wrong.
-
-Paragraph headings should not be capitalized and should have a
-trailing period. That is, you should write
-
-\begin{quote}
-\verb|\paragraph{EasyChair is cool.}|
-\end{quote}
-rather than
-
-\begin{quote}
-\verb|\paragraph{EasyChair is cool}|
-\end{quote}
-unless your aim was to write something like
-
-\begin{quote}
-\verb|\paragraph{EasyChair is cool} when you use it for publishing.| 
-\end{quote}
-Of course, you are welcome to replace the trailing period by any other
-punctuation mark, for example
-
-\begin{quote}
-\verb|\paragraph{EasyChair is cool!}|
-\end{quote}
-
-See the examples in this document, e.g.,
-Section~\ref{sect:typesetting} is a section, this 
-(Section~\ref{sect:section-headings}) is a subsection, and
-Section~\ref{sect:subsubsection-headings} is a subsubsection.
-
-%------------------------------------------------------------------------------
-\subsubsection{Subsubsection Header}
-\label{sect:subsubsection-headings}
-
-This is a subsubsection.
-
-\paragraph{Paragraph header.}
-
-This is a paragraph. 
-One way of saving space when hyper-references are not essential is to 
-use paragraphs instead of subsubsections.
-
-%------------------------------------------------------------------------------
-\subsection{Mathematics}
-\label{sect:mathematics}
-
-Mathematics can be done inline for simple things, e.g., an equation
-$x = 0$, possibly with super and subscripts, e.g., $x^2_k \approx 27$,
-Greek letters, e.g., $\alpha \cup \Theta \ne \gamma$, etc.
-Larger formulae must be done using {\tt \verb|\|[}~~{\tt \verb|\|]}
-bracketing, e.g.,
-\[
-\int_{0}^{1} x dx = \left[ \frac{1}{2}x^2 \right]_{0}^{1} = \frac{1}{2}
-\]
-or using {\tt \verb|\|begin\{equation\}} and {\tt \verb|\|end\{equation\}} for
-numbered equations, e.g.,
-\begin{equation}
-e^x = \sum_{n=0}^\infty \frac{x^n}{n!} = \lim_{n\rightarrow\infty} (1+x/n)^n
-\end{equation}
-
-Use {\tt \verb|\|begin\{align*\}} and {\tt \verb|\|end\{align*\}} (or without
-the {\tt *} include number) to align equations, e.g.,
-\begin{align*}
-x^2 + y^2 &= 1 \\
-y &= \sqrt{1 - x^2}
-\end{align*}
-
-Fonts, using \verb|\matcal| and others can also be used in the math mode: $\mathcal{ALC}$.
-
-%------------------------------------------------------------------------------
-\subsection{Tables}
-\label{sect:tables}
-
-Table~\ref{tab:ltbexample} shows an example of a table of data that was
-conveniently available (i.e., the data has nothing to do with
-{\easychair}) apart from being related to the research of two of its
-authors.
-
-\begin{table}[htp]
-	\begin{centering}
-		\begin{tabular}{lrrrrrrrr}
-		\hline
-		ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
-		                      & /100& time & out & Con. & Eff. & /35 & /40 & /25 \\
-		\hline
-		Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
-		iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
-		SInE 0.4              &  64 & 75.3 &  64 & 0.32 &  8.5 &  26 &  13 &  25 \\
-		leanCoP-SInE 2.1      &  35 &110.8 &  35 & 0.23 &  3.2 &  23 &   1 &  11 \\
-		E-LTB 1.1pre          &  18 & 63.4 &   0 & 0.21 &  2.8 &   7 &   9 &   2 \\
-		EP-LTB 1.1pre         &  18 & 77.8 &  18 & 0.21 &  2.3 &   7 &   9 &   2 \\
-		E-KRH'-LTB 1.1.3      &   0 &   -- &  -- &   -- &   -- &   0 &   0 &   0 \\
-		\hline
-		\end{tabular}
-		\caption{LTB division results}
-		\label{tab:ltbexample}
-	\end{centering}
-\end{table}
-
-%------------------------------------------------------------------------------
-\subsection{References}
-\label{sect:references}
-
-References must be provided in a {\tt .bib} file, so that \BibTeX\ can
-be used to generate the references in a consistent style in a volume.
-The preferred styles are {\tt plain} and {\tt alpha}.
-For example, the references for this paper are generated from the
-lines
-\begin{verbatim}
-    \bibliographystyle{plain}
-    \bibliography{easychair}
-\end{verbatim}
-and a way to compose the entires, e.g. citing this class style~\cite{easychair-latex-class}
-is below:
-\small
-\begin{verbatim}
-    @misc
-    {
-      easychair-latex-class,
-      author       = {Serguei A. Mokhov and Geoff Sutcliffe and Andrei Voronkov},
-      title        = {The {\sf easychair} Class File Documentation and Guide
-                      for Authors},
-      year         = {2008--2011},
-      howpublished = {[online]},
-      note         = {Available at \url{http://easychair.org/easychair.zip}}
-    }
-\end{verbatim}
-\normalsize
-
-%------------------------------------------------------------------------------
-\section{Installation and Usage Instructions}
-\label{sect:installation-usage}
-
-%------------------------------------------------------------------------------
-\subsection{Installation}
-
-The ``installation'' of the {\easychair} document class is easy.
-Download the latest version of the \texttt{easychair.zip} package 
-from \url{http://www.easychair.org/easychair.zip}
-and unzip it in the directory where you will prepare your paper.
-You will get the following files, out of which you may need to keep only 
-the \texttt{easychair.cls} style class if you are familiar with the rest 
-of the files and do not require them to get started.
-We are also working to make {\easychair} available from CTAN \cite{ctan},
-such that it can be installed with the popular \TeX Live \cite{texlive} and
-{\miktex} \cite{miktex} \LaTeX\ package management systems.
-
-\begin{itemize}
-\item
-\texttt{easychair.cls} -- the class file that this is all about.
-
-\item
-\texttt{easychair-a4.pdf} -- the PDF version of this guide rendered using 
-\texttt{a4paper} option.
-
-\item
-\texttt{easychair.tex} -- the {\LaTeX} source of this guide, 
-and
-\texttt{easychair.bib} -- the supporting bibliography entries found starting 
-on page~\pageref{sect:bib}.
-
-\item
-\texttt{logoEC.pdf} -- the PDF version of the EasyChair logo rendered in
-Figure~\ref{fig:easychair-logo} and
-\texttt{throneEC.jpg} -- the JPG version of the easy throne rendered in 
-Figure~\ref{fig:easythrone}.
-
-\end{itemize}
-
-%------------------------------------------------------------------------------
-\subsection{Required Packages}
-
-The {\easychair} class relies only on packages deemed standard and shipped by
-most {\LaTeX} distributions in the worlds of Linux (current \texttt{texlive} \cite{texlive}
-or older \texttt{tetex}), MacOS X,
-and Windows (via Cygwin or {\miktex}).
-If for some reason your distribution is old or doesn't have the packages
-listed below, you can always obtain a copy from CTAN \cite{ctan}. Note
-that EasyChair loads this packages automatically so you don't have to
-use \verb|\usepackage| to use any of these packages.
-
-\begin{itemize}
-\item
-\texttt{url} \cite{url-package} (included also by \texttt{hyperref} automatically) -- to provide URL rendering support for the
-monospaced font, which takes care of special characters as well as line 
-wrapping.
-
-\item
-\texttt{hyperref} \cite{hyperref-package} -- to allow hyperlinking of URLs and
-cross references within an article.
-Its options are set to either \verb+letterpaper+ or \verb+a4paper+, depending
-on the \verb+\documentclass+ options.
-
-\item
-\texttt{graphicx} \cite{graphicx-package} -- the standard package for rendering
-PNG, JPG, and PDF graphic images, primarily in \texttt{figure} environments.
-
-\item
-optional \texttt{mathptmx} \cite{mathptmx-package} -- Times base font for compactness
-(use with the \texttt{withtimes} {\easychair} option).
-
-\item
-\texttt{listings} \cite{listings-package} -- to allow highlighted source code 
-listing styles.
-
-\item
-\texttt{latexsym} \cite{latexsym-package} -- to provide common math and other 
-symbols.
-
-\item
-\texttt{amsthm} \cite{amsthm-package} -- to provide {\AmS} theorem-like 
-environments.
-
-\item
-\texttt{empheq} \cite{empheq-package} -- to provide equation environments, etc.
-
-\item
-\texttt{geometry} \cite{geometry-package} -- to set {\easychair} margins, 
-outlined in Section~\ref{sect:generalities}.
-
-\item
-\texttt{lastpage} \cite{lastpage-package} -- to allow computationally
-referencing the last page.
-
-\item
-\texttt{fancyhdr} \cite{fancyhdr-package} -- for running heads.
-
-\item
-\texttt{footmisc} \cite{footmisc-package} -- to ensure that footnotes are
-always at the bottom.
-
-\item
-optional \texttt{makeidx} \cite{makeidx-package} -- for index generation
-(use with the \texttt{thesis} {\easychair} option).
-
-\item
-\texttt{eso-pic} \cite{eso-pic-package} -- for draft versions and checking page
-overlflows vs. a border drawn around the headers, footers, and the main body of
-the article.
-
-\end{itemize}
-
-%------------------------------------------------------------------------------
-\subsection{Recommended Packages}
-\label{sect:recommended-packages}
-
-Here is a list of some packages that this guide's authors have experimented 
-with, and which are suitable for inclusion if needed by article authors.
-These packages must be loaded using \verb+\usepackage+.
-In general, authors may use any standard packages provided they do not change
-the basic layout and font settings established by the {\easychair} class.
-Such packages must be provided with the submission of articles.
-
-\begin{itemize}
-\item
-\texttt{rotating} \cite{rotating-package} -- to rotate floats (figures and
-tables) on the page, when wide tables or figures do not fit in portrait layout.
-
-\item
-\texttt{pdflscape} \cite{pdflscape-package} -- similar to \texttt{rotating}, 
-but also allows rotating text to make it conveniently viewable in a PDF 
-viewer that supports individual rotated pages.
-A possible disadvantage is that a page break is forced, which may create
-gaps before or after the landscape page.
-
-\item
-\texttt{algorithm2e} \cite{algorithm2e-package} -- provides a figure-like
-algorithm environment for formal algorithm presentation with highlighting.
-
-\end{itemize}
-
-%------------------------------------------------------------------------------
-\subsection{Compiling}
-\label{sect:compiling}
-
-\texttt{pdflatex} \cite{pdflatex-instructions} is the preferred tool for
-producing PDF files with {\easychair} class documents.
-The author kit (\texttt{easychair.zip}) includes some minimal automation 
-that authors can use at their discretion.
-
-\begin{itemize}
-\item
-Linux and UNIX-like platforms (also works under Cygwin and MacOS X):
-A {\makefile} is provided for the GNU \texttt{make} \cite{gmake} utility,
-so this document can be compiled by typing \texttt{make} at the terminal 
-prompt (on the systems where both GNU and non-GNU versions of \texttt{make} 
-are installed, one may need to use \texttt{gmake}).
-
-\item
-Microsoft Windows:
-{\texniccenter} \cite{texniccenter} or {\latexeditor} \cite{led} and
-{\miktex} \cite{miktex} as their backend are common tools
-for {\LaTeX} processing under Microsoft Windows. 
-The former provide a GUI front-end to {\LaTeX}, and the latter is the 
-Windows native-compiled binaries and standard packages with 
-a comprehensive package update tool. 
-The \texttt{easychair.tcp} project file is provided for {\texniccenter} users,
-as well as \texttt{easychair.lpr} for {\latexeditor} users.
-
-\item
-MacOS X:
-TeXShop \cite{texshop} is a tool for {\LaTeX} processing under Mac OS X.
-It provides a GUI front-end to {\LaTeX}. The backend can be installed
-through the \texttt{fink} \cite{fink} repository or the Darwin Ports.
-\end{itemize}
-
-Some authors use packages that require the use of \texttt{latex}
-instead of \texttt{pdflatex}, most notably \texttt{pstricks} and its
-derivatives. If switching to modern packages, such as
-\texttt{tikz} is not an option, these authors can use the {\easychair}
-class with \texttt{latex}.
-
-%------------------------------------------------------------------------------
-\subsection{Bug Reports}
-\label{sec:bug-reports}
-
-Please report bugs, errors, and omissions you find with the {\easychair} 
-class to its current maintainer, Andrei Voronkov,
-at \url{andrei@voronkov.com}. Any \emph{constructive} feedback is
-always welcome. If anybody has a very hard question, we might be
-unable to answer it without Sergei. If anybody knows his whereabouts,
-Geoff and Andrei have the ransom money ready.
-
-%------------------------------------------------------------------------------
-
-\section{Avoiding Text Overflows}
-
-When used in EasyChair, the {\easychair} class is intended for
-publishing, including printing. This implies that the text of your
-document should not overflow the page dimensions. This section gives
-some practical advice on how you can avoid text overflows. When you
-run \texttt{pdflatex} or \texttt{latex} on your files, you sometimes
-get warnings like
-
-\small
-\begin{verbatim}
-  LaTeX Warning: Overfull \hbox (7.43138pt too wide) in paragraph at lines 994--994
-\end{verbatim}
-\normalsize
-or
-
-\small
-\begin{verbatim}
-  LaTeX Warning: Float too large for page by 55.21666pt on input line 638.
-\end{verbatim}
-\normalsize
-They mean what they say: your text overflows the allowed space. When
-you prepare a PDF document that will only be used online or as a
-draft, in many cases you can ignore these warnings, since they only
-make your document uglier than it could be but it remains still
-readable. However, for documents that are supposed to be printed,
-overflows make a lot of difference. There is no universal recipe on
-how to make your text fit into the page. However, there are some
-common cases that can be easily fixed using the tricks described
-here.
-
-\subsection{Using the \texttt{debug} Option to Check for Overflows}
-
-A general guideline is that you should first check how the alleged
-overflows looks in reality. For example, some LaTeX overflows are
-invisible or hardly noticeable for people and overlows by
-less than 4pt look OK anyhow. The {\easychair} class has an option
-\texttt{debug} that adds to each page of your documents a red frame
-corresponding to the allowed page
-dimensions. Figure~\ref{fig:redframe} on page~\pageref{fig:redframe}
-shows an example of a page created using this style. With such a red
-frame, you can clearly see any overflows apart from very minor ones.
-
-\begin{figure}
-  \begin{centering}
-    \includegraphics[width=0.89\textwidth]{debug.pdf}
-    \caption{A page of a document created using the \texttt{debug}
-      option} 
-    \label{fig:redframe}
-  \end{centering}
-\end{figure}
-
-\subsection{Tables}
-
-Many page overflows happen because of large tables. In many case these
-overflows can be easily removed by slightly reducing padding added by
-\LaTeX\ to every column. It is controlled by the \LaTeX\ command
-\verb|\tabcolsep| whose value by default is 6pt. Even small changes in
-the value of this command may give drastic reductions in the width of
-tables. This is illustrated in Figure~\ref{fig:tabcolsep} on
-page~\pageref{fig:tabcolsep}. Note though that there is no free lunch:
-smaller values for this command may result in lower redability.
-
-%------------------------------------------------------------
-\begin{figure}[tb]\small
-\begin{center}
-  \begin{tabular}{lrrrrrrrr}
-    \hline
-    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
-   \hline
-    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
-    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
-   \hline
-  \end{tabular}
-\end{center}
-
-\begin{center}\renewcommand{\tabcolsep}{5pt}
-  \begin{tabular}{lrrrrrrrr}
-    \hline
-    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
-   \hline
-    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
-    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
-   \hline
-  \end{tabular}
-\end{center}
-
-\begin{center}\renewcommand{\tabcolsep}{3pt}
-  \begin{tabular}{lrrrrrrrr}
-    \hline
-    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
-   \hline
-    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
-    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
-   \hline
-  \end{tabular}
-\end{center}
-
-\begin{center}\renewcommand{\tabcolsep}{1pt}
-  \begin{tabular}{lrrrrrrrr}
-    \hline
-    ATP System            & LTB & Avg  &Prfs & SOTA & \multicolumn{1}{c}{$\mu$} & CYC & MZR & SMO \\
-   \hline
-    Vampire-LTB 11.0      &  69 & 24.5 &  69 & 0.37 & 28.1 &  23 &  22 &  24 \\
-    iProver-SInE 0.7      &  67 & 76.5 &   0 & 0.36 &  8.8 &  28 &  14 &  25 \\
-   \hline
-  \end{tabular}
-\end{center}
-\normalsize
-
-\caption{Original table and tables with \texttt{tabcolsep} set to 5pt,
-  3pt, and 1pt
-  \label{fig:tabcolsep}}
-
-\end{figure}
-%------------------------------------------------------------
-
-\subsection{Images}
-
-Images included using \verb|\includegraphics| are easy to resize since
-one can specify the size of the result explicitly. For example,
-Figure~\ref{fig:easythrone} shows three copies of the same image
-having different sizes obtained using the following commands:
-
-\begin{verbatim}
-  \includegraphics[width=0.5\textwidth]{throneEC.jpg}
-  \includegraphics[width=0.3\textwidth]{throneEC.jpg}
-  \includegraphics[width=0.15\textwidth]{throneEC.jpg}
-\end{verbatim}
-
-\subsection{A Universal Recipe}
-
-\LaTeX\ has a very powerful weapon for reducing the size of almost
-anythings. More precisely, it can reduce anything producing what
-\LaTeX\ considers a box. This weapon is called
-\verb|\scalebox|. Consider an example (check the source of this file
-to see how it was produced).
-
-\begin{center}
-\begin{tabular}{|c|c|}
-\hline \Huge
-\begin{tabular}[b]{cr}
-year & users \\ \hline
-2007 &    47,753 \\
- 2008 &    114,494 \\
- 2009 &    207,506 \\
- 2010 &   371,054 \\
-\end{tabular}
-&
-\includegraphics[width=0.28\textwidth]{chairEC}\\
-\hline
-\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
-  logos,}\\
-\multicolumn{2}{|c|}{scaled to the number of users in 2010} \\
-\hline
-\end{tabular}
-\end{center}
-This is what happens when we put (almost) the same \LaTeX\ code in 
-\verb|\scalebox{0.55923}{...}| to scale it down to the number of users
-in 2009:
-
-\begin{center}
-\scalebox{0.55923}{%
-\begin{tabular}{|c|c|}
-\hline \Huge
-\begin{tabular}[b]{cr}
-year & users \\ \hline
-2007 &    47,753 \\
- 2008 &    114,494 \\
- 2009 &    207,506 \\
- 2010 &   371,054 \\
-\end{tabular}
-&
-\includegraphics[width=0.28\textwidth]{chairEC}\\
-\hline
-\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
-  logos,}\\
-\multicolumn{2}{|c|}{scaled to the number of users in 2009} \\
-\hline
-\end{tabular}}
-\end{center}
-We can scale it down even further to the 2008 figure using
-\verb|\scalebox{0.30856}{...}|:
-
-\begin{center}
-\scalebox{0.30856}{%
-\begin{tabular}{|c|c|}
-\hline \Huge
-\begin{tabular}[b]{cr}
-year & users \\ \hline
-2007 &    47,753 \\
- 2008 &    114,494 \\
- 2009 &    207,506 \\
- 2010 &   371,054 \\
-\end{tabular}
-&
-\includegraphics[width=0.28\textwidth]{chairEC}\\
-\hline
-\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
-  logos,}\\
-\multicolumn{2}{|c|}{scaled to the number of users in 2008} \\
-\hline
-\end{tabular}}
-\end{center}
-or further down to 2007:
-
-\begin{center}
-\scalebox{0.12870}{%
-\begin{tabular}{|c|c|}
-\hline \Huge
-\begin{tabular}[b]{cr}
-year & users \\ \hline
-2007 &    47,753 \\
- 2008 &    114,494 \\
- 2009 &    207,506 \\
- 2010 &   371,054 \\
-\end{tabular}
-&
-\includegraphics[width=0.28\textwidth]{chairEC}\\
-\hline
-\multicolumn{2}{|c|}{The number of users of EasyChair and one of its
-  logos,}\\
-\multicolumn{2}{|c|}{scaled to the number of users in 2008} \\
-\hline
-\end{tabular}}
-\end{center}
-
-This size reduction technique is very efficient: using the right scale
-you may post your whole article on Twitter in a single tweet. However,
-it may also may parts of your text virtually unreadable with an
-unfortunate side effect of annoying reviewers.
-
-\section{Submitting Your Article Through EasyChair}
-
-This section is intended only for the authors and editors of EasyChair
-proceedings and Procedia volumes.
-
-When you prepare an article for either of these, it should be
-submitted through EasyChair. EasyChair automates the submission
-process as much as possible and goes to a great length to ensure that
-your article can be published and printed. Publication for EasyChair
-means much more than just putting a PDF of your article online. It
-collects a some meta-information about the article to classify it,
-find similar articles, make it easily searchable, and index it in
-various Web services, such as DBLP. This section explains how
-EasyChair processes your article.
-
-\subsection{Submitting the Article}
-
-Submitting the article is easy. All you should do is to put the source
-of your article in a single zip file. The source must contain all
-auxiliary files required to create a PDF file of your article: this
-includes images, bibliography, and all non-standard \LaTeX\ packages
-you used\footnote{ 
-  A non-standard \LaTeX\ package is a package that is not included in
-  CTAN. 
-} For example, suppose that your main \LaTeX\ file is
-\texttt{main.tex}, it inputs another file \texttt{macros.tex} and uses
-the file \texttt{biblio.bib} to produce the bibliography. Suppose it
-also uses two images \texttt{images/easy.jpg} and
-\texttt{images/easy.jpg}. Then you should create a zip archive
-containing all these files. Suppose all these files are put in a
-directory \texttt{mypaper} on your computer, where \texttt{images} is
-a subdirectory of \texttt{mypaper}
-
-On almost any operating system (Linux,
-Windows, or Mac) you can achieve this by using the following sequence
-of commands:
-
-\begin{verbatim}
-cd mypaper
-zip -r mypaper.zip *
-\end{verbatim}
-This will create a zip archive \texttt{mypaper.zip} including all
-files in the directory \texttt{mypaper} and its subdirectories.
-
-\subsection{EPiC Series and Kalpa Publications}
-\label{subsec:EPiC}
-
-If you publish your article in the EasyChair EPiC series, our aim is to produce
-a nicely looking EPiC Series header, see Figure~\ref{fig:epicheader} on page
-\pageref{fig:epicheader} for an idea. The images and information in the header
-will be added by EasyChair automatically and depend on the subseries and the
-volume information.
-
-\begin{figure}
-  \begin{centering}
-    \includegraphics[width=0.89\textwidth]{epic.pdf}
-    \caption{This document with an EPiC Series header} 
-    \label{fig:epicheader}
-  \end{centering}
-\end{figure}
-
-To ensure that the layout of your document will not change when the header is
-added, you should use the option \texttt{EPiC} as follows:
-
-\small
-\begin{verbatim}
-\documentclass[EPiC]{easychair}
-\end{verbatim}
-\normalsize
-This will reserve space for the header on the top of the first page of your
-article.
-
-%------------------------------------------------------------------------------
-\section{Future Work}
-\label{sect:future-work}
-
-We plan to further strengthen the {\easychair} class and promote it for 
-electronic publishing for EasyChair-powered conferences and workshops,
-and take over the world, as shown in Figure~\ref{fig:easythrone}. We
-aim at creating a new model of \emph{affordable publishing}, where
-anybody can become a low-cost publisher. The
-\href{http://www.easychair.org/publications/EPiC}{EasyChair EPiC
-  Series} and
-the
-\href{http://www.easychair.org/publications/Kalpa}{EasyChair Kalpa Publications} 
-are our first contributions to affordable open access publishing.
-
-\begin{figure}[tb]
-  \begin{centering}
-    \includegraphics[width=0.5\textwidth]{throneEC.jpg}
-    \includegraphics[width=0.3\textwidth]{throneEC.jpg}
-    \includegraphics[width=0.15\textwidth]{throneEC.jpg}
-  \end{centering}
-  \caption{Why one should use EasyChair}
-  \label{fig:easythrone}
-\end{figure}
-
-%------------------------------------------------------------------------------
-\section{Acknowledgments}
-\label{sect:acks}
-
-\begin{itemize}
-\item
-Aleksander Kosenkov for the graphics that are used here and on the EasyChair 
-website~\cite{easychair}.
-
-\item
-Graham Gough (The University of Manchester) for implementing title-related
-commands.
-
-\item
-Thomas F. Sturm (Universit\"at der Bundeswehr M\"unchen) for implementing the
-\LaTeX\ part of the headers used for the EasyChair EPiC series.
-
-\item
-The CTAN \cite{ctan} and {\LaTeX} communities \cite{texniccenter,miktex}.
-
-\item
-Guilin Qi, Jasmin Christian Blanchette, Leslie Lamport, Uwe Pfeiffer,
-and others for constructive feedback on the style, most of which had been
-incorporated in version 2 of the class style.
-\end{itemize}
-
-\label{sect:bib}
-\bibliographystyle{plain}
-%\bibliographystyle{alpha}
-%\bibliographystyle{unsrt}
-%\bibliographystyle{abbrv}
-\bibliography{easychair}
-
-%------------------------------------------------------------------------------
-\appendix
-\section{{\easychair} Requirements Specification}
-\label{sect:easychair-requirements}
-
-The following high-level requirements were set for the development of 
-the {\easychair} class, and were refined as development went along.
-
-\begin{enumerate}
-\item
-The style should be easy to use. 
-The average {\LaTeX} user should not need to read a long manual.
-
-\item
-It should be economical in space but the text should be nice-to-read.
-
-\item
-It should use fonts producing a reasonable-quality PDF.
-
-\item
-The bibliography should produce hyperlinks.
-
-\item
-Sections should produce menu sections in PDF.
-
-\item
-The text should look good on both A4 and letter paper.
-
-\item
-The style should be single-column for convenience of scrolling.
-
-\item
-The print area should be convenient for printing using print-on-demand publishers.
-
-\item
-Running heads.
-\end{enumerate}
-
-%------------------------------------------------------------------------------
-% Index
-%\printindex
-
-%------------------------------------------------------------------------------
-\end{document}

### 2020-04-24 coqws/easychair.tex
cleanup

@@
 
-\section{Context}
+\section{Context and Talk proposal}
 
@@
 construction of a hierarchy hard even for expert users. The difficulty
-begins with the non orthogonal choices of the encoding style, from bundled to
+begins with the non-orthogonal choices of the encoding style, from bundled to
 unbundled, and inference mechanism, from Type Classes to Canonical Structures.
-To this one may add the concerns about performance and finally the ones about
-the usability of the final hierarchy.
+To this, one may add the concerns about performance and about
+the usability, by a non expert, of the final hierarchy.
 
@@
 to the primitive ones, a feature that can also be used to accommodate changes
-without breaking user code.
+to the hierarchy without breaking user code.
 
@@
 are given in~\cite{cohen:hal-02478907}.
-In this talk we focus on how to take advantage
-of \HB{} in a Coq development, and we plan to not detail how it works
-behind the scenes, unless specific questions arise from the audience.
+In this talk we focus on how to use \HB{} in a Coq development, and we plan
+to not detail how it works behind the scenes, unless specific questions
+arise from the audience.
 
@@
 
-For space constraints we omit the 30 lines in which we use
-the \coq{HB.factory} command to declare an alternative interface
-to build examples of the \coq{Ring} structure (the green
-arrow in the picture) by providing all proofs at once but for the
-one on the commutativity of the Monoid operation:
-the other Ring axioms imply it~\cite{nearrings}. Factories can also be used to
-preserve backward compatibility when an hierarchy evolves to accommodate
-new intermediate  structures.
+For space constraints we omit the declaration of the green arrow
+in the picture (about 30 lines) which is an alternative interface
+to build examples of the \coq{Ring} structure by providing all
+proofs at once. In that case the user is spared to prove the commutativity
+of the Monoid operation since it is implied by the other axioms~\cite{nearrings}.
 

### 2020-04-25 hb.tex
color HB commands differently
fix

--- 2020-04-24 hb.tex
@@
 
-Definition Z_Ring_axioms : Ring_of_Monoid Z _ :=                        #\label{demo:theory:z:ring:axioms}#
+Definition Z_Ring_axioms : Ring_of_Monoid Z :=                        #\label{demo:theory:z:ring:axioms}#
   Ring_of_Monoid.Build Z 1 Z.opp Z.mul

### 2020-04-25 hb.tex
appendix with full V4 code

@@
 
+\section{Full V4 code}
+\begin{coqcode}
+From Coq Require Import ssreflect ssrfun ZArith.
+From HB Require Import structures.
+
+Declare Scope hb_scope.
+Delimit Scope hb_scope with G.
+Open Scope hb_scope.
+
+HB.mixin Record Monoid_of_Type A := {
+  zero : A;
+  add : A -> A -> A;
+  addrA : associative add;
+  add0r : left_id zero add;
+  addr0 : right_id zero add;
+}.
+HB.structure Definition Monoid := { A of Monoid_of_Type A }.
+Notation "0" := zero : hb_scope.
+Infix "+" := (@add _) : hb_scope.
+
+HB.mixin Record AbelianGroup_of_Monoid A of Monoid A := {
+  opp : A -> A;
+  addrC : commutative (add : A -> A -> A);
+  addNr : left_inverse zero opp add;
+}.
+HB.structure Definition AbelianGroup := { A of Monoid A & AbelianGroup_of_Monoid A }.
+Notation "- x" := (@opp _ x) : hb_scope.
+Notation "x - y" := (x + - y) : hb_scope.
+
+HB.mixin Record SemiRing_of_Monoid A of Monoid A := {
+  one : A;
+  mul : A -> A -> A;
+  mulrA : associative mul;
+  mul1r : left_id one mul;
+  mulr1 : right_id one mul;
+  mulrDl : left_distributive mul add;
+  mulrDr : right_distributive mul add;
+  mul0r : left_zero zero mul;
+  mulr0 : right_zero zero mul;
+}.
+HB.structure Definition SemiRing := { A of Monoid A & SemiRing_of_Monoid A }.
+Notation "1" := one : hb_scope.
+Infix "*" := (@mul _) : hb_scope.
+
+Lemma addrN {R : AbelianGroup.type} : right_inverse (zero : R) opp add.
+Proof. by move=> x; rewrite addrC addNr. Qed.
+
+HB.factory Record Ring_of_AbelianGroup A of AbelianGroup A := {
+  one : A;
+  mul : A -> A -> A;
+  mulrA : associative mul;
+  mul1r : left_id one mul;
+  mulr1 : right_id one mul;
+  mulrDl : left_distributive mul add;
+  mulrDr : right_distributive mul add;
+}.
+
+HB.builders Context (A : Type) (f : Ring_of_AbelianGroup A).
+
+  Fact mul0r : left_zero zero mul_f.
+  Proof.
+  move=> x; rewrite -[LHS]add0r addrC.
+  rewrite -{2}(addNr (mul_f x x)) (addrC (opp _)) addrA.
+  by rewrite -mulrDl_f add0r addrC addNr.
+  Qed.
+
+  Fact mulr0 : right_zero zero mul_f.
+  Proof.
+  move=> x; rewrite -[LHS]add0r addrC.
+  rewrite -{2}(addNr (mul_f x x)) (addrC (opp _)) addrA.
+  by rewrite -mulrDr_f add0r addrC addNr.
+  Qed.
+
+  Definition to_SemiRing_of_Monoid := SemiRing_of_Monoid.Build A _ mul_f mulrA_f
+     mul1r_f mulr1_f mulrDl_f mulrDr_f mul0r mulr0.
+  HB.instance A to_SemiRing_of_Monoid.
+
+HB.end.
+HB.structure Definition Ring := { A of AbelianGroup A & Ring_of_AbelianGroup A }.
+
+HB.factory Record Ring_of_Monoid A of Monoid A := {
+  one : A;
+  opp : A -> A;
+  mul : A -> A -> A;
+  addNr : left_inverse zero opp add;
+  addrN : right_inverse zero opp add;
+  mulrA : associative mul;
+  mul1r : left_id one mul;
+  mulr1 : right_id one mul;
+  mulrDl : left_distributive mul add;
+  mulrDr : right_distributive mul add;
+}.
+
+HB.builders Context (A : Type) (f : Ring_of_Monoid A).
+
+  Lemma addrC : commutative (add : A -> A -> A).
+  Proof.
+  have addKl (a b c : A) : a + b = a + c -> b = c.
+    apply: can_inj (add a) (add (opp_f a)) _ _ _.
+    by move=> x; rewrite addrA addNr_f add0r.
+  have addKr (a b c : A) : b + a = c + a -> b = c.
+    apply: can_inj (add ^~ a) (add ^~ (opp_f a)) _ _ _.
+    by move=> x; rewrite /= -addrA addrN_f addr0.
+  have innerC (a b : A) : a + b + (a + b) = a + a + (b + b).
+    by rewrite -[a+b]mul1r_f -mulrDl_f mulrDr_f !mulrDl_f !mul1r_f.
+  move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
+  by rewrite -!addrA [in RHS]addrA innerC !addrA.
+  Qed.
+
+  Definition to_AbelianGroup_of_Monoid := AbelianGroup_of_Monoid.Build A opp_f addrC addNr_f.
+  HB.instance A to_AbelianGroup_of_Monoid.
+
+  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build A one_f mul_f
+    mulrA_f mul1r_f mulr1_f mulrDl_f mulrDr_f.
+  HB.instance A to_Ring_of_AbelianGroup.
+HB.end.
+\end{coqcode}
+
 \end{document}

### 2020-04-25 hb.tex
typesetting + ref to appendixes

@@
 The first version of our hierarchy (that we name V1) features only two
-structures: \coq{Monoid} and \coq{Ring}.
+structures: \coq{Monoid} and \coq{Ring}. We use notations from Appendix~\ref{sec:coq-reference} to define them.
 %\marginpar{remove hb.end for \mixins{}, use next release of coq elpi}
@@
 Lemma addrC : forall R : Ring.type, commutative (@add R).                       #\label{demo:theory:state:addrC}#
-Proof. (* Proof by Hankel 1867, omitted for brevity *) Qed.
+Proof. (* Proof by Hankel 1867, in Appendix B *) Qed.
 
@@
 \begin{coqcode}
-#\hbfactory{}# Record Ring_of_Monoid A of Monoid A := {
-  ... (* unchanged, see page 2 line 13 *) ...
-}.
+#\hbfactory{}# Record Ring_of_Monoid A of Monoid A := { ... (* unchanged *) ...}.
+
 #\hbbuilders{}# Context A (f : Ring_of_Monoid A).                          #\label{demo3:variable:f}#
@@
 
-  Definition to_Ring_of_AbelianGroup :=
-    Ring_of_AbelianGroup.Build A (one f) (mul f)
-      (mulrA f) (mul1r f) (mulr1 f) (mulrDl f) (mulrDr f).
+  Definition to_Ring_of_AbelianGroup :=  Ring_of_AbelianGroup.Build A
+    (one f) (mul f) (mulrA f) (mul1r f) (mulr1 f) (mulrDl f) (mulrDr f).
   #\hbinstance{}# A to_Ring_of_AbelianGroup.
@@
 multiplication operator \coq{mul} and its identity element \coq{one}.
+The full code of this V4 is available in Appendix~\ref{sec:full-v4-code}.
 
@@
 \section{Coq reference}
-
+\label{sec:coq-reference}
 \begin{coqcode}
 Section OperationProperties.
-Variable T : Type.
-Variable e : T.
-Variable inv : T -> T.
-Variable op : T -> T -> T.
-Variable add : T -> T -> T.
+
+Variables (T : Type) (e : T) (inv : T -> T) (op : T -> T -> T) (add : T -> T -> T).
 
@@
 Definition left_inverse := forall x, op (inv x) x = e.
+Definition right_inverse := forall x, op x (inv x) = e.
 
@@
 Definition right_distributive := forall x y z, op x (add y z) = add (op x y) (op x z).
+
+End OperationProperties.
 \end{coqcode}
@@
 \section{Proof of addrC}
+\label{sec:proof-addrc}
 \begin{coqcode}
@@
 \section{Full V4 code}
+\label{sec:full-v4-code}
 \begin{coqcode}
@@
 
+(* Bottom mixin in Fig. 2. *)
 HB.mixin Record Monoid_of_Type A := {
@@
 
+(* Bottom right mixin in Fig. 2. *)
 HB.mixin Record AbelianGroup_of_Monoid A of Monoid A := {
@@
 
+(* Bottom left mixin in Fig. 2. *)
 HB.mixin Record SemiRing_of_Monoid A of Monoid A := {
@@
 
+(* Top right factory in Fig. 2. *)
 HB.factory Record Ring_of_AbelianGroup A of AbelianGroup A := {
@@
 
+(* Builder arrow from top right to bottom left *)
 HB.builders Context (A : Type) (f : Ring_of_AbelianGroup A).
@@
 
+(* Top left factory in Fig. 2. *)
+(* It is an exact copy of the bottom right mixin. *)
+HB.factory Definition Ring_of_SemiRing A of SemiRing A := AbelianGroup_of_Monoid A.
+(* The corresponding builder is the identity. *)
+HB.builders Context (A : Type) (f : Ring_of_SemiRing A).
+  Definition to_AbelianGroup_of_Monoid : AbelianGroup_of_Monoid A := f.
+  HB.instance A to_AbelianGroup_of_Monoid.
+HB.end.
+
+(* Right-most factory in Fig. 2. *)
 HB.factory Record Ring_of_Monoid A of Monoid A := {
@@
 
+  (* Builder to the bottom right mixin. *)
   Definition to_AbelianGroup_of_Monoid := AbelianGroup_of_Monoid.Build A opp_f addrC addNr_f.
@@
 
+  (* Builder to the top right factory, which is compiled to the bottom left mixin. *)
   Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build A one_f mul_f
@@
   HB.instance A to_Ring_of_AbelianGroup.
+
 HB.end.

### 2020-04-25 hb.tex
fixing overfull hbox
more typesetting + clarifications about Appendix C

@@
 
-\supplement{Source code of the Coq package: \url{https://github.com/math-comp/hierarchy-builder}}
+\supplement{Coq package source: \url{https://github.com/math-comp/hierarchy-builder}}
 \keywords{Algebraic Hierarchy, Packed Classes, Coq, Elpi, Metaprogramming, $\lambda$Prolog}%TODO mandatory; please add comma-separated list of keywords
@@
 
-As depicted in figure~\ref{fig:puzzle} we change again the hierarchy
+As depicted in Figure~\ref{fig:puzzle} we change again the hierarchy
 by declaring a \coq{Ring_of_Monoid} \factory{}, that, from the user point of view,
@@
 We describe mathematical structures with three kinds of dependent records: mixins, classes, and structures.
-As shown in section \ref{sec:example}, a \mixin{} gathers operators and axioms newly introduced by a structure.
-As in \cite[section 2.4]{DBLP:conf/tphol/GarillotGMR09}\cite[section 2]{KSdraft}, a class record is parametrized by the carrier type \coq{(T : Type)} and gathers all the operators and axioms of a structure by assembling mixins, and a \coq{Structure type} (a record) bundles a carrier and its class instance, as follows.
+As shown in Section \ref{sec:example}, a \mixin{} gathers operators and axioms newly introduced by a structure.
+As in \cite[Section 2.4]{DBLP:conf/tphol/GarillotGMR09}\cite[Section 2]{KSdraft}, a class record is parametrized by the carrier type \coq{(T : Type)} and gathers all the operators and axioms of a structure by assembling mixins, and a \coq{Structure type} (a record) bundles a carrier and its class instance, as follows.
 \begin{coqcode}
@@
 In order to introduce multiple inheritance, we extend the hierarchy described in
-section~\ref{subsec:v3} with the structure of semirings%, which is another
+Section~\ref{subsec:v3} with the structure of semirings %, which is another
 %intermediate structure between monoids and rings
-as depicted in figure \ref{fig:puzzle_v4}. Semirings introduce the binary
-multiplication operator \coq{mul} and its identity element \coq{one}.
-The full code of this V4 is available in Appendix~\ref{sec:full-v4-code}.
+as depicted in Figure~\ref{fig:puzzle_v4}, and name it V4. Semirings introduce the
+binary multiplication operator \coq{mul} and its identity element
+\coq{one}.  For the sake of completeness, the full code of V4 is available in
+Appendix~\ref{sec:full-v4-code}.
 
@@
 
-(* Builder arrow from top right to bottom left *)
+(* Builder arrow from top right to bottom left in Fig. 2. *)
 HB.builders Context (A : Type) (f : Ring_of_AbelianGroup A).

### 2020-04-26 hb.tex
my 2 C

@@
 \Mixins{} are declared by the user as the fundamental building blocks of
-a hierarchy. As the next proposition shows they shall not be distinguished
-from regular \factories{}, since they are a special cases.
+a hierarchy. As the next proposition shows they shall not be regarded as
+different from regular \factories{}, since they are a special cases.
 
@@
 \end{coqcode}
-  Creates a module \coq{M} with a record which is exported as
-  \coq{M}, and which depends on the \mixins{}
+  Declares the record \coq{M} inside an homonymous module \coq{M}.
+  The record \coq{M.M} depends on the \mixins{}
   \(\requires{}(\coq{M}) = \dep{}(\coq{M}) =
-  \provides{}^\star\left(\enum{f_1}{f_n}\right)\), and registers this
-  new record both as a \mixin{} and a \factory{}. It also exports an
-  \phantterm{} \coq{M.Build.}
+  \provides{}^\star\left(\enum{f_1}{f_n}\right)\)
+  and is registered both as a \mixin{} and a \factory{}. Finally It exports an
+  \phantterm{} \coq{M.Build} and a notation
+  \coq{M} standing for \coq{M.M}, so that the module name can be used to
+  denote the homonymous record it contains.
 \end{command}
@@
 \end{coqcode}
-  Creates a module \coq{F} with a record which is exported as
-  \coq{F}, and which depends on the mixins
+  Declares the record \coq{F} inside an homonymous module \coq{F}.
+  The record \coq{F.F} depends on the mixins
   \(\requires{}(\coq{M}) =
-  \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and registers this
-  new record as a factory.  It also exports an
-  \phantterm{} \coq{F.Build.} We then use the \factoryinstances{}
-  \(b_{n+1}, \ldots, b_{n+k}\) provided by the user and derive
-  \mixinbuilders{} from them, so that
+  \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and is registered
+  as a factory.  It exports an
+  \phantterm{} \coq{F.Build.} and a notation
+  \coq{F} standing for \coq{F.F}, so that the module name can be used to
+  denote the homonymous record it contains.
+  Finally it uses the \factoryinstances{}
+  \(b_{n+1}, \ldots, b_{n+k}\) provided by the user in order to derive
+  \mixinbuilders{}, so that
   \(\provides{}(\coq{F}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)
-
   It is thus necessary that
@@
 \factories{} provide \mixins{} through \mixinbuilders{}, we
-obtain \mixinbuilders{} out of each \(b_i\) by function composition.
+can obtain \mixinbuilders{} out of each \(b_i\) by function composition.
 
@@
 \begin{definition}[\C{}, class]
-A \class{} \coq{c} \(\in \C{}\) is distinguished a Coq record with one parameter
+A \class{} \coq{c} \(\in \C{}\) is a distinguished Coq record with one parameter
 \coq{(T : Type)}. The type of each field is a \mixin{} in \M{} applied to

### 2020-04-26 hb.tex
typo

@@
 In the second case,
-in order to show Coq that a data type and some operations for a structure,
+in order to show Coq that a data type and some operations forms a structure,
 one must pick the axioms of the structure an prove them for that specific

### 2020-04-26 hb.tex
a last check on section names
my other 2 C ;)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\HB{} language}\label{sec:language}
+\section{The \HB{} language}\label{sec:language}
 
@@
 
-\subsection{\HB{} \mixins{}, \factories{} and instances}
+\subsection{\Mixins{}, \factories{} and instances}
 \label{sec:hb-mixins-factories}
@@
 a hierarchy. As the next proposition shows they shall not be regarded as
-different from regular \factories{}, since they are a special cases.
+different from regular \factories{}, since they are a special case.
 
@@
 \end{coqcode}
-  Declares the record \coq{M} inside an homonymous module \coq{M}.
-  The record \coq{M.M} depends on the \mixins{}
-  \(\requires{}(\coq{M}) = \dep{}(\coq{M}) =
+  Declares the record \coq{axioms} inside a module \coq{M}.
+  This record \coq{M.axioms} depends on the \mixins{}
+  \(\requires{}(\coq{M.axioms}) = \dep{}(\coq{M.axioms}) =
   \provides{}^\star\left(\enum{f_1}{f_n}\right)\)
@@
   \phantterm{} \coq{M.Build} and a notation
-  \coq{M} standing for \coq{M.M}, so that the module name can be used to
-  denote the homonymous record it contains.
+  \coq{M} standing for \coq{M.axioms}, so that the module name can be used to
+  denote the \coq{axioms} record it contains.
 \end{command}
@@
 \end{coqcode}
-  Declares the record \coq{F} inside an homonymous module \coq{F}.
-  The record \coq{F.F} depends on the mixins
-  \(\requires{}(\coq{M}) =
+  Declares the record \coq{axioms} inside a module \coq{F}.
+  This record \coq{F.axioms} depends on the mixins
+  \(\requires{}(\coq{F.axioms}) =
   \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and is registered
   as a factory.  It exports an
-  \phantterm{} \coq{F.Build.} and a notation
-  \coq{F} standing for \coq{F.F}, so that the module name can be used to
-  denote the homonymous record it contains.
+  \phantterm{} \coq{F.Build} and a notation
+  \coq{F} standing for \coq{F.axioms}, so that the module name can be used to
+  denote the \coq{axioms} record it contains.
   Finally it uses the \factoryinstances{}
@@
 \factories{} provide \mixins{} through \mixinbuilders{}, we
-can obtain \mixinbuilders{} out of each \(b_i\) by function composition.
-
-\subsection{\HB{} classes and structures}
+obtain \mixinbuilders{} out of each \(b_i\) by function composition.
+
+\subsection{Classes and structures}
 \label{sec:hb-class-struct}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The implementation in Coq-Elpi}\label{sec:implementation}
+\section{The implementation of \HB{} in Coq-Elpi}\label{sec:implementation}
 

### 2020-04-27 coqws/easychair.tex
spacing + missing property in Abelian groups

@@
   addNr : left_inverse zero opp add; (* add is available here *)
+  addrN : right_inverse zero opp add;
 }.
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "easychair.tex"
+%%% TeX-command-extra-options: "-shell-escape"
+%%% ispell-local-dictionary: "english"
+%%% End:

### 2020-04-27 hb.tex
Explain the flat variant of packed classes

@@
 
-The language of packed classes were introduced in
-\cite{DBLP:conf/tphol/GarillotGMR09} and used directly to describe the
-algebraic hierarchy of the Mathematical Components library.
-It is based on a disciplined use of records and projections and
-on the possibility of extending the elaborator of Coq
-via the declaration of Canonical Structures instances.
-In this section we describe the \emph{flat} variant of
-Packed Classes, the target language of \HB{}.
+The language of Packed Classes~\cite{DBLP:conf/tphol/GarillotGMR09} is used
+directly to describe the algebraic hierarchy of the Mathematical Components
+library. It is based on a disciplined use of records and projections and on the
+possibility of extending the elaborator of Coq via the declaration of Canonical
+Structures~\cite{DBLP:conf/itp/MahboubiT13} instances. In this section, we
+describe the \emph{flat} variant of Packed Classes, the target language of
+\hb{}, through the hierarchy V3 extended with semirings.
+
+The \hb{} addon can generate all the Coq declarations in this section
+automatically, but some details are omitted for brevity in this section.
 
@@
 \end{coqcode}
+In the flat variant of Packed Classes, a class record gathers mixins as its
+fields directly as in above. In contrast, a class record of the non-flat variant
+packs a class instance of one of the superclasses with remaining mixins, which
+reduces the amount of code to implement inheritance significantly.
+Nested class records and chains of their projections exponentially slow down the
+conversion while the flat variant can mitigate the issue. Since we do not need
+to care about the amount of code in automated generation, \hb{} uses the flat
+variant of Packed Classes as its target language.
+
 As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
@@
 \subsection{Multiple inheritance}
+
+\begin{wrapfigure}[6]{r}{.4\textwidth}
+ \includegraphics[width=.4\textwidth]{v4.pdf}
+ \caption{V4 introduces multiple inheritance. For brevity we omit the factories/builders for the upper arrows.}
+ \label{fig:puzzle_v4}
+\end{wrapfigure}
 In order to introduce multiple inheritance, we extend the hierarchy described in
@@
 \begin{coqcode}
-#\hbmixin{}# Record SemiRing_of_Monoid A of Monoid A := {
+#\hbmixin{}# Record SemiRing_of_Monoid A
+  of Monoid A := {
   one : A;
@@
 }.
-\end{coqcode}
-
-\begin{wrapfigure}[3]{r}{.4\textwidth}
-  \includegraphics[width=.4\textwidth]{v4.pdf}
-  \caption{V4 introduces multiple inheritance. For brevity we omit the factories/builders for the upper arrows.}
-  \label{fig:puzzle_v4}
-\end{wrapfigure}
-\
-\begin{coqcode}
+
 #\hbfactory{}# Record Ring_of_AbelianGroup A
@@
 Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin :
-    Monoid_of_Type T;
-  SemiRing_of_Monoid_mixin :
-    SemiRing_of_Monoid T Monoid_of_Type_mixin;
+  Monoid_of_Type_mixin : Monoid_of_Type T;
+  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin;
 }.

### 2020-04-27 coqws/easychair.tex
update coqws paper raping TeX

@@
 
-\section{Context and Talk proposal}
+\paragraph{Context and Talk proposal}
 
@@
 
-\section{A hierarchy with a diamond in 50 lines}
+\paragraph{A hierarchy with a diamond in 50 lines}
 
 \begin{coqcode}
-HB.mixin Record CMonoid_of_Type A := { (* The set of axioms making A a commutative monoid *)
+HB.mixin Record CMonoid_of_Type A := { (* The set of axioms making A a commutative monoid. *)
   zero  : A;
   add   : A -> A -> A;
-  addrA : associative add;
-  addrC : commutative add;
-  add0r : left_id zero add;
+  addrA : associative add;  (* `add` is associative  *)
+  addrC : commutative add;  (* `add` is commutative  *)
+  add0r : left_id zero add; (* `zero` is the neutral *)
 }.
-
-HB.structure Definition CMonoid := { A of CMonoid_of_Type A }. (* The structure thereof *)
-
+HB.structure Definition CMonoid := { A of CMonoid_of_Type A }. (* The structure thereof. *)
 Notation "0" := zero.
@@
 
-Check 0.     (* ?M : CMonoid.type |- 0 : ?M *)
+(* The type of the operations and axioms depend on a CMonoid.type structure. *)
 Check addrC. (* ?M : CMonoid.type |- forall x y : ?M, x + y = y + x *)
+
+HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
+  (* We can write `add` here since A is a  CMonoid   *)
+  opp   : A -> A;
+  addNr : left_inverse zero opp add; (* `opp` is the additive inverse *)
+}.
 \end{coqcode}
@@
 
-\begin{wrapfigure}[0]{r}{.4\textwidth}
-  \includegraphics[width=.4\textwidth]{diagram.pdf}
-  %\caption{The hierarchy}
-  %\label{fig:hierarchy}
+\begin{wrapfigure}[0]{r}{.35\textwidth}
+  \vspace{1em}
+  \includegraphics[width=.35\textwidth]{diagram.pdf}
+  \vspace{-1.5em}
+  %\caption{
+    \center{\small{Figure 1: The hierarchy}}
+   % }
+  \label{fig:hierarchy}
 \end{wrapfigure}
@@
 \begin{coqcode}
-HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
-  opp   : A -> A;
-  addNr : left_inverse zero opp add; (* add is available here *)
-  addrN : right_inverse zero opp add;
-}.
 
-HB.structure Definition AbelianGrp :=
-  { A of AbelianGrp_of_CMonoid A & }.
+HB.structure Definition AbelianGrp := { A of AbelianGrp_of_CMonoid A & }.
+Notation "- x" := (opp x).
+Notation "x - y" := (add x (opp y)).
 
-Notation "- x" := (opp x).
-
-Check forall G : AbelianGrp.type, forall x : G, x + - x = 0.
+(* A statement in the signature of an Abelian group G. *)
+Check forall G : AbelianGrp.type, forall x : G, x - x = 0.
 
@@
   mul    : A -> A -> A;
-  mulrA  : associative mul;
-  mul1r  : left_id one mul;
-  mulr1  : right_id one mul;
-  mulrDl : left_distributive mul add;
-  mulrDr : right_distributive mul add;
-  mul0r  : left_zero zero mul;
-  mulr0  : right_zero zero mul;
+  mulrA  : associative mul;  (* `mul` is associative   *)
+  mul1r  : left_id one mul;  (* `one` is left neutral  *)
+  mulr1  : right_id one mul; (* `one` is right neutral *)
+  mulrDl : left_distributive mul add;  (* `mul` distributes over *)
+  mulrDr : right_distributive mul add; (*   `add` on both sides  *)
+  mul0r  : left_zero zero mul;  (* `zero` is absorbing `mul`     *)
+  mulr0  : right_zero zero mul; (*   on both sides               *)
 }.
-
 HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & }.
-
 Notation "1" := one.
@@
 
-Check forall S : SemiRing.type, forall x : S, x * x = 0 + 1. (* can I use - here? *)
+(* A statement in the signature of a Semiring S.  *)
+Check forall S : SemiRing.type, forall x : S, x * 1 + 0 = x.
 
+(* We need no additional mixin to declare the structure of rings. *)
 HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A & }.
 
-Check forall R : Ring.type, forall x : R, x * - (1 + x) = 0 + 1.
+(* A statement in the signature of a Ring R.  *)
+Check forall R : Ring.type, forall x : R, x * -1 = - x.
 
+(* We equip Z with the structures of CMonoid, AbelianGrp and SemiRing; HB equips it with Ring. *)
 Definition Z_CMonoid    := CMonoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
@@
 
+(* A statement in the signature of the Z ring *)
 Check forall x : Z, x * - (1 + x) = 0 + 1.
@@
 
-For space constraints we omit the declaration of the green arrow
-in the picture (about 30 lines) which is an alternative interface
-to build examples of the \coq{Ring} structure by providing all
-proofs at once. In that case the user is spared to prove the commutativity
-of the Monoid operation since it is implied by the other axioms~\cite{nearrings}.
-
+For space constraints we omit the declaration
+of the green arrows (Figure 1) which are alternative interfaces
+called factories. They provide specialized interfaces to build structures for others.
+Factories are used to accommodate the insertion of
+intermediate structures in a hierarchy without breaking user code, as we all
+as to implement shortcuts: the factory to build a \coq{Ring} from \coq{Type}
+(rightmost green arrow)
+can omit the one for the commutativity of the addition, since it is implied
+by the other axioms~\cite{nearrings}.
 \bibliography{biblio}

### 2020-04-27 coqws/easychair.tex
minor

@@
 
-\paragraph{A hierarchy with a diamond in 50 lines}
+\paragraph{Example: A hierarchy with a diamond in \HB{}}
 
@@
 For space constraints we omit the declaration
-of the green arrows (Figure 1) which are alternative interfaces
-called factories. They provide specialized interfaces to build structures for others.
+of the green arrows in Figure 1 which are alternative interfaces
+called factories. They provide specialized interfaces to
+build structures for others.
 Factories are used to accommodate the insertion of
@@
 (rightmost green arrow)
-can omit the one for the commutativity of the addition, since it is implied
+can omit the commutativity axiom of the addition, since it is implied
 by the other axioms~\cite{nearrings}.

### 2020-04-27 coqws/easychair.tex
fix caption
fix Ring

@@
 \usepackage{wrapfig}
+\usepackage[font=small,labelfont=bf]{caption}
 
@@
   \vspace{-1.5em}
-  %\caption{
-    \center{\small{Figure 1: The hierarchy}}
-   % }
+  \caption{The hierarchy}
   \label{fig:hierarchy}
@@
 
-(* A statement in the signature of a Semiring S.  *)
+(* A statement in the signature of a `SemiRing` S.  *)
 Check forall S : SemiRing.type, forall x : S, x * 1 + 0 = x.
 
-(* We need no additional mixin to declare the structure of rings. *)
-HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A & }.
+(* We need no additional mixin to declare the `Ring` structure. *)
+HB.structure Definition Ring := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A & }.
 
@@
 For space constraints we omit the declaration
-of the green arrows in Figure 1 which are alternative interfaces
+of the green arrows in Fig.~\ref{fig:hierarchy} which are alternative interfaces
 called factories. They provide specialized interfaces to

### 2020-04-27 coqws/easychair.tex
link the github

@@
 are given in~\cite{cohen:hal-02478907}.
-In this talk we focus on how to use \HB{} in a Coq development, and we plan
-to not detail how it works behind the scenes, unless specific questions
-arise from the audience.
+In this talk we focus on how to use \HB{} in a Coq development,
+% and we plan to not detail how it works behind the scenes, 
+unless questions on its internals arise from the audience.\\
+\HB{} is available for download at \url{https://github.com/math-comp/hierarchy-builder}
 

### 2020-04-27 coqws/easychair.tex
making the code compile

@@
 \begin{coqcode}
-
 HB.structure Definition AbelianGrp := { A of AbelianGrp_of_CMonoid A & }.
@@
 Notation "1"  := one.
+Notation "-1" := (- one).
 Infix    "*"  := mul.
@@
 (* We need no additional mixin to declare the `Ring` structure. *)
-HB.structure Definition Ring := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A & }.
+HB.structure Definition Ring := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A }.
 
@@
 HB.instance Z Z_AbelianGrp.
-Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l ...
+Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l
+  Z.mul_1_r Z.mul_add_distr_r Z.mul_add_distr_l (fun _ => erefl) Zmult_0_r.
 HB.instance Z Z_SemiRing.

### 2020-04-27 coqws/easychair.tex
final
fixes

@@
 unless questions on its internals arise from the audience.\\
-\HB{} is available for download at \url{https://github.com/math-comp/hierarchy-builder}
+\HB{} is available for download at \url{https://github.com/math-comp/hierarchy-builder}.
 
@@
 Notation "1"  := one.
-Notation "-1" := (- one).
 Infix    "*"  := mul.
 
-(* A statement in the signature of a `SemiRing` S.  *)
+(* A statement in the signature of a Semiring S.  *)
 Check forall S : SemiRing.type, forall x : S, x * 1 + 0 = x.
 
-(* We need no additional mixin to declare the `Ring` structure. *)
+(* We need no additional mixin to declare the Ring structure. *)
 HB.structure Definition Ring := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A }.
@@
 (* A statement in the signature of a Ring R.  *)
-Check forall R : Ring.type, forall x : R, x * -1 = - x.
+Check forall R : Ring.type, forall x : R, x * - (1) = - x.
 
-(* We equip Z with the structures of CMonoid, AbelianGrp and SemiRing; HB equips it with Ring. *)
+(* We equip Z with CMonoid, AbelianGrp and SemiRing structures; HB equips it with Ring. *)
 Definition Z_CMonoid    := CMonoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
@@
 HB.instance Z Z_AbelianGrp.
-Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l
-  Z.mul_1_r Z.mul_add_distr_r Z.mul_add_distr_l (fun _ => erefl) Zmult_0_r.
+Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l ...
 HB.instance Z Z_SemiRing.

### 2020-04-27 coqws/easychair.tex

@@
 For space constraints we omit the declaration
-of the green arrows in Fig.~\ref{fig:hierarchy} which are alternative interfaces
+of the green arrows in Figure~\ref{fig:hierarchy} which are alternative interfaces
 called factories. They provide specialized interfaces to
-build structures for others.
+build structures from others.
 Factories are used to accommodate the insertion of
-intermediate structures in a hierarchy without breaking user code, as we all
-as to implement shortcuts: the factory to build a \coq{Ring} from \coq{Type}
+intermediate structures in a hierarchy without breaking user code, as well
+as to implement shortcuts: e.g. the factory to build a \coq{Ring} from a \coq{Type}
 (rightmost green arrow)

### 2020-04-27 hb.tex
fix overfull line
overfull hbox code + making addrC proofs look like each other
comment sentence about flat PC performance
acknowledgments

@@
   \mixinbuilders{}, so that
-  \(\provides{}(\coq{F}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)
+  \(\provides{}(\coq{F}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)\\
   It is thus necessary that
@@
 reduces the amount of code to implement inheritance significantly.
-Nested class records and chains of their projections exponentially slow down the
-conversion while the flat variant can mitigate the issue. Since we do not need
+%Nested class records and chains of their projections exponentially slow down the
+%conversion while the flat variant can mitigate the issue. 
+Since we do not need
 to care about the amount of code in automated generation, \hb{} uses the flat
@@
 
+\section*{Acknowledgments}
+
+We are particularly grateful to Georges Gonthier, Assia Mahboubi and
+Florent Hivert for the many discussions we had that helped to shape
+various parts of this work, as well as other developers of the
+Mathematical Components library and members of Dagstuhl seminar
+18341. We also thank many Coq developers and users who showed interest
+and support in this work. We thank the anonymous reviewers of this
+paper for their useful comments.
+
 \bibliography{biblio}
@@
   Proof.
+  have innerC (a b : A) : a + b + (a + b) = a + a + (b + b).
+    by rewrite -[a+b]mul1r_f -mulrDl_f mulrDr_f !mulrDl_f !mul1r_f.
   have addKl (a b c : A) : a + b = a + c -> b = c.
@@
     by move=> x; rewrite /= -addrA addrN_f addr0.
-  have innerC (a b : A) : a + b + (a + b) = a + a + (b + b).
-    by rewrite -[a+b]mul1r_f -mulrDl_f mulrDr_f !mulrDl_f !mul1r_f.
   move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
@@
   (* Builder to the bottom right mixin. *)
-  Definition to_AbelianGroup_of_Monoid := AbelianGroup_of_Monoid.Build A opp_f addrC addNr_f.
+  Definition to_AbelianGroup_of_Monoid :=
+    AbelianGroup_of_Monoid.Build A opp_f addrC addNr_f.
   HB.instance A to_AbelianGroup_of_Monoid.

### 2020-04-29 hb.tex
refactoring concl + ack

@@
 Elpi code and tree hundred lines of Coq vernacular. It took less than one month
-to implement \HB{} while it took several years of attempts and fruitful
-discussions with the other developers of the Mathematical Components library
-to design it, as well as Coq users and developers and members of Dagstuhl seminar 18341.
-In particular, discussions with Georges Gonthier, Assia Mahboubi and
-Florent Hivert were instrumental in coining the concepts formalized here.
+to implement \HB{} but it took several years of attempts and fruitful
+discussions with the people we thank below.
 
@@
 We are particularly grateful to Georges Gonthier, Assia Mahboubi and
-Florent Hivert for the many discussions we had that helped to shape
-various parts of this work, as well as other developers of the
-Mathematical Components library and members of Dagstuhl seminar
-18341. We also thank many Coq developers and users who showed interest
-and support in this work. We thank the anonymous reviewers of this
-paper for their useful comments.
+Florent Hivert for the many discussions we had that were instrumental
+in coining the concepts formalized here. We thank as well other
+developers of the Mathematical Components library and members of
+Dagstuhl seminar 18341, for the helpful discussions we had. We also
+thank many Coq developers and users who showed interest and support in
+this work. Finally, we thank the anonymous reviewers of this paper for
+their useful comments.
 

### 2020-05-04 hb.tex
Update hb.tex
Fix some typesetting issues

@@
 \newcommand{\pmp}[1]{\ensuremath{\vect{\left(p : m\ T\ p_\sigma\right)}^{#1}}}
+\newcommand{\aux}{\mathit{aux}}
 \newcommand{\hbmixin}{{\tt\color{dkgreen}HB.mixin}}
@@
 do not matter here, what is worth pointing out is that in a single
-statement we mix monoid (e.g. \coq{+}) and ring (e.g. \coq{-}) operations and in the
-proof we use monoid axioms (e.g. \coq{addrA}), ring axioms
-(e.g. \coq{addrN}) and ring lemmas (e.g. \coq{addrC}), all seamlessly.
+statement we mix monoid (e.g.~\coq{+}) and ring (e.g.~\coq{-}) operations and in the
+proof we use monoid axioms (e.g.~\coq{addrA}), ring axioms
+(e.g.~\coq{addrN}) and ring lemmas (e.g.~\coq{addrC}), all seamlessly.
 
@@
 arguments of \(m_i \in \M{}\) consist in $q_i$ of the previously quantified
-mixin parameters $p_k$, i.e. for all \(k \in \enum1{q_i}\), we
+mixin parameters $p_k$, i.e.~for all \(k \in \enum1{q_i}\), we
 have \(\sigma(i,k) \in \enum1{i-1}\).
@@
   Given \(m \in \M{}\), we define \(\dep{}(m) \in \powerset\M{}\)
-  as the set of all \mixins{} that occur as parameters of \(m\), i.e.
-  \(\dep{}(m) = \enum{m_1}{m_n}\).
+  as the set of all \mixins{} that occur as parameters of \(m\),
+  i.e.~\(\dep{}(m) = \enum{m_1}{m_n}\).
 \end{definition}
@@
   Given \(f \in \F{}\), we define \(\requires{}(f) \in \powerset\M{}\)
-  as the set of all \mixins{} that occur as parameters of \(f\), i.e.
-  \(\dep{}(f) = \enum{m_1}{m_n}\).
+  as the set of all \mixins{} that occur as parameters of \(f\),
+  i.e.~\(\dep{}(f) = \enum{m_1}{m_n}\).
 \end{definition}
@@
 \begin{proposition}[\requires{} is closed under \dep{}]
-\(\forall f \in \F{}, \dep{}^\star(\requires{}(f)) \subseteq \requires{}(f)\)
+\(\forall f \in \F{}, \dep{}^\star(\requires{}(f)) \subseteq \requires{}(f).\)
 \end{proposition}
@@
 \begin{definition}[$\provides{}$]
-  \( \provides{}(f) = \{m | \from{}(f,m) \textrm{ is defined}\}
+  \( \provides{}(f) = \{m \mid \from{}(f,m) \textrm{ is defined}\}
     \in \powerset\M{}\) the set of mixins that a factory
@@
   \(\from{}(m,m) = \left(\coq{fun T }\ \pmp{n} \ \coq{(x :}\ m\
-    \coq{T }\ p_1\ \ldots p_n\ \coq{) => x}\right) \in \B{}.\)
+    \coq{T }\ p_1\ \ldots p_n\coq{) => x}\right) \in \B{}.\)
 \end{proof}
@@
   We call \(\cdef{}(c) \in \powerset\M{}\) the set of \mixins{}
-  mentioned in the fields of the \class{}, i.e.
-  \(\{m_1, \ldots,m_n\}\).  Given that \class{} records are well typed
+  mentioned in the fields of the \class{}, i.e.~\(\{m_1, \ldots,m_n\}\).
+  Given that \class{} records are well typed
   in the empty context the set of \mixin{} records is closed
@@
   records contains the same set of \mixins{} (disregarding the order of
-  the fields), i.e. \cdef{} is injective.
+  the fields), i.e.~\cdef{} is injective.
 \end{definition}
@@
 \begin{proposition}[\( \C{} \subseteq \F{} \)]
-  forall \(m \in \cdef{}(c)\) we have \(\from{}(c,m_i) = p_i,\) and it
+  For all \(m \in \cdef{}(c)\) we have \(\from{}(c,m_i) = p_i,\) and it
   follows that \(\requires{}(c)=\emptyset \),
@@
 \phantterms{} for the user to replace uses of constructors of \factories{}.
-These commands first create a record \(f_{aux}\) with a
-constructor \(F_{aux}\) and then create abbreviations \(f\) and \(F\)
-that automatically fill the mixin arguments of \(f_{aux}\) and \(F_{aux}\)
+These commands first create a record \(f_{\aux}\) with a
+constructor \(F_{\aux}\) and then create abbreviations \(f\) and \(F\)
+that automatically fill the mixin arguments of \(f_{\aux}\) and \(F_{\aux}\)
 respectively.
@@
 \begin{coqcode}
-Record #\(f_{aux}\)# T #\(\pmp{n}\)# := #\(F_{aux}\)# { ... }.
-Notation #$f$# T := (#\(f_{aux}\)# T ...(* $p_1 \ldots p_n$ inferred when $T$ is known *)).
-Notation #$F$# T  #\(x_1\ \ldots\ x_k\)# := (#\(F_{aux}\)# T ...(* $p_1 \ldots p_n$ inferred when $T$ is known *) #\(x_1\ \ldots\ x_k\)#).
-Definition b : #$f$# T := #$F$# T #\(x_1\ \ldots\ x_n\)#
+Record #\(f_{\aux}\)# T #\(\pmp{n}\)# := #\(F_{\aux}\)# { ... }.
+Notation #$f$# T := (#\(f_{\aux}\)# T ...(* $p_1 \ldots p_n$ inferred when $T$ is known *)).
+Notation #$F$# T  #\(x_1\ \ldots\ x_k\)# := (#\(F_{\aux}\)# T ...(* $p_1 \ldots p_n$ inferred when $T$ is known *) #\(x_1\ \ldots\ x_k\)#).
+Definition b : #$f$# T := #$F$# T #\(x_1\ \ldots\ x_n\)#.
 \end{coqcode}
@@
 Elpi code and tree hundred lines of Coq vernacular. It took less than one month
-to implement \HB{} but it took several years of attempts and fruitful
+to implement \HB{}, but its design took several years of attempts and fruitful
 discussions with the people we thank below.

### 2020-05-04 hb.tex
remove _f from factory ops

--- 2020-04-29 hb.tex
@@
 
-  Fact mul0r : left_zero zero mul_f.
+  Fact mul0r : left_zero zero mul.
   Proof.
   move=> x; rewrite -[LHS]add0r addrC.
-  rewrite -{2}(addNr (mul_f x x)) (addrC (opp _)) addrA.
-  by rewrite -mulrDl_f add0r addrC addNr.
+  rewrite -{2}(addNr (mul x x)) (addrC (opp _)) addrA.
+  by rewrite -mulrDl add0r addrC addNr.
   Qed.
 
-  Fact mulr0 : right_zero zero mul_f.
+  Fact mulr0 : right_zero zero mul.
   Proof.
   move=> x; rewrite -[LHS]add0r addrC.
-  rewrite -{2}(addNr (mul_f x x)) (addrC (opp _)) addrA.
-  by rewrite -mulrDr_f add0r addrC addNr.
+  rewrite -{2}(addNr (mul x x)) (addrC (opp _)) addrA.
+  by rewrite -mulrDr add0r addrC addNr.
   Qed.
 
-  Definition to_SemiRing_of_Monoid := SemiRing_of_Monoid.Build A _ mul_f mulrA_f
-     mul1r_f mulr1_f mulrDl_f mulrDr_f mul0r mulr0.
+  Definition to_SemiRing_of_Monoid := SemiRing_of_Monoid.Build A _ mul mulrA
+     mul1r mulr1 mulrDl mulrDr mul0r mulr0.
   HB.instance A to_SemiRing_of_Monoid.
@@
   have innerC (a b : A) : a + b + (a + b) = a + a + (b + b).
-    by rewrite -[a+b]mul1r_f -mulrDl_f mulrDr_f !mulrDl_f !mul1r_f.
+    by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
   have addKl (a b c : A) : a + b = a + c -> b = c.
-    apply: can_inj (add a) (add (opp_f a)) _ _ _.
-    by move=> x; rewrite addrA addNr_f add0r.
+    apply: can_inj (add a) (add (opp a)) _ _ _.
+    by move=> x; rewrite addrA addNr add0r.
   have addKr (a b c : A) : b + a = c + a -> b = c.
-    apply: can_inj (add ^~ a) (add ^~ (opp_f a)) _ _ _.
-    by move=> x; rewrite /= -addrA addrN_f addr0.
+    apply: can_inj (add ^~ a) (add ^~ (opp a)) _ _ _.
+    by move=> x; rewrite /= -addrA addrN addr0.
   move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
@@
   Definition to_AbelianGroup_of_Monoid :=
-    AbelianGroup_of_Monoid.Build A opp_f addrC addNr_f.
+    AbelianGroup_of_Monoid.Build A opp addrC addNr.
   HB.instance A to_AbelianGroup_of_Monoid.
@@
   (* Builder to the top right factory, which is compiled to the bottom left mixin. *)
-  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build A one_f mul_f
-    mulrA_f mul1r_f mulr1_f mulrDl_f mulrDr_f.
+  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build A one mul
+    mulrA mul1r mulr1 mulrDl mulrDr.
   HB.instance A to_Ring_of_AbelianGroup.

### 2020-05-04 hb.tex
fix category/relatedversion
Renaming A -> R, switching to code that compiles as such.

@@
 
-\category{}%optional, e.g. invited paper
-
-\relatedversion{System Description}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+\category{System Description}%optional, e.g. invited paper
+
+%\relatedversion{System Description}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
 %\relatedversion{A full version of the paper is available at \url{...}.}
@@
 
-Lemma addrC : forall R : Ring.type, commutative (@add R).                       #\label{demo:theory:state:addrC}#
+Lemma addrC {R : Ring.type} : commutative (@add R).                       #\label{demo:theory:state:addrC}#
 Proof. (* Proof by Hankel 1867, in Appendix B *) Qed.
@@
 Definition Z_Monoid_axioms : Monoid_of_Type Z :=                         #\label{demo:theory:z:monoid:axioms}#
-  Monoid_of_Type.Build Z 0 Z.add Z.add_assoc Z.add_0_l Z.add_0_r.
+  Monoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_0_l Z.add_0_r.
 
@@
 Definition Z_Ring_axioms : Ring_of_Monoid Z :=                        #\label{demo:theory:z:ring:axioms}#
-  Ring_of_Monoid.Build Z 1 Z.opp Z.mul
+  Ring_of_Monoid.Build Z 1%Z Z.opp Z.mul
     Z.add_opp_diag_l Z.add_opp_diag_r Z.mul_assoc Z.mul_1_l Z.mul_1_r
@@
 Proof.
-have innerC (a b : R) : (a + b) + (a + b) = (a + a) + (b + b).
+have innerC (a b : R) : a + b + (a + b) = a + a + (b + b).
   by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
 have addKl (a b c : R) : a + b = a + c -> b = c.
-  apply: can_inj (add a) (add (-a)) _ _ _.
+  apply: can_inj (add a) (add (opp a)) _ _ _.
   by move=> x; rewrite addrA addNr add0r.
 have addKr (a b c : R) : b + a = c + a -> b = c.
-  apply: can_inj (add ^~ a) (add ^~ (-a)) _ _ _.
+  apply: can_inj (add ^~ a) (add ^~ (opp a)) _ _ _.
   by move=> x; rewrite /= -addrA addrN addr0.
@@
 
-HB.builders Context (A : Type) (f : Ring_of_Monoid A).
-
-  Lemma addrC : commutative (add : A -> A -> A).
-  Proof.
-  have innerC (a b : A) : a + b + (a + b) = a + a + (b + b).
+HB.builders Context (R : Type) (f : Ring_of_Monoid R).
+
+  Lemma addrC : commutative (add : R -> R -> R).
+  Proof. (* Exactly the same as in Appendix B. *)
+  have innerC (a b : R) : a + b + (a + b) = a + a + (b + b).
     by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
-  have addKl (a b c : A) : a + b = a + c -> b = c.
+  have addKl (a b c : R) : a + b = a + c -> b = c.
     apply: can_inj (add a) (add (opp a)) _ _ _.
     by move=> x; rewrite addrA addNr add0r.
-  have addKr (a b c : A) : b + a = c + a -> b = c.
+  have addKr (a b c : R) : b + a = c + a -> b = c.
     apply: can_inj (add ^~ a) (add ^~ (opp a)) _ _ _.
@@
   Definition to_AbelianGroup_of_Monoid :=
-    AbelianGroup_of_Monoid.Build A opp addrC addNr.
-  HB.instance A to_AbelianGroup_of_Monoid.
+    AbelianGroup_of_Monoid.Build R opp addrC addNr.
+  HB.instance R to_AbelianGroup_of_Monoid.
 
   (* Builder to the top right factory, which is compiled to the bottom left mixin. *)
-  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build A one mul
+  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build R one mul
     mulrA mul1r mulr1 mulrDl mulrDr.
-  HB.instance A to_Ring_of_AbelianGroup.
+  HB.instance R to_Ring_of_AbelianGroup.
 

### 2020-05-04 hb.tex
Consistent variable names

@@
 \begin{coqcode}
-#\hbmixin{}# Record Monoid_of_Type A := {                       #\label{demo:mixin:mot}#
-  zero : A;
-  add : A -> A -> A;
+#\hbmixin{}# Record Monoid_of_Type M := {                       #\label{demo:mixin:mot}#
+  zero : M;
+  add : M -> M -> M;
   addrA : associative add;             (* `add` is associative.                   *)
@@
 }.                                                             #\label{demo:mixin:mot:end}#
-#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type A }.        #\label{demo:structure:monoid}#
-
-#\hbmixin{}# Record Ring_of_Monoid A of Monoid A := {       #\label{demo:mixin:rom}#
-  one : A;
-  opp : A -> A;
-  mul : A -> A -> A;
+#\hbstructure{}# Definition Monoid := { M of Monoid_of_Type M }.        #\label{demo:structure:monoid}#
+
+#\hbmixin{}# Record Ring_of_Monoid R of Monoid R := {       #\label{demo:mixin:rom}#
+  one : R;
+  opp : R -> R;
+  mul : R -> R -> R;
   addNr : left_inverse zero opp add;   (* `opp x` is the left and right additive  *)
@@
 }.
-#\hbstructure{}# Definition Ring := { A of Monoid A & Ring_of_Monoid A }. #\label{demo:structure:ring}#
+#\hbstructure{}# Definition Ring := { R of Monoid R & Ring_of_Monoid R }. #\label{demo:structure:ring}#
 
@@
 and~\ref{demo:mixin:mot:end} declares a \mixin{} that can turn a naked
-type \coq{A} into a monoid, hence we chose the name to be \coq{Monoid_of_Type}.
-
-The \hbstructure{} command takes in input a definition for a sigma type S
-that equips \coq{A} with a list of \factories{}.
-It registers a structure S in the hierarchy placing
+type \coq{M} into a monoid, hence we chose the name to be \coq{Monoid_of_Type}.
+
+The \hbstructure{} command takes in input a definition for a sigma type \coq{S}
+that equips a type with a list of \factories{}.
+It registers a structure \coq{S} in the hierarchy placing
 any definition specific to that structure inside a Coq module named \coq{S}.
@@
 and properties that are needed in order to enrich a monoid to a ring, hence
-the name \coq{Ring_of_Monoid}. Indeed this time the type variable \coq{A}
+the name \coq{Ring_of_Monoid}. Indeed this time the type variable \coq{R}
 is followed by \coq{Monoid} that enriches
-\coq{A} with the operations and properties of monoids. As a consequence
+\coq{R} with the operations and properties of monoids. As a consequence
 \coq{add} and \coq{zero} can be used to express the new properties.
@@
 In particular they are quantified over a \coq{Monoid.type} or \coq{Ring.type},
-and not a simple type \coq{A} as in the \mixins{}. Moreover we evince that
+and not a simple type as in the \mixins{}. Moreover we evince that
 the \coq{sort} projection is declared as an implicit coercion~\cite{Saibi97}
@@
 \begin{coqcode}
-#\hbmixin{}# Record Monoid_of_Type A := { ... (* unchanged *) ... }.
-#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type A }.
+#\hbmixin{}# Record Monoid_of_Type M := { ... (* unchanged *) ... }.
+#\hbstructure{}# Definition Monoid := { M of Monoid_of_Type M }.
 
@@
 
-#\hbmixin{}# Record Ring_of_AbelianGroup A of AbelianGroup A := {
-  one : A;
-  mul : A -> A -> A;
+#\hbmixin{}# Record Ring_of_AbelianGroup R of AbelianGroup R := {
+  one : R;
+  mul : R -> R -> R;
   mulrA : associative mul;
@@
 }.
-#\hbstructure{}# Definition Ring := { A of AbelianGroup A & Ring_of_AbelianGroup A }.
-
-Lemma addrN {R : AbelianGroup.type} : right_inverse zero opp add.               #\label{demo2:proof:addrN}#
-Proof. by move=>x; rewrite addrC addNr. Qed.
+#\hbstructure{}# Definition Ring := { R of AbelianGroup R & Ring_of_AbelianGroup R }.
+
+Lemma addrN {A : AbelianGroup.type} : right_inverse zero opp add.               #\label{demo2:proof:addrN}#
+Proof. by move=> x; rewrite addrC addNr. Qed.
 \end{coqcode}
@@
 for rings is now called \coq{Ring_of_AbelianGroup} (instead of \coq{Ring_of_Monoid})
-since it expects the type \coq{A} to be already an Abelian group and hence
+since it expects the type \coq{R} to be already an Abelian group and hence
 gathers fewer axioms.
@@
 \begin{coqcode}
-#\hbfactory{}# Record Ring_of_Monoid A of Monoid A := { ... (* unchanged *) ...}.
-
-#\hbbuilders{}# Context A (f : Ring_of_Monoid A).                          #\label{demo3:variable:f}#
+#\hbfactory{}# Record Ring_of_Monoid R of Monoid R := { ... (* unchanged *) ...}.
+
+#\hbbuilders{}# Context R (f : Ring_of_Monoid R).                          #\label{demo3:variable:f}#
 
@@
   Definition to_AbelianGroup_of_Monoid :=
-    AbelianGroup_of_Monoid.Build A (opp f) addrC (addNr f). (* addrN unused *)
-  #\hbinstance{}# A to_AbelianGroup_of_Monoid.
-
-  Definition to_Ring_of_AbelianGroup :=  Ring_of_AbelianGroup.Build A
-    (one f) (mul f) (mulrA f) (mul1r f) (mulr1 f) (mulrDl f) (mulrDr f).
-  #\hbinstance{}# A to_Ring_of_AbelianGroup.
+    AbelianGroup_of_Monoid.Build R opp addrC addNr. (* addrN unused *)
+  #\hbinstance{}# R to_AbelianGroup_of_Monoid.
+
+  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build R one mul
+    mulrA mul1r mulr1 mulrDl mulrDr.
+  #\hbinstance{}# R to_Ring_of_AbelianGroup.
 
@@
 axioms (named \coq{f}, line~\ref{demo3:variable:f}).
-This construction is also registered as canonical for \coq{A},
+This construction is also registered as canonical for \coq{R},
 so that the next construction \coq{to_Ring_of_AbelianGroup} can call
-the \coq{Ring_of_AbelianGroup.Build} \phantterm{} that requires \coq{A} to be
+the \coq{Ring_of_AbelianGroup.Build} \phantterm{} that requires \coq{R} to be
 an \coq{AbelianGroup}.
@@
 \begin{coqcode}
-#\hbstructure{}# Definition Ring := { A of Monoid A & Ring_of_Monoid A }.
-#\hbstructure{}# Definition Ring := { A of AbelianGroup A & Ring_of_AbelianGroup A }.
+#\hbstructure{}# Definition Ring := { R of Monoid R & Ring_of_Monoid R }.
+#\hbstructure{}# Definition Ring := { R of AbelianGroup R & Ring_of_AbelianGroup R }.
 \end{coqcode}
@@
 Module Monoid.
-Record axioms (T : Type) : Type :=
-  Class { Monoid_of_Type_mixin : Monoid_of_Type T; }.
+Record axioms (M : Type) : Type :=
+  Class { Monoid_of_Type_mixin : Monoid_of_Type M; }.
 Structure type : Type := Pack { sort : Type; class : axioms sort; }.
@@
 
-Definition zero {T : Monoid.type} : T :=
-  Monoid_of_Type.zero T (Monoid.Monoid_of_Type_mixin T (Monoid.class T)).
-Definition add {T : Monoid.type} : T -> T -> T :=
-  Monoid_of_Type.add T (Monoid.Monoid_of_Type_mixin T (Monoid.class T)).
-Lemma addrA {T : Monoid.type} : associative (@add T).
+Definition zero {M : Monoid.type} : M :=
+  Monoid_of_Type.zero M (Monoid.Monoid_of_Type_mixin M (Monoid.class M)).
+Definition add {M : Monoid.type} : M -> M -> M :=
+  Monoid_of_Type.add M (Monoid.Monoid_of_Type_mixin M (Monoid.class M)).
+Lemma addrA {M : Monoid.type} : associative (@add M).
 (* Two monoid axioms `add0r` and `addr0` are omitted. *)
@@
 Module AbelianGroup.
-Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin : Monoid_of_Type T;
-  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
+Record axioms (A : Type) : Type := Class {
+  Monoid_of_Type_mixin : Monoid_of_Type A;
+  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid A Monoid_of_Type_mixin; }.
 Structure type : Type := Pack { sort : Type; class : axioms sort }.
@@
 Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
-Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup T) : #\label{demo4:coe:ag:monoid:class}#
-  Monoid T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
-Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
-  Monoid.Pack (AbelianGroup.sort T) (AbelianGroup.class T).
-
-Definition opp {T : AbelianGroup.type} : T -> T := ... .
+Coercion AbelianGroup_class_to_Monoid_class (A : Type) (c : AbelianGroup A) : #\label{demo4:coe:ag:monoid:class}#
+  Monoid A := Monoid.Class A (AbelianGroup.Monoid_of_Type_mixin A c).
+Coercion AbelianGroup_to_Monoid (A : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
+  Monoid.Pack (AbelianGroup.sort A) (AbelianGroup.class A).
+
+Definition opp {A : AbelianGroup.type} : A -> A := ... .
 (* Two Abelian group axioms `addrC` and `addNr` are omitted. *)
@@
 
-Generally, a coercion from a structure A to another structure B can (and should) have the following form, thanks to the corresponding coercion between classes \coq{A_class_to_B_class}.
-\begin{coqcode}
-Coercion A_to_B (T : A.type) : B.type :=
-  B.Pack (A.sort T) ((* A_class_to_B_class _ *) (A.class T)).
+Generally, a coercion from a structure X to another structure Y can (and should) have the following form, thanks to the corresponding coercion between classes \coq{X_class_to_Y_class}.
+\begin{coqcode}
+Coercion X_to_Y (T : X.type) : Y.type :=
+  Y.Pack (X.sort T) ((* X_class_to_Y_class _ *) (X.class T)).
 \end{coqcode}
@@
 \begin{coqcode}
-#\hbmixin{}# Record SemiRing_of_Monoid A
-  of Monoid A := {
-  one : A;
-  mul : A -> A -> A;
+#\hbmixin{}# Record SemiRing_of_Monoid S
+  of Monoid S := {
+  one : S;
+  mul : S -> S -> S;
   (* 7 axioms are omitted. *)
@@
 
-#\hbfactory{}# Record Ring_of_AbelianGroup A
-  of AbelianGroup A := {
+#\hbfactory{}# Record Ring_of_AbelianGroup R
+  of AbelianGroup R := {
   (* 2 operators and 5 axioms are omitted. *)
 }.
-#\hbbuilders{}# Context A (a : Ring_of_AbelianGroup A).
+#\hbbuilders{}# Context R (f : Ring_of_AbelianGroup R).
 ..
@@
 Module SemiRing.
-Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin : Monoid_of_Type T;
-  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin;
+Record axioms (S : Type) : Type := Class {
+  Monoid_of_Type_mixin : Monoid_of_Type S;
+  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid S Monoid_of_Type_mixin;
 }.
@@
 
-Definition one {T : SemiRing.type} : T := ... .
-Definition mul {T : SemiRing.type} : T -> T -> T := ... .
+Definition one {S : SemiRing.type} : S := ... .
+Definition mul {S : SemiRing.type} : S -> S -> S := ... .
 (* 7 semiring axioms are omitted. *)
@@
 Module Ring.
-Record axioms (T : Type) : Type := Class {
-  Monoid_of_Type_mixin : Monoid_of_Type T;
-  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin;
-  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
+Record axioms (R : Type) : Type := Class {
+  Monoid_of_Type_mixin : Monoid_of_Type R;
+  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid R Monoid_of_Type_mixin;
+  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid R Monoid_of_Type_mixin; }.
 Structure type : Type := Pack { sort : Type; class : axioms sort; }.
@@
 \begin{coqcode}
-Canonical AbelianGroup_to_SemiRing (T : Ring.type) :=
-  SemiRing.Pack (AbelianGroup.sort (Ring_to_AbelianGroup T)) (Ring.class T)).
+Canonical AbelianGroup_to_SemiRing (S : Ring.type) :=
+  SemiRing.Pack (AbelianGroup.sort (Ring_to_AbelianGroup S)) (Ring.class S)).
 \end{coqcode}
@@
 (* Bottom mixin in Fig. 2. *)
-HB.mixin Record Monoid_of_Type A := {
-  zero : A;
-  add : A -> A -> A;
+HB.mixin Record Monoid_of_Type M := {
+  zero : M;
+  add : M -> M -> M;
   addrA : associative add;
@@
 }.
-HB.structure Definition Monoid := { A of Monoid_of_Type A }.
+HB.structure Definition Monoid := { M of Monoid_of_Type M }.
 Notation "0" := zero : hb_scope.
@@
 (* Bottom left mixin in Fig. 2. *)
-HB.mixin Record SemiRing_of_Monoid A of Monoid A := {
-  one : A;
-  mul : A -> A -> A;
+HB.mixin Record SemiRing_of_Monoid S of Monoid S := {
+  one : S;
+  mul : S -> S -> S;
   mulrA : associative mul;
@@
 }.
-HB.structure Definition SemiRing := { A of Monoid A & SemiRing_of_Monoid A }.
+HB.structure Definition SemiRing := { S of Monoid S & SemiRing_of_Monoid S }.
 Notation "1" := one : hb_scope.
@@
 
-Lemma addrN {R : AbelianGroup.type} : right_inverse (zero : R) opp add.
+Lemma addrN {A : AbelianGroup.type} : right_inverse (zero : A) opp add.
 Proof. by move=> x; rewrite addrC addNr. Qed.
@@
 (* Top right factory in Fig. 2. *)
-HB.factory Record Ring_of_AbelianGroup A of AbelianGroup A := {
-  one : A;
-  mul : A -> A -> A;
+HB.factory Record Ring_of_AbelianGroup R of AbelianGroup R := {
+  one : R;
+  mul : R -> R -> R;
   mulrA : associative mul;
@@
 (* It is an exact copy of the bottom right mixin. *)
-HB.factory Definition Ring_of_SemiRing A of SemiRing A := AbelianGroup_of_Monoid A.
+HB.factory Definition Ring_of_SemiRing R of SemiRing R := AbelianGroup_of_Monoid R.
 (* The corresponding builder is the identity. *)
-HB.builders Context (A : Type) (f : Ring_of_SemiRing A).
-  Definition to_AbelianGroup_of_Monoid : AbelianGroup_of_Monoid A := f.
-  HB.instance A to_AbelianGroup_of_Monoid.
+HB.builders Context (R : Type) (f : Ring_of_SemiRing R).
+  Definition to_AbelianGroup_of_Monoid : AbelianGroup_of_Monoid R := f.
+  HB.instance R to_AbelianGroup_of_Monoid.
 HB.end.
@@
 (* Right-most factory in Fig. 2. *)
-HB.factory Record Ring_of_Monoid A of Monoid A := {
-  one : A;
-  opp : A -> A;
-  mul : A -> A -> A;
+HB.factory Record Ring_of_Monoid R of Monoid R := {
+  one : R;
+  opp : R -> R;
+  mul : R -> R -> R;
   addNr : left_inverse zero opp add;

### 2020-06-19 coqws/easychair.tex
Improving the presentation.

@@
 %\documentclass[letterpaper]{easychair}
+\usepackage{amsmath, amssymb}
+
 \usepackage{wrapfig}
@@
 \newmintinline[elpi]{elpi.py:ElpiLexer -x}{fontsize=\small}
-\newminted[coqcode]{elpi.py:CoqElpiLexer -x}{firstnumber=last,linenos=true,fontsize=\footnotesize,mathescape,escapeinside=\#\#}
+\newminted[coqcode]{elpi.py:CoqElpiLexer -x}{firstnumber=last,linenos=false,fontsize=\footnotesize,mathescape,escapeinside=\#\#}
 \newmintinline[coq]{elpi.py:CoqElpiLexer -x}{fontsize=\small}
@@
 
-There is no questioning the importance of abstract algebra in the mathematical
-vernacular. And it is with no surprise that libraries of machine checked code
-are, nowadays, organized around hierarchies of structures. Unfortunately the
-language of Type Theory and the features provided by the Coq system make the
-construction of a hierarchy hard even for expert users. The difficulty
-begins with the non-orthogonal choices of the encoding style, from bundled to
-unbundled, and inference mechanism, from Type Classes to Canonical Structures.
-To this, one may add the concerns about performance and about
-the usability, by a non expert, of the final hierarchy.
+Abstract algebra is extremely important in the mathematical vernacular
+and libraries of machine checked code are, nowadays, organized around
+hierarchies of structures. Unfortunately the language of Type Theory
+and the features provided by the Coq system make the construction of a
+hierarchy hard even for expert users. The difficulty begins with the
+non-orthogonal choices, between storing information as record fields
+or parameters, and between using type classes and canonical structures
+for inference. To this, one may add the concerns about performance and
+about the usability, by a non expert, of the final hierarchy.
 
-In this talk we give an extensive demo of the Hierarchy Builder (\HB{})
-extension for the Coq system version 8.10 and above.
-\HB{} gives the library designer a language to describe the building blocks
-of algebraic structures and to assemble them into a hierarchy. Similarly it
-provides the final user linguistic constructs to build instances (examples) of
-structures and to teach the elaborator of Coq how to take advantage of this
-knowledge during type inference. Finally \HB{} lets the library designer
-improve the usability of his library by providing alternative interfaces
-to the primitive ones, a feature that can also be used to accommodate changes
-to the hierarchy without breaking user code.
+\HB{} gives the library designer a language to describe the building
+blocks of algebraic structures and to assemble them into a
+hierarchy. Similarly it provides the final user linguistic constructs
+to build instances (examples) of structures and to teach the
+elaborator of Coq how to take advantage of this knowledge during type
+inference. Finally \HB{} lets the library designer improve the
+usability of his library by providing alternative interfaces to the
+primitive ones, a feature that can also be used to accommodate changes
+to the hierarchy without breaking user code. The details of \HB{},
+including its compilation to a variant of the Packed Classes
+discipline and its implementation in the Elpi extension language for
+Coq, are given in~\cite{cohen:hal-02478907}.
 
-The details of \HB{}, including its compilation to the (Flat) Packed Classes
-discipline and its implementation in the Elpi extension language for Coq,
-are given in~\cite{cohen:hal-02478907}.
-In this talk we focus on how to use \HB{} in a Coq development,
-% and we plan to not detail how it works behind the scenes, 
-unless questions on its internals arise from the audience.\\
+In this talk we focus on how to use \HB{} in a Coq development without
+any prior knowledge, walking slowly through the example below,
+describing the bricks of the language from mixins -- i.e. bare bone building
+block of the hierarchy, it packs operations and axioms -- to structures.
+
 \HB{} is available for download at \url{https://github.com/math-comp/hierarchy-builder}.
@@
 
+We first introduce the base structure of our example,
+\verb-CMonoid-, which is defined in two phases: first its contents
+using \verb-HB.mixin- and then the structure itself is declared using
+\verb-HB.structure.-
+
 \begin{coqcode}
 HB.mixin Record CMonoid_of_Type A := { (* The set of axioms making A a commutative monoid. *)
-  zero  : A;
-  add   : A -> A -> A;
+  zero  : A; add   : A -> A -> A;
   addrA : associative add;  (* `add` is associative  *)
   addrC : commutative add;  (* `add` is commutative  *)
-  add0r : left_id zero add; (* `zero` is the neutral *)
+  add0r : left_id zero add; (* `zero` is a neutral element *)
 }.
@@
 (* The type of the operations and axioms depend on a CMonoid.type structure. *)
-Check addrC. (* ?M : CMonoid.type |- forall x y : ?M, x + y = y + x *)
-
-HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
-  (* We can write `add` here since A is a  CMonoid   *)
-  opp   : A -> A;
-  addNr : left_inverse zero opp add; (* `opp` is the additive inverse *)
-}.
+Check addrC. (* ?M : CMonoid.type |- commutative (@add ?M) *)
 \end{coqcode}
 
-\newpage
+We extend this hierarchy with abelian group, semi ring and ring
+structures, which happen to form a diamond (cf
+Figure~\ref{fig:hierarchy}), using only the bare minimum information:
+we use the same pattern as before to create the abelian group and
+semi-ring structures, then a ring being no more than the combination
+of the former structures, we do not even need to provide a mixin.
 
 \begin{wrapfigure}[0]{r}{.35\textwidth}
-  \vspace{1em}
+  \vspace{-27em}
   \includegraphics[width=.35\textwidth]{diagram.pdf}
@@
 \end{wrapfigure}
-\
+
 \begin{coqcode}
-HB.structure Definition AbelianGrp := { A of AbelianGrp_of_CMonoid A & }.
+HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
+  opp   : A -> A;
+  (* We can write `add` here since A is a  CMonoid   *)
+  addNr : left_inverse zero opp add; (* `opp` is the additive inverse *)
+}.
+HB.structure Definition AbelianGrp := { A of AbelianGrp_of_CMonoid A }.
 Notation "- x"   := (opp x).
 Notation "x - y" := (add x (opp y)).
-
-(* A statement in the signature of an Abelian group G. *)
-Check forall G : AbelianGrp.type, forall x : G, x - x = 0.
 
@@
 }.
-HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & }.
+HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A }.
 Notation "1"  := one.
@@
 
-(* A statement in the signature of a Semiring S.  *)
-Check forall S : SemiRing.type, forall x : S, x * 1 + 0 = x.
-
 (* We need no additional mixin to declare the Ring structure. *)
 HB.structure Definition Ring := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A }.
+\end{coqcode}
 
-(* A statement in the signature of a Ring R.  *)
-Check forall R : Ring.type, forall x : R, x * - (1) = - x.
+Remark we can write abstract statements combining successive signatures:
+\begin{coqcode}
+(* An example statement in the signature of an Abelian group G, combining 0 and -. *)
+Check forall G : AbelianGrp.type, forall x : G, x - x = 0.
+(* An example statement in the signature of a Semiring S, combining 0, +, and *.  *)
+Check forall S : SemiRing.type, forall x : S, x * 1 + 0 = x.
+(* An example statement in the signature of a Ring R, combining -, 1 and *.  *)
+Check forall R : Ring.type, forall x y : R, x * - (1 * y) = - x * y.
+\end{coqcode}
 
-(* We equip Z with CMonoid, AbelianGrp and SemiRing structures; HB equips it with Ring. *)
-Definition Z_CMonoid    := CMonoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
-HB.instance Z Z_CMonoid.
-Definition Z_AbelianGrp := AbelianGrp_of_CMonoid.Build Z Z.opp Z.add_opp_diag_l.
-HB.instance Z Z_AbelianGrp.
-Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l ...
-HB.instance Z Z_SemiRing.
+Now, we can equip $\mathbb{Z}$ with commutative monoid, abelian group and
+semi ring instances, and \HB{} equips it automatically with the ring
+one, because there is only one possible choice.
 
-(* A statement in the signature of the Z ring *)
+\begin{coqcode}
+HB.instance Definition Z_CMonoid    := CMonoid_of_Type.Build Z 0%Z Z.add
+  Z.add_assoc Z.add_comm Z.add_0_l.
+HB.instance Definition Z_AbelianGrp := AbelianGrp_of_CMonoid.Build Z Z.opp Z.add_opp_diag_l.
+HB.instance Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul
+  Z.mul_assoc Z.mul_1_l Z.mul_1_r Z.mul_add_distr_r Z.mul_add_distr_l Z.mul_0_l Z.mul_0_r.
+
+(* An example statement in the signature of the Z ring, combining Z, 0, +, -, 1 and * *)
 Check forall x : Z, x * - (1 + x) = 0 + 1.
@@
 
-For space constraints we omit the declaration
-of the green arrows in Figure~\ref{fig:hierarchy} which are alternative interfaces
-called factories. They provide specialized interfaces to
-build structures from others.
-Factories are used to accommodate the insertion of
-intermediate structures in a hierarchy without breaking user code, as well
-as to implement shortcuts: e.g. the factory to build a \coq{Ring} from a \coq{Type}
-(rightmost green arrow)
-can omit the commutativity axiom of the addition, since it is implied
-by the other axioms~\cite{nearrings}.
+For space constraints we omit the declaration of the dashed arrows in
+Figure~\ref{fig:hierarchy}, which are Factories, i.e. alternative
+ways of building the same structures, but from different inputs.
+
+Factories are used to accommodate the insertion of intermediate
+structures in a hierarchy without breaking user code, as well as to
+implement shortcuts: e.g. the factory to build a \coq{Ring} from a
+\coq{Type} (rightmost dashed arrow) can omit the commutativity axiom
+of the addition, since it is implied by the other axioms.
+
 \bibliography{biblio}

### 2020-06-19 coqws/easychair.tex
Better sectioning

@@
 
-Remark we can write abstract statements combining successive signatures:
+\paragraph{Usage} We can write abstract statements combining successive signatures:
 \begin{coqcode}
@@
 
-Now, we can equip $\mathbb{Z}$ with commutative monoid, abelian group and
-semi ring instances, and \HB{} equips it automatically with the ring
-one, because there is only one possible choice.
+We can also equip $\mathbb{Z}$ with commutative monoid, abelian group
+and semi ring instances, and \HB{} equips it automatically with a ring
+instance, because there is only one possible choice.
 
@@
 
+\paragraph{Factories}
 For space constraints we omit the declaration of the dashed arrows in
 Figure~\ref{fig:hierarchy}, which are Factories, i.e. alternative
-ways of building the same structures, but from different inputs.
+ways of building the same structures from different inputs.
 
-Factories are used to accommodate the insertion of intermediate
+Factories are used to allow the insertion of intermediate
 structures in a hierarchy without breaking user code, as well as to
 implement shortcuts: e.g. the factory to build a \coq{Ring} from a
-\coq{Type} (rightmost dashed arrow) can omit the commutativity axiom
+\coq{Type} can omit the commutativity axiom
 of the addition, since it is implied by the other axioms.

### 2020-06-19 coqws/easychair.tex
small fix
being a bit more verbose about CMonoid

@@
 describing the bricks of the language from mixins -- i.e. bare bone building
-block of the hierarchy, it packs operations and axioms -- to structures.
+block of the hierarchy, packing operations and axioms -- to structures.
 
@@
 
-We first introduce the base structure of our example,
-\verb-CMonoid-, which is defined in two phases: first its contents
-using \verb-HB.mixin- and then the structure itself is declared using
-\verb-HB.structure.-
+We first introduce the base structure of our example, the structure
+\verb-CMonoid- of commutative monoids, which is defined in two phases:
+first its contents using \verb-HB.mixin- and then the structure itself
+is declared using \verb-HB.structure.-
 

### 2020-06-26 talk-fscd/main.tex
Talk fscd intial import

\documentclass[11pt]{beamer}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{sourcesanspro}
\usepackage{sourcecodepro}
\usepackage{euler}
\usepackage[normalem]{ulem}
\usepackage{natbib}
\usepackage{apalike}
% \usepackage{bibentry}
% \usepackage{enumitem}
\usepackage[font={scriptsize},labelfont={scriptsize}]{caption}
\usepackage{lmodern,tgtermes,tabularx,ragged2e}
% \newcolumntype{Y}{>{\arraybackslash\RaggedRight}X}
% \newcolumntype{P}[1]{>{\arraybackslash\RaggedRight}p{#1}}
\usepackage{hhline}
\usepackage{booktabs}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{makecell}
\usepackage{pdfpages}
% \usepackage[overlay,absolute]{textpos}
%   \setlength{\TPHorizModule}{1mm}
%   \setlength{\TPVertModule}{1mm}
\usepackage{graphicx}

% \nobibliography*
\setcitestyle{authoryear,comma,square,aysep={},yysep={}}
% \setbeameroption{hide notes}
% \setbeameroption{notes on second screen=right}
% \setbeameroption{use timer}
% \setbeameroption{use timer countdown=30}

\usetheme{inria}
%\usepackage{helvet}
\AtBeginSection[]{
  \begin{frame}[plain]
    \partpage
  \end{frame}
}
\newcommand{\inriaswitchcolors}[1]{%
\pgfaliasimage{figfootline}{figfootline-#1}% !!!
\pgfaliasimage{figbackground}{figbackground-#1}% !!!
\pgfaliasimage{figbackground}{figbackground-#1}% !!!
}
\setbeamertemplate{blocks}[rounded][shadow=true]
\setbeamercolor{block title}{fg=white,bg=inriaRed1}
\setbeamerfont{block title}{parent=frametitle,size=\small}
\setbeamerfont{section in toc}{parent=structure, size=\large}

% %remove the icon
% \setbeamertemplate{bibliography item}{}

% %remove line breaks
% \setbeamertemplate{bibliography entry title}{}
% \setbeamertemplate{bibliography entry location}{}
% \setbeamertemplate{bibliography entry note}{}

\usepackage{listings}
\usepackage{amsmath,amssymb}
% \usepackage{courier}

\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{decorations.pathreplacing,angles,quotes}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{calc}
\usetikzlibrary{arrows,patterns}
\usetikzlibrary{intersections}
\usetikzlibrary{tikzmark,fit,shapes.geometric}

\definecolor{myviolet}{rgb}{0.6,0.0,0.65}
\definecolor{myblue}{rgb}{0.1,0.0,0.8}
\definecolor{mygreen}{rgb}{0.1,0.8,0.0}
\definecolor{myred}{rgb}{0.8,0.0,0.0}

\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{lightblue}{rgb}{0,0.5,0.5}
\definecolor{dkgreen}{rgb}{0,0.4,0}
\definecolor{dk2green}{rgb}{0.4,0,0}
\definecolor{dkviolet}{rgb}{0.6,0,0.8}
\definecolor{shadethmcolor}{rgb}{0.9, 0.9,1}

\let\L=\lstinline
\def\lstlanguagefiles{defManSSR.tex}
\lstset{language=SSR}
\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
\lstset{framextopmargin=50pt,frame=bottomline}
\lstset{moredelim=[is][\color{red}\bfseries\ttfamily\underbar]{|*}{*|}}
\lstset{moredelim=[is][\color{blue}\bfseries\ttfamily]{/*}{*/}}

\def\mathcomp{{\small\textsc{Mathematical Components}}}
\def\mydef#1{{\sl #1}}
\def\myem#1{{\bf\sl \color{red}{#1}}}
% \def\myem#1{\emph{#1}}

\newcommand{\cvg}[2]{#1 \rightarrow #2}
\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\scal}[2]{\left\langle #1, #2 \right\rangle}
\newcommand\IR{\ensuremath{\mathbb{R}}}
\newcommand\IC{\ensuremath{\mathbb{C}}}
\newcommand\nat{\ensuremath{\mathbb{N}}}
\newcommand\IN{\nat}

\renewcommand{\leq}{\leqslant}

\def\eps{\varepsilon}
\def\lip{LaSalle's invariance principle}
\def\o{\circ}
\def\coq{{\sc Coq}}
\def\ssr{{\sc SSReflect}}
\def\mathcomp{{\sc Mathematical Components}}
\def\coquelicot{{\sc Coquelicot}}
\def\isahol{{\sc Isabelle/HOL}}
\def\hol4{{\sc HOL4}}
\def\hollight{{\sc HOL Light}}
\def\lean{{\sc Lean}}
\def\pvs{{\sc PVS}}
\def\mizar{{\sc Mizar}}
\def\analysis{{\sc Mathematical Components Analysis}}
\def\plim{\Gamma^{+}}
\def\loc{\textrm{locally}}
\def\ball{\textrm{ball}}

\def\us{\char`\_}

\def\newterm#1{{\sl #1}} % to introduce new terms such as ``entourage'', etc.
\def\eg{\textit{e.g.}} % NB(rei): btw, the lncs authors' instructions uses the american convention (``e.g.,'' in roman)
\def\ie{\textit{i.e.}} % NB(rei): for uniformity (there was one with \em instead of \it)
\def\subset{\subseteq}
\def\supset{\supseteq}

% \makeatletter
% \renewenvironment{table}%
%   {\renewcommand\familydefault\sfdefault
%    \@float{table}}
%   {\end@float}
% \makeatother

\newcommand\doubleRule{\toprule\toprule}
\newcommand\doublerule{\toprule\specialrule{\heavyrulewidth}{\doublerulesep}{0.95em}}

\newcommand\coqPR[1]{\coq{} PR \href{https://github.com/coq/coq/issues/#1}{\##1}}
\newcommand\coqIssue[1]{\coq{} Issue \href{https://github.com/coq/coq/issues/#1}{\##1}}

\begin{document}

\title[Hierarchy Builder]{{Hierarchy Builder:}\\ {Algebraic
    hierarchies} \\ {made easy} {in {\sc Coq} with {\sc Elpi}}}
\subtitle{FSCD 2020}

\author[\underline{Cohen}, Sakaguchi, Tassi]{
  \underline{Cyril Cohen \textit{(Inria)}}, Kazuhiko Sakaguchi, Enrico
Tassi}

% \institute[]{\inst{1} Inria, France, \inst{2} University of Tsukuba,
% Japan}

\date{June 2nd, 2020}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}[fragile]
  \frametitle{Structures in Mathematics}

  \begin{itemize}
  \item A \textbf{carrier} in Set / Type,
  \item A set of \textbf{constants} in the carrier, and \textbf{operations},
  \item Proofs of the \textbf{axioms} of the structure
  \end{itemize}

  \pause

\begin{lstlisting}
Record is_ring R := mk_ring {
    zero : R; add : R -> R -> R; opp : R -> R;
    one : R;  mul : R -> R -> R;
    addrA : associative add;
    addrC : commutative add;
    add0r : left_id zero add;
    addNr : left_inverse zero opp add;
    mulrA : associative mul;
    mul1r : left_id one mul;
    mulr1 : right_id one mul;
    mulrDl : left_distributive mul add;
    mulrDr : right_distributive mul add;
  }.
\end{lstlisting}

\end{frame}

\begin{frame}
  \frametitle{Implementations in DTT (unbundled classes)}

\end{frame}

\begin{frame}
  \frametitle{Implementations in DTT (semi-bundled classes)}

\end{frame}

\begin{frame}
  \frametitle{Implementations in DTT (bundled record)}

\end{frame}

\begin{frame}
  \frametitle{Implementations in DTT (packed classes)}

\end{frame}

\begin{frame}
  \frametitle{Implementations in proof assistants}

  A variety of representations:

  % more details comming
  \begin{itemize}
  \item {\sc Coq/Mathcomp}:     Classes packed in (canonical) structures
  \item {\sc Coq/Math-Classes}: Fully unbundled (typeclass) records $+$
                                special case for varieties.
  \item {\sc Lean/Mathlib}:     Variably-bundled (typeclass) records.
  \item {\sc Agda}:             bundled and semi-bundled records.
  \end{itemize}
  \vfill
  \pause

  Many other possibilities:
  \begin{itemize}
  \item Modules a la Ocaml \emph{(not first class!)},
  \item Fully-bundled typeclasses \emph{(bad!)},
  \item Telescopes \emph{(bad!)},
  \item Records without inference \emph{(tedious!)}, ...
  \end{itemize}
  \vfill
  \pause

\center{  \textbf{Representations work hand in hand with an inference
  mechanism.}}

\end{frame}

\begin{frame}
  \frametitle{Inference of structures}

  \begin{itemize}
  \item {\sc Coq/Mathcomp}:     canonical instances $+$ coercions $+$ checker
  \item {\sc Coq/Math-Classes}: typeclass instances $+$ hint extern
  \item {\sc Lean/Mathlib}:     typeclass instances $+$ linter
  \item {\sc Agda}:             \L{open} and renaming directives
  \end{itemize}

  \vfill
  \pause
  None of these encoding are straightforward, they all need
  expert knowledge or checkers/linters, and additionally

  \begin{itemize}
  \item Some encodings are unnecessarily verbose,
  \item Some known design problems might be detected too late
    (e.g. priority of instance, typeclass indexing, forgetful
    inheritance, etc)
  \end{itemize}

  % TODO: maybe insert here showing various encoding artifacts

  \vfill
  \pause
  \center{\textbf{Hierarchy Builder provides a DSL!}}
\end{frame}

\begin{frame}
  \frametitle{Hierarchies in formalization}

  Purpose:
  \begin{itemize}
  \item factor theorems across instances, using the \textbf{theory} of
    each structure,
  \item \textbf{automatically find} which structures hold on which types.
  \end{itemize}

  \pause
  \vfill

  Requirements:
  \begin{itemize}
  \item declare a \textbf{new instance},
  \pause
  \item declare a \textbf{new structure}
    \begin{itemize}
    \item above, below or in the middle
    \item handle diamonds (e.g. monoid, group, commutative or not),
    \item by amending existing code, or not,
    \end{itemize}
  \pause
  \item \textbf{predictability} of inferred instance,
  \pause
  \item \textbf{robustness} of user code with regard to \textit{new declarations}.
  \end{itemize}

\end{frame}

\begin{frame}
  \frametitle{Hierarchy Builder in theory}

  \begin{enumerate}
  \item \textbf{Hierarchy Builder provides a DSL to generates/extends a
      hierarchy from minimal input.}
    \vfill
  \item \textbf{Hierarchy Builder lets you amend a hierarchy without
      breaking your code.}
  \end{enumerate}

  \vfill

  That way, Hierarchy Builder is fully in control of the design
  patterns used throughout your library.
\end{frame}

\begin{frame}
  \frametitle{Hierarchy Builder in practice}

  \begin{itemize}
  \item Hierarchy Builder generates/extends a hierarchy using \mathcomp{}
    packed class methodology.
    \vfill
  \item Hierarchy Builder enforces a discipline of \emph{mixins} and
    \emph{factories} to make client code robust to hierarchy changes.
    \vfill
  \item Hierarchy Builders lets us encode built-in safety measures
    (e.g. detection of overlapping instances)
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Structures relating to each other}

  Examples:

  \begin{itemize}
  \item Monoid $\leftarrow$ Group $\leftarrow$ Ring $\leftarrow$ Field
    $\leftarrow$ ...

  \item Euclidean Spaces $\rightarrow$ Normed Spaces $\rightarrow$
    Complete Space $\rightarrow$ Metric Spaces $\rightarrow$ Topological
    Spaces $\rightarrow$ ...

  \end{itemize}

  \pause
  \vfill{}

  \textbf{Going through arrows must be automated.}

  \pause
  \vfill{}

  Arrows represent both
  \begin{itemize}
  \item Extensions: add operations, axioms or combine structures
  \item Entailment/Induction/Deduction/Generalization.
  \end{itemize}

\end{frame}

\begin{frame}
  \frametitle{More examples}

\begin{tikzpicture}[line width=1pt, structure/.style={draw, rounded corners=1mm, fill=white}, new/.style={preaction={fill, white}, pattern=crosshatch dots, pattern color=lightgray}, scale=0.7]
 \useasboundingbox (-6, -.5) rectangle (12, 10.5);
 % Replace `fill` with `pattern`s for the final version.
 %\draw[black!50, pattern=north east lines, pattern color=red!50, rounded corners=15mm] (-7.5, -.5) -- (-7.5, 7) -- (-4, 7) -- (2, 3) -- (2, -.5) -- cycle;
 %\draw[black!50, pattern=north west lines, pattern color=blue!50, rounded corners=15mm] (-8, -.5) -- (2, 4) -- (-4, 8) -- (0, 10) -- (10, 5) -- (10, -.5) -- cycle;
 %
 \node at (-1, 1) {\Large \begin{tabular}{c}''Calculus''\\structures\end{tabular}};
 \node at (6, 1) {\Large \begin{tabular}{c}''Algebraic''\\structures\end{tabular}};
 %
 \node[structure] (PartialOrder)          at (9,  9) {\L+PartialOrder+};
 \node[structure] (Lattice)         at (9,  7.67) {\L+Lattice+};
 \node[structure] (TotalOrder)           at (9,  6.33) {\L+TotalOrder+};
 %
 \node[structure] (AddGroup)             at (3,  9.5) {\L+AddGroup+};
 \node[structure] (Lmodule)              at (2, 7.3) {\L+Lmodule+};
 \node[structure] (Rings)                at (5,  8) {\L+(Com)(Unit)Ring+};
 \node[structure] (IntegralDomain)       at (5,  6.5) {\L+IntegralDomain+};
 \node[structure] (Field)                at (2,  5) {\L+Field+};
 \node[structure] (OrderedDomain)        at (8,  5) {\L+OrderedDomain+};
 \node[structure] (OrderedField)         at (5,  4) {\L+OrderedField+};
 \node[structure] (RealClosedField)      at (3,  3) {\L+RealClosedField+};
 \node[structure] (ArchimedeanField)     at (8,  3) {\L+ArchimedeanField+};
 %
 \node[structure] (TopologicalSpace)          at (-4, 9) {\L+TopologicalSpace+};
 \node[structure] (UniformSpace)              at (-4, 7.5) {\L+UniformSpace+};
 \node[structure] (Complete)             at (-5, 5) {\L+Complete+};
 \node[structure] (NormedAddGroup)       at (-2,  6.1) {\L+NormedAddGroup+};
 \node[structure] (NormedModule)         at (-1, 4.5) {\L+NormedModule+};
 \node[structure] (CompleteNormedModule) at (-3, 3) {\L+CompleteNormedModule+};
 %
 \draw[<-] (PartialOrder) to (Lattice);
 \draw[<-] (Lattice) to (TotalOrder);
 \draw[<-] (TopologicalSpace) to (UniformSpace);
 \draw[<-] (UniformSpace) to (Complete);
 \draw[<-] (UniformSpace) to (NormedAddGroup);
 \draw[<-] (Complete) to (CompleteNormedModule);
 \draw[<-] (Lmodule) to (NormedModule);
 \draw[<-] (NormedAddGroup) to (NormedModule);
 \draw[<-] (NormedModule) to (CompleteNormedModule);
 \draw[<-] (AddGroup) to (Lmodule);
 \draw[<-] (AddGroup) to (NormedAddGroup);
 \draw[<-] (AddGroup) to (Rings);
 \draw[<-] (Rings) to (IntegralDomain);
 \draw[<-] (IntegralDomain) to (OrderedDomain);
 \draw[<-] (NormedAddGroup) to (OrderedDomain);
 \draw[<-] (TotalOrder) to (OrderedDomain);
 \draw[<-] (IntegralDomain) to (Field);
 \draw[<-] (Field) to (OrderedField);
 \draw[<-] (OrderedDomain) to (OrderedField);
 \draw[<-] (OrderedField) to (RealClosedField);
 \draw[<-] (OrderedField) to (ArchimedeanField);

 % \draw[opacity=.5, black, fill=red!50, rounded corners=15mm] (-7.5, -.5) -- (-7.5, 7) -- (-4, 7) -- (2.5, 3) -- (2, -.5) -- cycle;
 % \draw[opacity=.5, black, fill=blue!50, rounded corners=15mm] (-8, -.5) -- (1.5, 4) -- (-4.3, 8) -- (0, 10) -- (10.5, 5.5) -- (10, -.5) -- cycle;
 \end{tikzpicture}

\end{frame}

\begin{frame}
  \frametitle{Structure extension vs Structure entailment}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{center}
        {\bf \emph{Structure extension}}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        {\bf \emph{Structure extension}}
      \end{center}
    \end{column}
  \end{columns}
  \vfill
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item \textbf{Compositional}: no need to start from scratch every time.
        (E.g. the product of two groups is a group)
        \pause\vfill
      \item \textbf{Noisy}: changes the internal definition of a structure.
        (E.g. defining a commutative monoid from a monoid, one gets an unnecessary axiom),
        \pause\vfill
      \item \textbf{Non-robust}: breakage of user code when adding new
        intermediate structures, \pause\vfill
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item \textbf{More flexible}: no need to cut structures
        into small bits,
        \pause\vfill
      \item \textbf{Robust}: lets us fix the operations and axioms
        once and for all.
        \pause\vfill
      \item \textbf{Not suitable for inference}: Major breakage when
        arbitrary entailment is automatic. (cf IJCAR \emph{Formalizing
          functional analysis structures in dependent type theory)}
        \pause\vfill
      \end{itemize}
    \end{column}
  \end{columns}
  \vfill

\end{frame}

\begin{frame}
  \frametitle{HB Design}

  The best of two the worlds:
  \begin{itemize}
  \item \textbf{Extension}, through \emph{mixins} for \textbf{internal declaration} and
    \textbf{automatic inference}

  \item \textbf{Entailment}, through \emph{factories} for any other use.

    Factories require mixins and can produces others.
    (e.g. a full axiomatic can provide all the pieces)
  \end{itemize}
  \vfill

  Five commands:
  \begin{enumerate}
  \item \L{HB.mixin     Record axioms T := \{..\}.}
  \item \L{HB.factory   Record axioms T := \{..\}.}
  \item \L{HB.builder   Context       T (f : axioms T). ... HB.end.}
  \item \L{HB.structure Definition    S := \{ T & axioms T \}}
  \item \L{HB.instance  Definition    f := Axioms X ...}
  \end{enumerate}
  \vfill
  \begin{center}
    \url{https://github.com/math-comp/hierarchy-builder}
  \end{center}

\end{frame}

\begin{frame}[fragile]
  \frametitle{A very short example}

\begin{lstlisting}
HB.mixin Record is_monoid (M : Type) := {
  zero : M;
  add : M -> M -> M;
  addrA : associative add;   (* add is associative. *)
  add0r : left_id zero add;  (* zero is the neutral *)
  addr0 : right_id zero add; (*    element wrt add. *)
}.
HB.structure Definition Monoid := { M & is_monoid M }.

HB.instance Definition Z_Monoid_axioms : is_monoid Z :=
  is_monoid.Build Z 0%Z Z.add
    Z.add_assoc Z.add_0_l Z.add_0_r.
\end{lstlisting}
  \vfill

  See a full demo during the Coq workshop.
\end{frame}

\begin{frame}[fragile]
  \frametitle{Breaking down monoid}

  We split the monoid structure into a semi-group and a monoid
  \begin{lstlisting}
HB.mixin Record is_semigroup (S : Type) := {
  add : S -> S -> S;
  addrA : associative add;   (* add is associative. *)
}.
HB.structure Definition SemiGroup :=
  { S & is_semigroup S }.

HB.mixin Record monoid_of_semigroup (M : Type) := {
  zero : M;
  add0r : left_id zero add;  (* zero is the neutral *)
  addr0 : right_id zero add; (*    element wrt add. *)
}.
HB.structure Definition Monoid :=
  { M & monoid_of_semigroup M }.
\end{lstlisting}
\center{\textbf{But we must provide \L{is_monoid} again.}}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Recovering \L{is_monoid}}

  It becomes a \emph{factory}.
\begin{lstlisting}
HB.factory Record is_monoid (M : Type) := {
  zero : M;
  add : M -> M -> M;
  addrA : associative add;
  add0r : left_id zero add;
  addr0 : right_id zero add;
}.

HB.builder Context (M : Type) (f : is_monoid M) :=
  HB.instance Definition is_monoid_semigroup :
    is_semigroup M := ...
  HB.instance Definition is_monoid_monoid :
    monoid_of_semigroup M := ...
HB.end
\end{lstlisting}
\end{frame}

\begin{frame}
  \frametitle{Meta programming in {\sc Coq-ELPI}}

\end{frame}

\begin{frame}
  \frametitle{Conclusion}

  \begin{itemize}
  \item High-level commands to declare structures and instances, \\
    easy to use.
    \vfill
  \item Predictable outcome of inference,
    \vfill
  \item Takes into account the evolution of knowledge
    \begin{itemize}
    \item which is formalized, and
    \item which the user has.
    \end{itemize}
    The two knowledge do not need to be correlated.
    \vfill
  \item Robustness with regard to new declaration \emph{and even changes of
    internal implementation}.
  \end{itemize}

  \vfill\pause
  Also, \textsc{Coq-ELPI} turned out to be a very comfortable
  meta-programming language for this.

\end{frame}

\begin{frame}
  \frametitle{Future work on Hierarchy Builder}
  \begin{itemize}
  \item Adding support for parameters. (debugging phase in progress)
    \vfill
  \item Detecting and fixing competing inheritance paths.
    \vfill
  \item Generating hierarchies of morphisms from structures.
    \vfill
  \item Generating hierarchies of subobjects from structures.
    \vfill
  \item Supporting multiple instances on the same carrier.
    \vfill
  \item Replacing all uses in math-comp and extensions.
    \vfill
  \item Get better error messages.
  \end{itemize}
\end{frame}

\end{document}

### 2020-07-01 talk-fscd/defManSSR.tex
update talk (pseudo-final version)

%=======================================================================
% Définit le style ssr Gallina pour les listings (Assia Mahboubi 2007)

\lstdefinelanguage{SSR} {

% Anything betweeen $ becomes LaTeX math mode
mathescape=true,						
% Comments may or not include Latex commands
texcl=false,

% Vernacular commands
morekeywords=[1]{
Context, Section, Module, End, Require, Import, Export, Coercions,
Variable, Variables, Parameter, Parameters, Axiom, Hypothesis, Hypotheses,
Notation, Local, Tactic, Reserved, Scope, Open, Close, Bind, Delimit,
Definition, Let, Ltac, Fixpoint, CoFixpoint, Add, Morphism, Relation,
Implicit, Arguments, Set, Unset, Contextual, Strict, Prenex, Implicits,
Inductive, CoInductive, Record, Structure, Canonical, Coercion,
Theorem, Lemma, Corollary, Proposition, Fact, Remark, Example,
Proof, Goal, Save, Qed, Defined, Hint, Resolve, Rewrite, View,
Search, Check, Show, Print, Printing, All, Graph, Projections, inside,
outside, Locate, Maximal, Types, Type, Class, Instance,
HB, mixin, factory, builders, end, structure, instance},
% HB.mixin, HB.factory, HB.builders, HB.structure, HB.instance,},

% Gallina
morekeywords=[2]{forall, exists, exists2, fun, fix, cofix, struct,
      match, with, end, as, ' in ', return, let, if, is, then, else,
      for, of, nosimpl},

% Sorts
morekeywords=[3]{Set, Prop, Type},

% Various tactics, some are std Coq subsumed by ssr, for the manual purpose
morekeywords=[4]{
         pose, set, move, case, elim, apply, clear,
            hnf, intro, intros, generalize, rename, pattern, after,
	    destruct, induction, using, refine, inversion, injection,
         rewrite, congr, unlock, compute, vm_compute, group, ring, field,
            replace, fold, unfold, change, cutrewrite, simpl,
         have, suff, wlog, suffices, without, loss, nat_norm,
            assert, cut, trivial, revert, bool_congr, nat_congr,
	 symmetry, transitivity, auto, split, left, right, autorewrite},

% Terminators
morekeywords=[5]{
         by, done, exact, reflexivity, tauto, romega, omega,
         assumption, solve, contradiction, discriminate},

% Control
morekeywords=[6]{do, last, first, try, idtac, repeat},

% Various symbols
% For the ssr manual we turn off the prettyprint of formulas
% literate=	
%	{->}{{$\rightarrow\,$}}2		
% 	{->}{{\tt ->}}3
%	{<-}{{$\leftarrow\,$}}2	
% 	{<-}{{\tt <-}}2			
% 	{>->}{{$\mapsto$}}3		
% 	{<=}{{$\leq$}}1			
% 	{>=}{{$\geq$}}1		
% 	{<>}{{$\neq$}}1
% 	{/\\}{{$\wedge$}}2	
% 	{\\/}{{$\vee$}}2 		
% 	{<->}{{$\leftrightarrow\;$}}3
% 	{<=>}{{$\Leftrightarrow\;$}}3
% 	{:nat}{{$~\in\mathbb{N}$}}3	
%	{fforall\ }{{$\forall_f\,$}}1		
%	{forall\ }{{$\forall\,$}}1		
%	{exists\ }{{$\exists\,$}}1
% 	{negb}{{$\neg$}}1
% 	{spp}{{:*:\,}}1
% 	{~}{{$\sim$}}1
% 	{\\in}{{$\in\;$}}1
% 	{/\\}{$\land\,$}1
% 	{:*:}{{$*$}}2
%	{=>}{{$\,\Rightarrow\ $}}1
% 	{=>}{{\tt =>}}2
% 	{:=}{{{\tt:=}\,\,}}2
% 	{==}{{$\equiv$}\,}2
% 	{!=}{{$\neq$}\,}2
% 	{^-1}{{$^{-1}$}}1
% 	{elt'}{elt'}1
%       {=}{{\tt=}\,\,}2
%       {+}{{\tt+}\,\,}2,

% Comments delimiters, we do turn this off for the manual
%comment=[s]{(*}{*)},

% Spaces are not displayed as a special character
showstringspaces=false,

% String delimiters
morestring=[b]",
morestring=[d]’,

% Size of tabulations
tabsize=3,

% Enables ASCII chars 128 to 255
extendedchars=true,

% Case sensitivity
sensitive=true,

% Automatic breaking of long lines
breaklines=true,

% Default style fors listings
basicstyle=\ttfamily,

% Position of captions is bottom
captionpos=b,

% Full flexible columns 
% columns=[l]fullflexible,

% Style for (listings') identifiers
identifierstyle={\ttfamily\color{black}},
% Note : highlighting of Coq identifiers is done through a new
% delimiter definition through an lstset at the begining of the
% document. Don't know how to do better.

% Style for declaration keywords
keywordstyle=[1]{\ttfamily\color{dkviolet}},

% Style for gallina keywords
keywordstyle=[2]{\ttfamily\color{dkgreen}},

% Style for sorts keywords
keywordstyle=[3]{\ttfamily\color{lightblue}},

% Style for tactics keywords
keywordstyle=[4]{\ttfamily\color{dkblue}},

% Style for terminators keywords
keywordstyle=[5]{\ttfamily\color{red}},

%Style for iterators
keywordstyle=[6]{\ttfamily\color{dkviolet}},

% Style for strings
stringstyle=\ttfamily,

% Style for comments
commentstyle=\rmfamily,

}

### 2020-07-01 talk-fscd/main.tex
update talk (pseudo-final version)

@@
 \usepackage{apalike}
+
 % \usepackage{bibentry}
@@
 \def\lstlanguagefiles{defManSSR.tex}
+\lstdefinelanguage{Prolog}
+{
+    basicstyle = \ttfamily\color{blue},
+    moredelim = [s][\color{black}]{(}{)},
+    literate =
+        {:-}{{\textcolor{black}{:-}}}2
+        {,}{{\textcolor{black}{,}}}1
+        {.}{{\textcolor{black}{.}}}1
+        {o:}{{\textcolor{black}{o:}}}1
+        {i:}{{\textcolor{black}{o:}}}1
+}
+
 \lstset{language=SSR}
@@
 
+  Standard definition:
   \begin{itemize}
@@
   \pause
+  \vfill
+
+  E.g. an (additive) monoid is given by
+  \begin{itemize}
+  \item a carrier \L{T : Type},
+  \item a constant \L{zero : T} and a binary operation %
+    \L{add : T -> T -> T}
+  \item three axioms: \\
+    associativity of the addition, left and right neutrality of zero.
+  \end{itemize}
+
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Implementations in DTT (unbundled classes)}
 
 \begin{lstlisting}
-Record is_ring R := mk_ring {
-    zero : R; add : R -> R -> R; opp : R -> R;
-    one : R;  mul : R -> R -> R;
+Class |*is_monoid*| T (zero : T) (add : T -> T -> T) := {
     addrA : associative add;
-    addrC : commutative add;
     add0r : left_id zero add;
-    addNr : left_inverse zero opp add;
-    mulrA : associative mul;
-    mul1r : left_id one mul;
-    mulr1 : right_id one mul;
-    mulrDl : left_distributive mul add;
-    mulrDr : right_distributive mul add;
+    add0r : right_id zero add;
   }.
@@
 
-\begin{frame}
-  \frametitle{Implementations in DTT (unbundled classes)}
+\begin{frame}[fragile]
+  \frametitle{Implementations in DTT (semi-bundled classes)}
+
+\begin{lstlisting}
+Class |*is_monoid*| (T : Type) : Type := {
+    zero  : T;
+    add   : T -> T -> T;
+    addrA : associative add;
+    add0r : left_id zero add;
+    add0r : right_id zero add;
+  }.
+\end{lstlisting}
+
+\vfill
+  \begin{onlyenv}<2>
+\begin{lstlisting}
+Class |*is_group*| (T : Type) : Type :=
+  { zero : T;  add : T -> T -> T; opp : T -> T; .. }.
+\end{lstlisting}
+  \end{onlyenv}
+  \begin{onlyenv}<3>
+\begin{lstlisting}
+Class |*group_of_monoid*| T : |*is_monoid*| T -> Type :=
+   { opp : T -> T; .. }.
+\end{lstlisting}
+  \end{onlyenv}
+
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Implementations in DTT (bundled record)}
+
+\begin{lstlisting}
+Structure |*monoidType*| : Type := {
+    sort  :> Type;
+    zero  : sort;
+    add   : sort -> sort -> sort;
+    addrA : associative add;
+    add0r : left_id zero add;
+    add0r : right_id zero add;
+  }.
+\end{lstlisting}
+
+\end{frame}
+\begin{frame}[fragile]
+  \frametitle{Implementations in DTT \\ \hfill (simplified packed classes)}
+
+\begin{lstlisting}
+Class |*is_monoid*| (T : Type) : Type := {
+    zero  : T;
+    add   : T -> T -> T;
+    addrA : associative add;
+    add0r : left_id zero add;
+    add0r : right_id zero add;
+  }.
+\end{lstlisting}
+
+\vfill
+
+\begin{lstlisting}
+Structure |*monoidType*| : Type := {
+    sort  :> Type;
+    class : |*is_monoid*| sort;
+  }.
+\end{lstlisting}
+
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Implementations in DTT (packed classes)}
+
+\begin{lstlisting}
+Record |*is_monoid*| (T : Type) : Type := { zero ; ..}.
+
+Structure |*monoidType*| : Type :=
+  { sort :> Type;     class : |*is_monoid*| sort }.
+\end{lstlisting}
+
+\begin{lstlisting}
+Record |*group_of_monoid*| T : |*is_monoid*| T -> Type := ...
+
+Record |*is_group*| (T : Type) := {
+    monoid_of_group : |*is_monoid*| T;
+    group_of_group  : |*group_of_monoid*| T monoid_of_group
+  }.
+
+Structure |*groupType*| : Type :=
+  { sort :> Type;     class : |*is_group*| sort }.
+\end{lstlisting}
 
@@
 \begin{frame}
-  \frametitle{Implementations in DTT (semi-bundled classes)}
+  \frametitle{Implementation in DTT (other)}
 
-\end{frame}
-
-\begin{frame}
-  \frametitle{Implementations in DTT (bundled record)}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Implementations in DTT (packed classes)}
+  Many other possibilities:
+  \vfill
+  \begin{itemize}
+  \item Modules a la {\sc OCaml} \emph{(not first class in {\sc Coq}!)},
+  \vfill
+  \item Fully-bundled typeclasses \emph{(bad!)},
+  \vfill
+  \item Telescopes \emph{(bad!)},
+  \vfill
+  \item Records without inference \emph{(tedious!)},
+  \vfill
+  \item ...
+  \end{itemize}
 
@@
 
-  A variety of representations:
+  The variety of representations is out there!
 
@@
   \begin{itemize}
-  \item {\sc Coq/Mathcomp}:     Classes packed in (canonical) structures
-  \item {\sc Coq/Math-Classes}: Fully unbundled (typeclass) records $+$
-                                special case for varieties.
-  \item {\sc Lean/Mathlib}:     Variably-bundled (typeclass) records.
-  \item {\sc Agda}:             bundled and semi-bundled records.
+  \item {\sc Coq/Mathcomp}:     Packed classes\only<3>{\alert{ inside canonical
+    structures}}.
+\item {\sc Coq/Math-Classes}: Fully unbundled \alt<3>{\alert{type classes}}{records} \\
+    \hfill ($+$ special case for varieties).
+  \item {\sc Lean3/Mathlib}:     Semi-bundled \alt<3>{\alert{type classes}}{records}.
+  \item {\sc Agda}:             Bundled and semi-bundled records.
   \end{itemize}
@@
 
-  Many other possibilities:
-  \begin{itemize}
-  \item Modules a la Ocaml \emph{(not first class!)},
-  \item Fully-bundled typeclasses \emph{(bad!)},
-  \item Telescopes \emph{(bad!)},
-  \item Records without inference \emph{(tedious!)}, ...
-  \end{itemize}
   \vfill
@@
 
-\center{  \textbf{Representations work hand in hand with an inference
+\center{\textbf{Representations work hand in hand with an inference
   mechanism.}}
@@
 \begin{frame}
-  \frametitle{Inference of structures}
+  \frametitle{More than ``just records''}
 
   \begin{itemize}
-  \item {\sc Coq/Mathcomp}:     canonical instances $+$ coercions $+$ checker
-  \item {\sc Coq/Math-Classes}: typeclass instances $+$ hint extern
-  \item {\sc Lean/Mathlib}:     typeclass instances $+$ linter
-  \item {\sc Agda}:             \L{open} and renaming directives
+  \item {\sc Coq/Mathcomp}:     canonicals \alert<1>{$+$ heavy boilerplate $+$ checker}
+  \item {\sc Coq/Math-Classes}: type classes \alert<1>{$+$ boilerplate $+$ hints}
+  \item {\sc Lean3/Mathlib}:     type classes \alert<1>{$+$ priorities $+$ linter}
+  \item {\sc Agda}:             records \alert<1>{$+$ open and renaming directives}
   \end{itemize}
@@
   \pause
-  None of these encoding are straightforward, they all need
-  expert knowledge or checkers/linters, and additionally
+  \vfill
 
+  None of these encoding are straightforward:
   \begin{itemize}
-  \item Some encodings are unnecessarily verbose,
-  \item Some known design problems might be detected too late
+  \item they all need expert knowledge and/or checkers/linters,
+  \item some encodings are unnecessarily verbose,
+  \item some known design problems might be detected too late
     (e.g. priority of instance, typeclass indexing, forgetful
@@
   \begin{itemize}
-  \item factor theorems across instances, using the \textbf{theory} of
-    each structure,
+  \item \textbf{factor theorems}, using the \emph{theory} of each structure,
   \item \textbf{automatically find} which structures hold on which types.
@@
     \pause
+  \item provide \textbf{several ways} to instantiate them
+  \pause
   \item \textbf{predictability} of inferred instance,
@@
 \begin{frame}
-  \frametitle{Hierarchy Builder in theory}
+  \frametitle{Hierarchy Builder in two bullets}
 
   \begin{enumerate}
-  \item \textbf{Hierarchy Builder provides a DSL to generates/extends a
+  \item \textbf{Hierarchy Builder provides a DSL to generate and extend a
       hierarchy from minimal input.}
@@
 
-  That way, Hierarchy Builder is fully in control of the design
-  patterns used throughout your library.
+  Hierarchy Builder takes care of the design pattern used throughout
+  your library.
 \end{frame}
@@
       \begin{center}
-        {\bf \emph{Structure extension}}
+        {\bf \emph{Structure entailment}}
       \end{center}
@@
       \begin{itemize}
-      \item \textbf{More flexible}: no need to cut structures
+      \item \textbf{Flexible}: no need to cut structures
         into small bits,
@@
       \item \textbf{Not suitable for inference}: Major breakage when
-        arbitrary entailment is automatic. (cf IJCAR \emph{Formalizing
-          functional analysis structures in dependent type theory)}
-        \pause\vfill
+        arbitrary entailment is automatic. (cf IJCAR \emph{Competing
+          Inheritance Paths in Dependent Type Theory)} \pause\vfill
       \end{itemize}
@@
   \begin{itemize}
-  \item \textbf{Extension}, through \emph{mixins} for \textbf{internal declaration} and
-    \textbf{automatic inference}
+  \item \textbf{Extension}, through \emph{mixins} for \textbf{automatic inference}
 
-  \item \textbf{Entailment}, through \emph{factories} for any other use.
+  \item \textbf{Entailment}, through \emph{factories} for
+    \textbf{smart instantiation}
 
-    Factories require mixins and can produces others.
-    (e.g. a full axiomatic can provide all the pieces)
+    % Factories require mixins and can produces others.
+    % (e.g. a full axiomatic can provide all the pieces)
   \end{itemize}
+  \vspace{2em}
+  \pause
   \vfill
 
-  Five commands:
+  Five primitives:
   \begin{enumerate}
-  \item \L{HB.mixin     Record axioms T := \{..\}.}
-  \item \L{HB.factory   Record axioms T := \{..\}.}
-  \item \L{HB.builder   Context       T (f : axioms T). ... HB.end.}
-  \item \L{HB.structure Definition    S := \{ T & axioms T \}}
-  \item \L{HB.instance  Definition    f := Axioms X ...}
+  \item \L{HB.mixin     Record |*<mixin name>*|$ $ T of |*<dependencies>*| := \{..\}.}
+    \vspace{1em}
+  \item \L{HB.factory   Record |*<factory name>*|$ $ T of |*<dependencies>*| := \{..\}.}
+  \item \L{HB.builders   Context       T (f : |*<factory name>*|$ $ T). ... HB.end.}
+    \vspace{1em}
+  \item \L{HB.structure Definition    |*<structure name>*|$ $ :=} \L{\{ T & |*<dependencies>*| \}}
+  \item \L{HB.instance  Definition    |*<name>*|$ $ : |*<axioms name>*| |*<type>*| := ...}
   \end{enumerate}
   \vfill
+
+  \pause
   \begin{center}
-    \url{https://github.com/math-comp/hierarchy-builder}
+    see \url{https://github.com/math-comp/hierarchy-builder}
   \end{center}
@@
   \frametitle{A very short example}
+{\scriptsize\url{https://github.com/math-comp/hierarchy-builder/tree/master/FSCD2020_talk}}
 
 \begin{lstlisting}
-HB.mixin Record is_monoid (M : Type) := {
+HB.mixin Record |*is_monoid*| (M : Type) := {
   zero  : M;
@@
 }.
-HB.structure Definition Monoid := { M & is_monoid M }.
+HB.structure Definition |*Monoid*| := { M & |*is_monoid*| M }.
 
-HB.instance Definition Z_Monoid_axioms : is_monoid Z :=
-  is_monoid.Build Z 0%Z Z.add
+HB.instance Definition Z_is_monoid : |*is_monoid*| Z
+  := |*is_monoid*|.Build Z 0%Z Z.add
     Z.add_assoc Z.add_0_l Z.add_0_r.
@@
 
-  See a full demo during the Coq workshop.
+  See a bigger demo during the Coq workshop.
 \end{frame}
@@
 \begin{lstlisting}
-HB.mixin Record is_semigroup (S : Type) := {
+HB.mixin Record |*is_semigroup*| (S : Type) := {
   add   : S -> S -> S;
-  addrA : associative add;   (* add is associative. *)
+  addrA : associative add;
 }.
-HB.structure Definition SemiGroup :=
-  { S & is_semigroup S }.
-
-HB.mixin Record monoid_of_semigroup (M : Type) := {
+HB.structure Definition |*SemiGroup*| :=
+  { S & |*is_semigroup*| S }.
+\end{lstlisting}
+\begin{lstlisting}
+HB.mixin Record |*monoid_of_semigroup*| (M : Type)
+  & |*is_semigroup*| M := {
   zero  : M;
-  add0r : left_id zero add;  (* zero is the neutral *)
-  addr0 : right_id zero add; (*    element wrt add. *)
+  add0r : left_id  zero add;
+  addr0 : right_id zero add;
 }.
-HB.structure Definition Monoid :=
-  { M & monoid_of_semigroup M }.
+HB.structure Definition |*Monoid*| :=
+  { M & |*monoid_of_semigroup*| M }.
 \end{lstlisting}
-\center{\textbf{But we must provide \L{is_monoid} again.}}
+  \begin{center}
+    \textbf{But we must provide \L{|*is_monoid*|} again.}
+  \end{center}
 \end{frame}
@@
 \begin{frame}[fragile]
-  \frametitle{Recovering \L{is_monoid}}
+  \frametitle{Recovering the lost mixin (\L{is_monoid})}
 
-  It becomes a \emph{factory}.
+  It becomes a \emph{factory} with the \textbf{exact} same contents as before
 \begin{lstlisting}
-HB.factory Record is_monoid (M : Type) := {
+HB.factory Record |*is_monoid*| (M : Type) := {
   zero  : M;
@@
 }.
-
-HB.builder Context (M : Type) (f : is_monoid M) :=
-  HB.instance Definition is_monoid_semigroup :
-    is_semigroup M := ...
-  HB.instance Definition is_monoid_monoid :
-    monoid_of_semigroup M := ...
+HB.builders Context (M : Type) (f : |*is_monoid*| M).
+  HB.instance Definition is_monoid_semigroup
+    : |*is_semigroup*| M := ... (* trivial *)
+  HB.instance Definition is_monoid_monoid
+    : |*monoid_of_semigroup*| M := ... (* trivial *)
 HB.end
+\end{lstlisting}
+  \begin{center}
+    \textbf{Factories can only be used at instantiation time:}
+  \end{center}
+\begin{lstlisting}
+HB.instance Definition Z_is_monoid : |*is_monoid*| Z := ...
 \end{lstlisting}
@@
   \frametitle{Meta programming in {\sc Coq-ELPI}}
+
+  \textbf{{\sc Elpi} is a programming language,}
+  \begin{itemize}
+  \item prolog-like: \textbf{programs and data are clauses}
+  \item with binders, unification and constraints
+    \pause
+  \item capable if representing {\sc Coq} terms in HOAS, its typing
+    judgements, evaluation and unification.
+  \end{itemize}
+    \pause
+    \vfill
+
+  \textbf{{\sc Coq-Elpi} is a plugin for {\sc Coq}} that lets one use {\sc Elpi} as a
+    meta-programming language, \pause in particular
+  \begin{itemize}
+  \item one can write new commands and tactics,
+  \item one can add new definitions, inductive, sections, modules, etc
+    to the environment,
+  \item one can maintain databases across {\sc Coq} files
+  \end{itemize}
+
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Two main HB databases}
+
+  \begin{itemize}
+  \item The predicate \L{from} stores an association between a factory
+    \L{F}, a mixin \L{M} and the term \L{B} that can be used to build
+    mixin \L{M} from factory \L{F}.
+\begin{lstlisting}[language=Prolog, basicstyle=\footnotesize]
+pred from o:factoryname, o:mixinname, o:term.
+\end{lstlisting}
+\vfill
+  \item The predicate {\footnotesize\tt factory-requires} stores an
+    association between a factory and a list of mixins that are
+    pre-requisites to inhabiting this factory.
+\begin{lstlisting}[language=Prolog, basicstyle=\footnotesize]
+pred factory-requires o:factoryname, o:list mixinname.
+\end{lstlisting}
+    e.g. \L{group_of_monoid T} has the prerequisite that \L{T} is a
+    monoid.
+  \end{itemize}
 
@@
     \vfill
-  \item Replacing all uses in math-comp and extensions.
+  \item Replacing all uses in math-comp and extensions. (just started)
     \vfill
-  \item Get better error messages.
+  \item Get better error messages. (work in progress)
   \end{itemize}
@@
 
+\begin{frame}
+  \begin{center}
+   {\huge Thanks! Questions?}
+  \end{center}
+\end{frame}
+
 \end{document}

### 2020-07-01 talk-fscd/main.tex
colors for elpi code

@@
 \usepackage{apalike}
+\usepackage{minted} % pip3 install pygments, pdflatex -shell-escape hb
+\newminted[elpicode]{../elpi.py:ElpiLexer -x}{linenos=false,fontsize=\footnotesize,frame=bottomline}
 
@@
     mixin \L{M} from factory \L{F}.
-\begin{lstlisting}[language=Prolog, basicstyle=\footnotesize]
+\begin{elpicode}
 pred from o:factoryname, o:mixinname, o:term.
-\end{lstlisting}
+\end{elpicode}
 \vfill
@@
     pre-requisites to inhabiting this factory.
-\begin{lstlisting}[language=Prolog, basicstyle=\footnotesize]
+\begin{elpicode}
 pred factory-requires o:factoryname, o:list mixinname.
-\end{lstlisting}
+\end{elpicode}
     e.g. \L{group_of_monoid T} has the prerequisite that \L{T} is a

### 2020-07-01 talk-fscd/main.tex
add Date/Time for the Coq WS
last minute fixes

@@
 \begin{frame}[fragile]
-  \frametitle{Implementations in DTT (unbundled classes)}
+  \frametitle{Implementations in DTT (unbundled classes) \\ \hfill [MSCS2011]}
 
@@
 \begin{frame}[fragile]
-  \frametitle{Implementations in DTT (packed classes)}
+  \frametitle{Implementations in DTT (packed classes) \\ \hfill [TPHOL
+    2009]}
 
@@
   \item {\sc Agda}:             Bundled and semi-bundled records.
+  \item ...
   \end{itemize}
@@
 
-\center{\textbf{Representations work hand in hand with an inference
-  mechanism.}}
+\center{\textbf{Representations work hand in hand with tooling.}}
 
@@
   \begin{itemize}
-  \item {\sc Coq/Mathcomp}:     canonicals \alert<1>{$+$ heavy boilerplate $+$ checker}
+  \item {\sc Coq/Mathcomp}:     canonicals \alert<1>{$+$ boilerplate
+      $+$ validator [IJCAR]}
   \item {\sc Coq/Math-Classes}: type classes \alert<1>{$+$ boilerplate $+$ hints}
@@
 
-  \item Euclidean Spaces $\rightarrow$ Normed Spaces $\rightarrow$
-    Complete Space $\rightarrow$ Metric Spaces $\rightarrow$ Topological
-    Spaces $\rightarrow$ ...
+  \item Normed Space $\rightarrow$ Metric Spaces $\rightarrow$
+    Topological Spaces $\rightarrow$ ...
 
@@
 \begin{frame}
-  \frametitle{More examples}
+  \frametitle{More examples \hfill [IJCAR]}
 
@@
         \pause\vfill
-      \item \textbf{Noisy}: changes the internal definition of a structure.
+      \item \textbf{Noisy} internal definition of a structure.
         (E.g. defining a commutative monoid from a monoid, one gets an unnecessary axiom),
         \pause\vfill
-      \item \textbf{Non-robust}: breakage of user code when adding new
+      \item \textbf{Non-robust} when adding new
         intermediate structures, \pause\vfill
@@
         \pause\vfill
-      \item \textbf{Robust}: lets us fix the operations and axioms
+      \item \textbf{Robust}: we can fix operations and axioms
         once and for all.
@@
 
-  See a bigger demo during the Coq workshop.
+  A bigger demo at the Coq Workshop, Monday 10AM GMT+2.
 \end{frame}
@@
 
-  \textbf{{\sc Elpi} is a programming language,}
+  \textbf{{\sc Elpi} is a programming language,} \hfill [LPAR-20]
   \begin{itemize}
@@
   Also, \textsc{Coq-ELPI} turned out to be a very comfortable
-  meta-programming language for this.
+  meta-programming language for this (approx. 2000 loc).
 
@@
     \vfill
+  \item Replacing all uses in math-comp and extensions. (just started)
+    \vfill
+  \item Get better error messages. (work in progress)
+    \vfill
   \item Detecting and fixing competing inheritance paths.
@@
   \item Supporting multiple instances on the same carrier.
-    \vfill
-  \item Replacing all uses in math-comp and extensions. (just started)
-    \vfill
-  \item Get better error messages. (work in progress)
   \end{itemize}

### 2020-07-02 talk-fscd/main.tex
Fix: TPHOL*s*
small fix

@@
 \begin{frame}[fragile]
-  \frametitle{Implementations in DTT (packed classes) \\ \hfill [TPHOL
+  \frametitle{Implementations in DTT (packed classes) \\ \hfill [TPHOLs
     2009]}
@@
   \begin{itemize}
-  \item {\sc Coq/Mathcomp}:     canonicals \alert<1>{$+$ boilerplate
-      $+$ validator [IJCAR]}
+  \item {\sc Coq/Mathcomp}:     canonicals \alert<1>{\\
+      $+$ heavy boilerplate $+$ validator [IJCAR K.S. paper]}
   \item {\sc Coq/Math-Classes}: type classes \alert<1>{$+$ boilerplate $+$ hints}
